### 2017-06-06 inst/paper/figs.Rmd
init commit

---
title: "F1000 paper figures"
author: "Kent Riemondy RBI"
date: "June 6, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load_libs}
library(valr)
library(tidyverse)
library(cowplot)
```
## Figure 1

```{r, fig1}

# theme to make pretty for publication

pub_theme <- theme(axis.text = element_text(size = 14),
                   plot.title = element_text(size = 14),
                   strip.text = element_text(size = 14))

x <- tibble::tribble(
  ~chrom, ~start, ~end,
  'chr1', 25,     50,
  'chr1', 100,    125
)

y <- tibble::tribble(
  ~chrom, ~start, ~end,
  'chr1', 30,     75
)

a <- bed_glyph(bed_intersect(x, y)) + pub_theme

x <- tibble::tribble(
  ~chrom, ~start, ~end,
  'chr1',      1,      50,
  'chr1',      10,     75,
  'chr1',      100,    120
)

b <- bed_glyph(bed_merge(x)) + pub_theme
plot_grid(a, b, align = 'h', nrow = 1, labels="AUTO", scale = c(.9, .9))
ggsave("glyphs.pdf", width = 8, height = 4)

```

## Figure 2
```{r fig2}
# `valr_example()` identifies the path of example files
bedfile <- valr_example('genes.hg19.chr22.bed.gz')
genomefile <- valr_example('hg19.chrom.sizes.gz')
bgfile  <- valr_example('hela.h3k4.chip.bg.gz')

genes <- read_bed(bedfile, n_fields = 6)
genome <- read_genome(genomefile)
y <- read_bedgraph(bgfile)

# generate 1 bp TSS intervals, `+` strand only
tss <- genes %>%
  filter(strand == '+') %>%
  mutate(end = start + 1)

# 1000 bp up and downstream
region_size <- 1000
# 50 bp windows
win_size <- 50

# add slop to the TSS, break into windows and add a group
x <- tss %>%
  bed_slop(genome, both = region_size) %>%
  bed_makewindows(genome, win_size)

x

# map signals to TSS regions and calculate summary statistics.
res <- bed_map(x, y, win_sum = sum(value, na.rm = TRUE)) %>%
  group_by(.win_id) %>%
  summarize(win_mean = mean(win_sum, na.rm = TRUE),
            win_sd = sd(win_sum, na.rm = TRUE))

res

x_labels <- seq(-region_size, region_size, by = win_size * 5)
x_breaks <- seq(1, 41, by = 5)

sd_limits <- aes(ymax = win_mean + win_sd, ymin = win_mean - win_sd)

p <- ggplot(res, aes(x = .win_id, y = win_mean)) +
  geom_point(size = 0.25) + geom_pointrange(sd_limits, size = 0.1) + 
  scale_x_continuous(labels = x_labels, breaks = x_breaks) + 
  xlab('Position (bp from TSS)') + ylab('Signal') + 
  pub_theme +
  theme_classic()

plot_grid(p, align = 'h', nrow = 1, labels="AUTO")
ggsave("tss.pdf", width = 3.5, height = 2.25)
```

## Figure 3

```{r, message=FALSE}
# Load required packages
library(valr)
library(tidyverse)
library(scales)
library(microbenchmark)
library(stringr)
library(cowplot)

# Set-up test genome
genome <- read_genome(valr_example('hg19.chrom.sizes.gz'))
write_tsv(genome, "genome.txt", col_names=FALSE)

# number of intervals
n <- 1e6
# number of timing reps
nrep <- 10

seed_x <- 1010486
x <- bed_random(genome, n = n, seed = seed_x)
write_tsv(x, "x.bed", col_names=FALSE)
seed_y <- 9283019
y <- bed_random(genome, n = n, seed = seed_y)
write_tsv(y, "y.bed", col_names=FALSE)

res <- microbenchmark(
  # randomizing functions
  bed_random(genome, n = n, seed = seed_x),
  bed_shuffle(x, genome, seed = seed_x),
  # # single tbl functions
  bed_slop(x, genome, both = 1000),
  bed_flank(x, genome, both = 1000),
  bed_merge(x),
  bed_cluster(x),
  bed_complement(x, genome),
  # multi tbl functions
  bed_closest(x, y),
  bed_intersect(x, y),
  bed_map(x, y, .n = n()),
  bed_subtract(x, y),
  # stats
  bed_reldist(x, y),
  bed_jaccard(x, y),
  bed_fisher(x, y, genome),
  #utils 
  bed_makewindows(x, genome, win_size = 100),
  times = nrep,
  unit = 's')

load_file <- function(filename) {
  if(endsWith(filename, ".txt")){
    read_genome(filename)
  } else if (endsWith(filename, ".bed")) {
    read_bed(filename)
  }
}

genome_file <- "genome.txt"
x_bed <- "x.bed"
y_bed  <- "y.bed"

res_hard <- microbenchmark(
  # randomizing functions
  bed_random(load_file(genome_file), n = n, seed = seed_x),
  bed_shuffle(load_file(x_bed), load_file(genome_file), seed = seed_x),
  # # single tbl functions
  bed_slop(load_file(x_bed), load_file(genome_file), both = 1000),
  bed_flank(load_file(x_bed), load_file(genome_file), both = 1000),
  bed_merge(load_file(x_bed)),
  bed_cluster(load_file(x_bed)),
  bed_complement(load_file(x_bed), load_file(genome_file)),
  # multi tbl functions
  bed_closest(load_file(x_bed), load_file(y_bed)),
  bed_intersect(load_file(x_bed), load_file(y_bed)),
  bed_map(load_file(x_bed), load_file(y_bed), .n = n()),
  bed_subtract(load_file(x_bed), load_file(y_bed)),
  # stats
  bed_reldist(load_file(x_bed), load_file(y_bed)),
  bed_jaccard(load_file(x_bed), load_file(y_bed)),
  bed_fisher(load_file(x_bed), load_file(y_bed), load_file(genome_file)),
  #utils 
  bed_makewindows(load_file(x_bed), load_file(genome_file), win_size = 100),
  times = nrep,
  unit = 's')

# covert nanoseconds to seconds
res <- res %>%
  as_tibble() %>%
  mutate(time = time / 1e9)

res_hard <- res_hard %>%
  as_tibble() %>%
  mutate(time = time / 1e9)

# futz with the x-axis
sts <- boxplot.stats(res$time)$stats
# filter out outliers (currently filtering fisher and map)
# res <- filter(res, time <= max(sts) * 1.05)

```

```{bash}
# Function to benchmark any bash command
repeats=10

bench_tests() {
    # --------------------------------------------------------------------------
    # Benchmark loop
    # --------------------------------------------------------------------------
    echo 'Benchmarking ' $1'...';

    # Run the given command [repeats] times
    for (( i = 1; i <= $repeats ; i++ ))
    do
        # Indicate the command we just ran in the csv file
        echo -ne $3"\t" >> $2
        # runs time function for the called script, output in a comma seperated
        # format output file specified with -o command and -a specifies append
        # requires gnu-time not BSD
        gtime -f %e -o ${2} -a ${1} > /dev/null 2>&1
    done;

}

# bedtools time
# set-up files for bedtools function
bedtools sort -i y.bed > y1.bed
bedtools sort -i x.bed > x1.bed
sort -k1,1 genome.txt > genome1.txt

## Run timing test
rm -f time.txt

bench_tests "bedtools random -g genome1.txt -n 1000000 -seed 1010486" "time.txt" "random"
bench_tests "bedtools shuffle -i x.bed -g genome1.txt -seed 1010486" "time.txt" "shuffle"
# single tbl function
bench_tests "bedtools slop -i x1.bed -g genome1.txt -b 1000" "time.txt" "slop"
bench_tests "bedtools flank -i x1.bed -g genome1.txt -b 1000" "time.txt" "flank"
bench_tests "bedtools merge -i x1.bed" "time.txt" "merge"
bench_tests "bedtools cluster -i x1.bed" "time.txt" "cluster"
bench_tests "bedtools complement -i x1.bed -g genome1.txt" "time.txt" "complement"
# multi tbl functions
bench_tests "bedtools closest -a x1.bed -b y1.bed" "time.txt" "closest"
bench_tests "bedtools intersect -a x1.bed -b y1.bed" "time.txt" "intersect"
bench_tests "bedtools map -a x1.bed -b y1.bed -c 1 -o count" "time.txt" "map"
bench_tests "bedtools subtract -a x1.bed -b y1.bed" "time.txt" "subtract"
# stats
bench_tests "bedtools reldist -a x1.bed -b y1.bed" "time.txt" "reldist"
bench_tests "bedtools jaccard -a x1.bed -b y1.bed" "time.txt" "jaccard"
bench_tests "bedtools fisher -a x1.bed -b y1.bed -g genome1.txt" "time.txt" "fisher"
#utils 
bench_tests "bedtools makewindows -b x1.bed -w 100" "time.txt" "makewindows"
```

```{r}
# Code to process data from previous chunks (above)

# This is to import the data from bedtools (bash) and graph it
times <- read_tsv("time.txt", col_names = c("expr", "time"))
dat <- list(times, res_hard)
names(dat) <- c("BEDTools", "valr")
dat <- bind_rows(dat, .id = "package")

mean_dat <- group_by(dat, expr, package) %>% 
             summarize(mean = mean(time), se = sd(time))

mean_dat$func <- ifelse(mean_dat$package == "bedtools", 
                        mean_dat$expr, 
                        str_split(mean_dat$expr, "\\(",  simplify = T)[, 1] %>% 
  str_split(., "_", simplify = T) %>% 
  .[,2])

ggplot(mean_dat, aes(x=reorder(func, mean), y = mean, fill = package)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se), 
                width = .1, position = position_dodge(1)) +
  scale_fill_brewer(palette = "Paired") + 
  coord_flip() +
  theme_bw() +
  labs(
    y='execution time (seconds)',
    x='',
    subtitle = paste(comma(n), "random 1 kilobase intervals,", 
                     comma(nrep), "repetitions")) +
  theme_classic() +
  pub_theme +
  theme(
    legend.text = element_text(size = 14),
    legend.title = element_text(size = 14)
  )

```

### 2017-06-06 inst/paper/main.tex
init commit

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Welcome to Overleaf --- just edit your article on the left,
% and we'll compile it for you on the right. If you give 
% someone the link to this page, they can edit at the same
% time. See the help menu above for more info. Enjoy!
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% For more detailed article preparation guidelines, please see:
% https://f1000research.com/for-authors/article-guidelines/software-tool-articles and http://f1000research.com/data-preparation

\documentclass[9pt,a4paper]{extarticle}
\usepackage{f1000_styles}

%% Default: numerical citations
\usepackage[numbers]{natbib}

%% Uncomment this lines for superscript citations instead
% \usepackage[super]{natbib}

%% Uncomment these lines for author-year citations instead
% \usepackage[round]{natbib}
% \let\cite\citep

%% syntax highlighting style kate 
\renewcommand{\KeywordTok}[1]{\textbf{{#1}}}
\renewcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.50,0.00,0.00}{{#1}}}
\renewcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,1.00}{{#1}}}
\renewcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,1.00}{{#1}}}
\renewcommand{\FloatTok}[1]{\textcolor[rgb]{0.50,0.00,0.50}{{#1}}}
\renewcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
\renewcommand{\CharTok}[1]{\textcolor[rgb]{1.00,0.00,1.00}{{#1}}}
\renewcommand{\SpecialCharTok}[1]{\textcolor[rgb]{1.00,0.00,1.00}{{#1}}}
\renewcommand{\StringTok}[1]{\textcolor[rgb]{0.87,0.00,0.00}{{#1}}}
\renewcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.87,0.00,0.00}{{#1}}}
\renewcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.87,0.00,0.00}{{#1}}}
\renewcommand{\ImportTok}[1]{{#1}}
\renewcommand{\CommentTok}[1]{\textcolor[rgb]{0.50,0.50,0.50}{\textit{{#1}}}}
\renewcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.50,0.50,0.50}{\textit{{#1}}}}
\renewcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.50,0.50,0.50}{\textbf{\textit{{#1}}}}}
\renewcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.50,0.50,0.50}{\textbf{\textit{{#1}}}}}
\renewcommand{\OtherTok}[1]{{#1}}
\renewcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.50}{{#1}}}
\renewcommand{\VariableTok}[1]{{#1}}
\renewcommand{\ControlFlowTok}[1]{{#1}}
\renewcommand{\OperatorTok}[1]{{#1}}
\renewcommand{\BuiltInTok}[1]{{#1}}
\renewcommand{\ExtensionTok}[1]{{#1}}
\renewcommand{\PreprocessorTok}[1]{\textbf{{#1}}}
\renewcommand{\AttributeTok}[1]{{#1}}
\renewcommand{\RegionMarkerTok}[1]{{#1}}
\renewcommand{\InformationTok}[1]{\textcolor[rgb]{0.50,0.50,0.50}{\textbf{\textit{{#1}}}}}
\renewcommand{\WarningTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{{#1}}}}
\renewcommand{\AlertTok}[1]{\textcolor[rgb]{0.00,1.00,0.00}{\textbf{{#1}}}}
\renewcommand{\ErrorTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{{#1}}}}
\renewcommand{\NormalTok}[1]{{#1}}

\begin{document}
\pagestyle{front}

\title{valr: Reproducible genome interval analysis in R}
\author[1]{Kent A. Riemondy}
\author[1]{Austin Gillen}
\author[2]{Ryan M. Sheridan}
\author[2]{Yinni Yu}
\author[3]{Christopher G. Bennett}
\author[1,2,*]{Jay R. Hesselberth}

\affil[1]{University of Colorado School of Medicine, RNA Bioscience Initiative, Aurora CO 80045}
\affil[2]{University of Colorado School of Medicine, Department of Biochemistry and Molecular Genetics}
\affil[3]{ComAnalyzeIT LLC, Fort Collins CO 80525}
\affil[*]{Corresponding author: jay.hesselberth@gmail.com}

\maketitle
\thispagestyle{front}

% 300 words
\begin{abstract}
New tools for genome internal analysis enable analysis of large genomic data sets
Availability of large genomic data sets. Need for new tools for exploratoratory data analysis. valr enables genomic interval analysis in the RStudio environment.
\end{abstract}

\section*{Keywords}

Genomics, Intervals, BEDtools, reproducibility, R, RStudio

\clearpage
\pagestyle{main}
\section*{Introduction}

Determining the correlation between sets of genomic features is a routine task in bioinformatics that requires sophisticated interval arithmetic. Command-line tools for interval analysis such as BEDtools \cite{quinlan_bedtools:_2010} and BEDOPS \cite{neph_bedops:_2012} are powerful tools that are central to many genomic analysis pipelines. A common workflow for genomic analysis begins with preprocessing interval analyses via command-line tools, followed by importing data into R for visualization and further statistical analyses. However, this workflow separation hinders exploratory data analysis, and the development reproducible research workflows built in the RMarkdown framework.

R packages developed for interval analysis include IRanges \cite{lawrence_software_2013}, GenometriCorr [@Favorov12], and bedr \cite{haider_bedr_2016}. IRanges is a Bioconductor package that provides specialized interval classes and methods to conduct interval arthimetic and is utilized extensively by Bioconductor packages. GenometriCorr provides a set of statistical tests to determine the relationships between interval sets and builds upon IRanges. bedr is a CRAN distributed package that provides wrapper functions to call BEDTools, BedOPS, and tabix functions providing out-of-memory support for interval analysis in R. These packages provide important functionality for interval analysis in R however they do not easily integrate with interactive data visualization features developed for R, such as Shiny\cite{chang_shiny_2017}, and the RStudio IDE. Additionally, the programmatic advances provided by the recent tidyverse suite of data processing and visualization tools (dplyr, purrr, broom, and ggplot2) do not easily interface with the R S4 objects implemented in many Bioconductor packages for data organization \cite{wickham_tidyverse_2017}.

We therefore sought to develop an R package that allows seamless and extensible interval arithmetic built to incorporate new R programming, visualization, and interactivity features.

\section*{Methods}
\subsection*{Implementation}
valr is implemented as an R package that makes extensive use of dplyr, a popular package that provides a flexible and high-performance framework for data manipulation \cite{wickham_dplyr_2016}. Computationally intensive functions in valr are written in C++ using Rcpp to enable fluid and quick interactive analysis of large datasets\cite{eddelbuettel_rcpp_2011}.

\subsection*{Operation}
valr is distributed as part of the CRAN R package repository and is compatible with Mac OS X, Windows, and major Linux operating systems. Package dependencies and system requirements are documented on the valr CRAN repository (\url{https://cran.r-project.org/web/packages/valr/index.html}).

\section*{Use Cases} % Optional - only if NO new datasets are included
To demonstrate the functionality and utility of \textbf{valr} presented below are some common use cases for genomic interval analysis.

\subsection{Basic Usage}

\subsubsection{Input data}\label{input-data}

\texttt{valr} provides a set of functions to read BED, BEDgraph, and VCF into R as convenient \texttt{tbl\_dfs()} dataframe objects. All tbls will have \texttt{chrom}, \texttt{start}, and\texttt{end} columns, and tbls from multi-column formats will have additional pre-determined column names.

Additionally \texttt{valr} supports connections to remote databases to programatically access the UCSC and Ensembl databases.

Show examples of read bed with files, url, and readucsc.

\subsection{Terse Syntax}

The functions in \texttt{valr} have similar names to their
\texttt{BEDtools} counterparts, and so will be familiar to users of
the \texttt{BEDtools} suite.

Similar to pybedtools\cite{dale_pybedtools:_2011}, a python wrapper for BEDTools, valr has a terse syntax. Shown below is an example of how to find all SNPs within 1 Kbp of genes using valr.

\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(valr)}
\KeywordTok{library}\NormalTok{(dplyr)}

\NormalTok{snps <-}\StringTok{ }\KeywordTok{read_bed}\NormalTok{(}\KeywordTok{valr_example}\NormalTok{(}\StringTok{'hg19.snps147.chr22.bed.gz'}\NormalTok{), }\DataTypeTok{n_fields =} \DecValTok{6}\NormalTok{)}
\NormalTok{genes <-}\StringTok{ }\KeywordTok{read_bed}\NormalTok{(}\KeywordTok{valr_example}\NormalTok{(}\StringTok{'genes.hg19.chr22.bed.gz'}\NormalTok{), }\DataTypeTok{n_fields =} \DecValTok{6}\NormalTok{)}

\CommentTok{# find snps in intergenic regions}
\NormalTok{intergenic <-}\StringTok{ }\KeywordTok{bed_subtract}\NormalTok{(snps, genes)}
\CommentTok{# distance from intergenic snps to nearest gene}
\NormalTok{nearby <-}\StringTok{ }\KeywordTok{bed_closest}\NormalTok{(intergenic, genes)}

\NormalTok{nearby %>%}
\StringTok{  }\KeywordTok{select}\NormalTok{(}\KeywordTok{starts_with}\NormalTok{(}\StringTok{'name'}\NormalTok{), .overlap, .dist) %>%}
\StringTok{  }\KeywordTok{filter}\NormalTok{(}\KeywordTok{abs}\NormalTok{(.dist) <}\StringTok{ }\DecValTok{1000}\NormalTok{)}
\CommentTok{#> # A tibble: 285 x 4}
\CommentTok{#>         name.x            name.y .overlap .dist}
\CommentTok{#>          <chr>             <chr>    <int> <int>}
\CommentTok{#>  1   rs2261631             P704P        0  -267}
\CommentTok{#>  2 rs570770556             POTEH        0  -912}
\CommentTok{#>  3 rs538163832             POTEH        0  -952}
\CommentTok{#>  4   rs9606135            TPTEP1        0  -421}
\CommentTok{#>  5  rs11912392 ANKRD62P1-PARP4P3        0   104}
\CommentTok{#>  6   rs8136454          BC038197        0   355}
\CommentTok{#>  7   rs5992556              XKR3        0  -455}
\CommentTok{#>  8 rs114101676              GAB4        0   473}
\CommentTok{#>  9  rs62236167             CECR7        0   261}
\CommentTok{#> 10   rs5747023             CECR1        0  -386}
\CommentTok{#> # ... with 275 more rows}
\end{Highlighting}

\subsubsection{Visual documentation}\label{visual-documentation}

The \texttt{bed\_glyph()} tool illustrates the results of operations in
\texttt{valr}, similar to those found in the \texttt{BEDtools}
documentation. Shown below is the code to produce glyphs that illustrate the result of intersecting \texttt{x}
and \texttt{y} intervals with \texttt{bed\_intersect()} , and the result of merging \texttt{x} intervals with \texttt{bed\_merge()} (Figure 1 \ref{fig:Figure 1}).

\begin{Highlighting}[]
\NormalTok{x <-}\StringTok{ }\NormalTok{tibble::}\KeywordTok{tribble}\NormalTok{(}
  \NormalTok{~chrom, ~start, ~end,}
  \StringTok{'chr1'}\NormalTok{, }\DecValTok{25}\NormalTok{,     }\DecValTok{50}\NormalTok{,}
  \StringTok{'chr1'}\NormalTok{, }\DecValTok{100}\NormalTok{,    }\DecValTok{125}
\NormalTok{)}

\NormalTok{y <-}\StringTok{ }\NormalTok{tibble::}\KeywordTok{tribble}\NormalTok{(}
  \NormalTok{~chrom, ~start, ~end,}
  \StringTok{'chr1'}\NormalTok{, }\DecValTok{30}\NormalTok{,     }\DecValTok{75}
\NormalTok{)}

\KeywordTok{bed_glyph}\NormalTok{(}\KeywordTok{bed_intersect}\NormalTok{(x, y))}
\end{Highlighting}

And this glyph illustrates \texttt{bed\_merge()}:

\begin{Highlighting}[]
\NormalTok{x <-}\StringTok{ }\NormalTok{tibble::}\KeywordTok{tribble}\NormalTok{(}
  \NormalTok{~chrom, ~start, ~end,}
  \StringTok{'chr1'}\NormalTok{,      }\DecValTok{1}\NormalTok{,      }\DecValTok{50}\NormalTok{,}
  \StringTok{'chr1'}\NormalTok{,      }\DecValTok{10}\NormalTok{,     }\DecValTok{75}\NormalTok{,}
  \StringTok{'chr1'}\NormalTok{,      }\DecValTok{100}\NormalTok{,    }\DecValTok{120}
\NormalTok{)}

\KeywordTok{bed_glyph}\NormalTok{(}\KeywordTok{bed_merge}\NormalTok{(x))}
\end{Highlighting}

\begin{figure}[!htb]
\centering
\includegraphics[width=\textwidth]{glyphs.pdf}
\caption{\label{fig:Figure 1}Visualizing interval operations in valr with \texttt{bed\_glyph()}.} 
\end{figure}

\subsubsection{Reproducible reports}\label{reproducible-reports}

\texttt{valr} can be used in RMarkdown documents to generate
reproducible work-flows for data processing. Because \texttt{valr} is
reasonably fast (see the \protect\hyperlink{benchmarks}{benchmarks}), we
now use it in lieu of other tools for exploratory analysis of genomic
data sets in R.

Command-line tools like \texttt{BEDtools} and \texttt{bedops} can be
used in reproducible workflows (e.g., with
\href{https://bitbucket.org/snakemake/snakemake/wiki/Home}{\texttt{snakemake}}),
but it is cumbersome to move from command-line tools to exploratory
analysis and plotting software.
\href{https://pythonhosted.org/pybedtools/}{\texttt{pybedtools}} can be
used within \texttt{ipython\ notebooks} to accomplish a similar goal,
but others have pointed out
\href{https://www.r-bloggers.com/why-i-dont-like-jupyter-fka-ipython-notebook/}{issues
with this approach}, including clunky version control. Because RMarkdown
files are text files, they are readily kept under version control.
Moreover, new features in RStudio, such as notebook viewing, and multiple language support enable similar functionality to \texttt{ipython}.

\subsubsection{Grouping data}\label{grouping-data}

The \texttt{group\_by} function in dplyr can be used to perform functions
on subsets of single and multiple \texttt{data\_frame}s. Functions in
\texttt{valr} leverage grouping to enable a variety of comparisons. For
example, intervals can be grouped by \texttt{strand} to perform
comparisons among intervals on the same strand.

\begin{Highlighting}[]
\NormalTok{x <-}\StringTok{ }\NormalTok{tibble::}\KeywordTok{tribble}\NormalTok{(}
  \NormalTok{~chrom, ~start, ~end, ~strand,}
  \StringTok{'chr1'}\NormalTok{, }\DecValTok{1}\NormalTok{,      }\DecValTok{100}\NormalTok{,  }\StringTok{'+'}\NormalTok{,}
  \StringTok{'chr1'}\NormalTok{, }\DecValTok{50}\NormalTok{,     }\DecValTok{150}\NormalTok{,  }\StringTok{'+'}\NormalTok{,}
  \StringTok{'chr2'}\NormalTok{, }\DecValTok{100}\NormalTok{,    }\DecValTok{200}\NormalTok{,  }\StringTok{'-'}
\NormalTok{)}

\NormalTok{y <-}\StringTok{ }\NormalTok{tibble::}\KeywordTok{tribble}\NormalTok{(}
  \NormalTok{~chrom, ~start, ~end, ~strand,}
  \StringTok{'chr1'}\NormalTok{, }\DecValTok{50}\NormalTok{,     }\DecValTok{125}\NormalTok{,  }\StringTok{'+'}\NormalTok{,}
  \StringTok{'chr1'}\NormalTok{, }\DecValTok{50}\NormalTok{,     }\DecValTok{150}\NormalTok{,  }\StringTok{'-'}\NormalTok{,}
  \StringTok{'chr2'}\NormalTok{, }\DecValTok{50}\NormalTok{,     }\DecValTok{150}\NormalTok{,  }\StringTok{'+'}
\NormalTok{)}

\CommentTok{# intersect tbls by strand}
\NormalTok{x <-}\StringTok{ }\KeywordTok{group_by}\NormalTok{(x, strand)}
\NormalTok{y <-}\StringTok{ }\KeywordTok{group_by}\NormalTok{(y, strand)}

\KeywordTok{bed_intersect}\NormalTok{(x, y)}
\CommentTok{#> # A tibble: 2 x 8}
\CommentTok{#>   chrom start.x end.x strand.x start.y end.y strand.y .overlap}
\CommentTok{#>   <chr>   <dbl> <dbl>    <chr>   <dbl> <dbl>    <chr>    <int>}
\CommentTok{#> 1  chr1       1   100        +      50   125        +       50}
\CommentTok{#> 2  chr1      50   150        +      50   125        +       75}
\end{Highlighting}

Comparisons between intervals on opposite strands are done using the
\texttt{flip\_strands()} function:

\begin{Highlighting}[]
\NormalTok{x <-}\StringTok{ }\KeywordTok{group_by}\NormalTok{(x, strand)}

\NormalTok{y <-}\StringTok{ }\KeywordTok{flip_strands}\NormalTok{(y)}
\NormalTok{y <-}\StringTok{ }\KeywordTok{group_by}\NormalTok{(y, strand)}

\KeywordTok{bed_intersect}\NormalTok{(x, y)}
\CommentTok{#> # A tibble: 3 x 8}
\CommentTok{#>   chrom start.x end.x strand.x start.y end.y strand.y .overlap}
\CommentTok{#>   <chr>   <dbl> <dbl>    <chr>   <dbl> <dbl>    <chr>    <int>}
\CommentTok{#> 1  chr2     100   200        -      50   150        -       50}
\CommentTok{#> 2  chr1       1   100        +      50   150        +       50}
\CommentTok{#> 3  chr1      50   150        +      50   150        +      100}
\end{Highlighting}

Both single set (e.g. \texttt{bed\_merge()}) and multi set operations
will respect groupings in the input intervals.

\subsubsection{Column specification}\label{column-specification}

Columns in \texttt{BEDtools} are referred to by position:

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# calculate the mean of column 6 for intervals in `b` that overlap with `a`}
\KeywordTok{bedtools} \NormalTok{map -a a.bed -b b.bed -c 6 -o mean}
\end{Highlighting}
\end{Shaded}

In \texttt{valr}, columns are referred to by name and can be used in
multiple name/value expressions for summaries.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# calculate the mean and variance for a `value` column}
\KeywordTok{bed_map}\NormalTok{(a, b, }\DataTypeTok{.mean =} \KeywordTok{mean}\NormalTok{(value), }\DataTypeTok{.var =} \KeywordTok{var}\NormalTok{(value))}

\CommentTok{# report concatenated and max values for merged intervals}
\KeywordTok{bed_merge}\NormalTok{(a, }\DataTypeTok{.concat =} \KeywordTok{concat}\NormalTok{(value), }\DataTypeTok{.max =} \KeywordTok{max}\NormalTok{(value))}
\end{Highlighting}
\end{Shaded}

\subsection{API}\label{api}
The major functions available in valr are shown in Table \ref{tab:tools}.

\begin{table}[h!]
\hrule \vspace{0.1cm}
\caption{\label{tab:tools} An overview of major functions available in valr.}
\centering
\begin{tabular}{lr}
 \bfseries \large Function Name & \bfseries  \large Purpose \\ 
  \hline
\bfseries Reading Data & \\ 
  read\_bed & Read BED files \\ 
  read\_bedgraph & Read bedGraph files \\
  read\_narrowpeak & Read narrowPeak files \\
  read\_broadpeak & Read broadPeak files \\
\bfseries Interval Transformation \\ 
  bed\_slop & Expand interval coordinates \\
  bed\_shift & Shift interval coordinates \\
  bed\_flank & Create flanking intervals \\
  bed\_merge & Merge overlapping intervals \\
  bed\_cluster & Identify (but not merge) overlapping intervals \\
  bed\_complement & Create intervals not covered by a query \\
\bfseries Interval Comparison & \\ 
  bed\_intersect & Report intersecting intervals from x and y tbls \\ 
  bed\_cluster & Cluster neighboring intervals \\
  bed\_map & Apply functions to selected columns for overlapping intervals \\
  bed\_subtract & Remove intervals based on overlaps \\
  bed\_window & Find overlapping intervals within a window \\
  bed\_closest & Find the closest intervals independent of overlaps \\
\bfseries Randomizing intervals & \\
  bed\_random & Generate random intervals from an input genome \\
  bed\_shuffle & Shuffle the coordinates of input intervals \\
\bfseries Interval statistics & \\
  bed\_fisher, bed\_projection & Calculate significance of overlaps between two sets of intervals \\
  bed\_reldist & Quantify relative distances between sets of intervals \\
  bed\_absdist & Quantify absolute distances between sets of intervals \\
  bed\_jaccard &Quantify extent of overlap between two sets of intervals \\
\bfseries Utilities &  \\ 
  bed\_glyph & Visualize the actions of valr functions \\ 
  bound\_intervals & Constrain intervals to a genome reference \\
  bed\_makewindows & Subdivide intervals \\
  bed12\_to\_exons & Convert BED12 to BED6 format \\
  interval\_spacing & Calculate spacing between intervals \\
  db\_ucsc, db\_ensembl & Access remote databases \\
  \end{tabular}
\end{table}

\subsection*{Summarizing interval coverage across genomic features}

This demonstration illustrates how to use \texttt{valr} tools to perform
a ``meta-analysis'' of signals relative to genomic features. Here we to
analyze the distribution of histone marks surrounding transcription
start sites.

First we load libraries and relevant data.

\begin{Highlighting}[]
\CommentTok{# `valr_example()` identifies the path of example files}
\NormalTok{bedfile <-}\StringTok{ }\KeywordTok{valr_example}\NormalTok{(}\StringTok{'genes.hg19.chr22.bed.gz'}\NormalTok{)}
\NormalTok{genomefile <-}\StringTok{ }\KeywordTok{valr_example}\NormalTok{(}\StringTok{'hg19.chrom.sizes.gz'}\NormalTok{)}
\NormalTok{bgfile  <-}\StringTok{ }\KeywordTok{valr_example}\NormalTok{(}\StringTok{'hela.h3k4.chip.bg.gz'}\NormalTok{)}

\NormalTok{genes <-}\StringTok{ }\KeywordTok{read_bed}\NormalTok{(bedfile, }\DataTypeTok{n_fields =} \DecValTok{6}\NormalTok{)}
\NormalTok{genome <-}\StringTok{ }\KeywordTok{read_genome}\NormalTok{(genomefile)}
\NormalTok{y <-}\StringTok{ }\KeywordTok{read_bedgraph}\NormalTok{(bgfile)}
\end{Highlighting}

Then we generate 1 bp intervals to represent transcription start sites
(TSSs). We focus on \texttt{+} strand genes, but \texttt{-} genes are
easily accomodated by filtering them and using
\texttt{bed\_makewindows()} with \texttt{reverse}d window numbers.

\begin{Highlighting}[]
\CommentTok{# generate 1 bp TSS intervals, `+` strand only}
\NormalTok{tss <-}\StringTok{ }\NormalTok{genes %>%}
\StringTok{  }\KeywordTok{filter}\NormalTok{(strand ==}\StringTok{ '+'}\NormalTok{) %>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{end =} \NormalTok{start +}\StringTok{ }\DecValTok{1}\NormalTok{)}

\CommentTok{# 1000 bp up and downstream}
\NormalTok{region_size <-}\StringTok{ }\DecValTok{1000}
\CommentTok{# 50 bp windows}
\NormalTok{win_size <-}\StringTok{ }\DecValTok{50}

\CommentTok{# add slop to the TSS, break into windows and add a group}
\NormalTok{x <-}\StringTok{ }\NormalTok{tss %>%}
\StringTok{  }\KeywordTok{bed_slop}\NormalTok{(genome, }\DataTypeTok{both =} \NormalTok{region_size) %>%}
\StringTok{  }\KeywordTok{bed_makewindows}\NormalTok{(genome, win_size)}

\NormalTok{x}
\CommentTok{#> # A tibble: 13,530 x 7}
\CommentTok{#>    chrom    start      end      name score strand .win_id}
\CommentTok{#>    <chr>    <int>    <int>     <chr> <chr>  <chr>   <int>}
\CommentTok{#>  1 chr22 16161065 16161115 LINC00516     3      +       1}
\CommentTok{#>  2 chr22 16161115 16161165 LINC00516     3      +       2}
\CommentTok{#>  3 chr22 16161165 16161215 LINC00516     3      +       3}
\CommentTok{#>  4 chr22 16161215 16161265 LINC00516     3      +       4}
\CommentTok{#>  5 chr22 16161265 16161315 LINC00516     3      +       5}
\CommentTok{#>  6 chr22 16161315 16161365 LINC00516     3      +       6}
\CommentTok{#>  7 chr22 16161365 16161415 LINC00516     3      +       7}
\CommentTok{#>  8 chr22 16161415 16161465 LINC00516     3      +       8}
\CommentTok{#>  9 chr22 16161465 16161515 LINC00516     3      +       9}
\CommentTok{#> 10 chr22 16161515 16161565 LINC00516     3      +      10}
\CommentTok{#> # ... with 13,520 more rows}
\end{Highlighting}

Now we use the \texttt{.win\_id} group with \texttt{bed\_map()} to
caluclate a sum by mapping \texttt{y} signals onto the intervals in
\texttt{x}. These data are regrouped by \texttt{.win\_id} and a summary
with \texttt{mean} and \texttt{sd} values is calculated.

\begin{Highlighting}[]
\CommentTok{# map signals to TSS regions and calculate summary statistics.}
\NormalTok{res <-}\StringTok{ }\KeywordTok{bed_map}\NormalTok{(x, y, }\DataTypeTok{win_sum =} \KeywordTok{sum}\NormalTok{(value, }\DataTypeTok{na.rm =} \OtherTok{TRUE}\NormalTok{)) %>%}
\StringTok{  }\KeywordTok{group_by}\NormalTok{(.win_id) %>%}
\StringTok{  }\KeywordTok{summarize}\NormalTok{(}\DataTypeTok{win_mean =} \KeywordTok{mean}\NormalTok{(win_sum, }\DataTypeTok{na.rm =} \OtherTok{TRUE}\NormalTok{),}
            \DataTypeTok{win_sd =} \KeywordTok{sd}\NormalTok{(win_sum, }\DataTypeTok{na.rm =} \OtherTok{TRUE}\NormalTok{))}

\NormalTok{res}
\CommentTok{#> # A tibble: 41 x 3}
\CommentTok{#>    .win_id win_mean    win_sd}
\CommentTok{#>      <int>    <dbl>     <dbl>}
\CommentTok{#>  1       1 100.8974  85.83423}
\CommentTok{#>  2       2 110.6829  81.13521}
\CommentTok{#>  3       3 122.9070  99.09635}
\CommentTok{#>  4       4 116.2800  96.30098}
\CommentTok{#>  5       5 116.3500 102.33773}
\CommentTok{#>  6       6 124.9048  95.08887}
\CommentTok{#>  7       7 122.9437  94.39792}
\CommentTok{#>  8       8 127.5946  91.47407}
\CommentTok{#>  9       9 130.2051  95.71309}
\CommentTok{#> 10      10 130.1220  88.82809}
\CommentTok{#> # ... with 31 more rows}
\end{Highlighting}

Finally, these summary statistics are used to construct a plot that
illustrates histone density surrounding TSSs (Figure 2\ref{fig:Figure 2}).

\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(ggplot2)}

\NormalTok{x_labels <-}\StringTok{ }\KeywordTok{seq}\NormalTok{(-region_size, region_size, }\DataTypeTok{by =} \NormalTok{win_size *}\StringTok{ }\DecValTok{5}\NormalTok{)}
\NormalTok{x_breaks <-}\StringTok{ }\KeywordTok{seq}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{41}\NormalTok{, }\DataTypeTok{by =} \DecValTok{5}\NormalTok{)}

\NormalTok{sd_limits <-}\StringTok{ }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{ymax =} \NormalTok{win_mean +}\StringTok{ }\NormalTok{win_sd, }\DataTypeTok{ymin =} \NormalTok{win_mean -}\StringTok{ }\NormalTok{win_sd)}

\KeywordTok{ggplot}\NormalTok{(res, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x =} \NormalTok{.win_id, }\DataTypeTok{y =} \NormalTok{win_mean)) +}
\StringTok{  }\KeywordTok{geom_point}\NormalTok{() +}\StringTok{ }\KeywordTok{geom_pointrange}\NormalTok{(sd_limits) +}\StringTok{ }
\StringTok{  }\KeywordTok{scale_x_continuous}\NormalTok{(}\DataTypeTok{labels =} \NormalTok{x_labels, }\DataTypeTok{breaks =} \NormalTok{x_breaks) +}\StringTok{ }
\StringTok{  }\KeywordTok{xlab}\NormalTok{(}\StringTok{'Position (bp from TSS)'}\NormalTok{) +}\StringTok{ }\KeywordTok{ylab}\NormalTok{(}\StringTok{'Signal'}\NormalTok{) +}\StringTok{ }
\StringTok{  }\KeywordTok{ggtitle}\NormalTok{(}\StringTok{'Human H3K4me3 signal near transcription start sites'}\NormalTok{) +}
\StringTok{  }\KeywordTok{theme_classic}\NormalTok{()}
\end{Highlighting}

\begin{figure}[!htb]
\centering
\includegraphics[width=\textwidth]{tss.pdf}
\caption{\label{fig:Figure 2}Meta-analysis of signals relative to genomic features with valr. \textnormal{Summarized coverage of human H3K4Me3 Chip-Seq coverage across positive strand transcription start sites. Data presented +/- SEM.}}
\end{figure}

\subsection*{Interval statistics}\label{interval-statistics}

Estimates of significance for interval overlaps can be obtained by
combining \texttt{bed\_shuffle()}, \texttt{bed\_random()} and the
\texttt{sample\_} functions from \texttt{dplyr} with interval statistics
in \texttt{valr}.

Here we examine the overlap of repeat classes in the human genome (on
\texttt{chr22} only, for simplicity) using \texttt{bed\_jaccard()}.

\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(purrr)}
\KeywordTok{library}\NormalTok{(tidyr)}

\NormalTok{repeats <-}\StringTok{ }\KeywordTok{read_bed}\NormalTok{(}\KeywordTok{valr_example}\NormalTok{(}\StringTok{'hg19.rmsk.chr22.bed.gz'}\NormalTok{), }\DataTypeTok{n_fields =} \DecValTok{6}\NormalTok{) }
\NormalTok{genome <-}\StringTok{ }\KeywordTok{read_genome}\NormalTok{(}\KeywordTok{valr_example}\NormalTok{(}\StringTok{'hg19.chrom.sizes.gz'}\NormalTok{))}

\NormalTok{shuffle_intervals <-}\StringTok{ }\NormalTok{function(n, .data, genome) \{}
  \KeywordTok{replicate}\NormalTok{(n, }\KeywordTok{bed_shuffle}\NormalTok{(.data, genome, }\DataTypeTok{seed =} \DecValTok{1010486}\NormalTok{), }\DataTypeTok{simplify =} \OtherTok{FALSE}\NormalTok{) %>%}
\StringTok{    }\KeywordTok{bind_rows}\NormalTok{(}\DataTypeTok{.id =} \StringTok{'rep'}\NormalTok{) %>%}
\StringTok{    }\KeywordTok{group_by}\NormalTok{(rep) %>%}\StringTok{ }\KeywordTok{nest}\NormalTok{()}
\NormalTok{\}}

\NormalTok{shuffled <-}\StringTok{ }\KeywordTok{shuffle_intervals}\NormalTok{(}\DataTypeTok{n =} \DecValTok{100}\NormalTok{, repeats, genome) %>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{jaccard =} \NormalTok{data %>%}
\StringTok{           }\KeywordTok{map}\NormalTok{(bed_jaccard, repeats) %>%}
\StringTok{           }\KeywordTok{map_dbl}\NormalTok{(}\StringTok{"jaccard"}\NormalTok{))}

\NormalTok{shuffled}
\CommentTok{#> # A tibble: 100 x 3}
\CommentTok{#>      rep                  data      jaccard}
\CommentTok{#>    <chr>                <list>        <dbl>}
\CommentTok{#>  1     1 <tibble [10,000 x 6]> 0.0004355887}
\CommentTok{#>  2     2 <tibble [10,000 x 6]> 0.0004355887}
\CommentTok{#>  3     3 <tibble [10,000 x 6]> 0.0004355887}
\CommentTok{#>  4     4 <tibble [10,000 x 6]> 0.0004355887}
\CommentTok{#>  5     5 <tibble [10,000 x 6]> 0.0004355887}
\CommentTok{#>  6     6 <tibble [10,000 x 6]> 0.0004355887}
\CommentTok{#>  7     7 <tibble [10,000 x 6]> 0.0004355887}
\CommentTok{#>  8     8 <tibble [10,000 x 6]> 0.0004355887}
\CommentTok{#>  9     9 <tibble [10,000 x 6]> 0.0004355887}
\CommentTok{#> 10    10 <tibble [10,000 x 6]> 0.0004355887}
\CommentTok{#> # ... with 90 more rows}
\end{Highlighting}

\subsection*{Analysis of DNA methylation}

\subsection*{Benchmarking against bedtools}
In order to ensure that valr performs fast enough to enable interactive analysis, key functionality is implemented in C++. We benchmarked major valr functions against corresponding commands in BEDTools. valr operates on tables already preloaded into RAM, whereas BEDTools performs file-reading, processing, and writing. To compare valr against bedtools we generated two files containing 1 million random 1 Kbp intervals derived from the human genome. For valr functions, we timed reading the table into R (e.g. with read\_bed()) and performing the respective function. For BEDTools commands we timed executing the command with the output written to /dev/null. valr functions performed similarly or slightly faster than BEDTools commands with the exception of the bed\_map and bed\_fisher (Figure 3\ref{fig:Figure 3}) .

\begin{figure}[!htb]
\centering
\includegraphics[width=\textwidth]{benchmarks.pdf}
\caption{\label{fig:Figure 3} Performance of major valr functionality benchmarked against BEDTools \textnormal{Timings calculated using microbenchmark and GNU time commands for valr and BEDTools respectively (A).}}
\end{figure}

\section*{Summary} % Optional - only if NO new datasets are included
valr provides a flexible framework for interval arithmetic in R/Rstudio. valr functions are written with a simple and terse syntax that promotes flexible interactive analysis. Additionally by providing an easy-to-use interface for interval arithmetic in R, valr is also a useful teaching tool to introduce the analyses necessary to investigate correlations between genomic intervals, without requiring familiarity with the command-line.

bedr provides a complementary set of tools for working with datasets where RAM limits loading all data into R. IRanges, is a mature package, but is a different philsophy. valr provides some of the functionality of Genometricorr, nbut is extensible to additional statistical test.

\section*{Software availability}
\begin{enumerate}
\item URL valr can be installed via CRAN using "install.packages('valr')". 
\item valr is maintained at http://github.com/rnabioco/valr
\item valr source code is available at http://github.com/rnabioco/valr
\item Latest stable version of source code is at: https://github.com/rnabioco/valr/archive/v0.2.0.tar.gz TO UPDATE
\item Software license: MIT 
\end{enumerate}

\section*{Author contributions}
In order to give appropriate credit to each author of an article, the individual
contributions of each author to the manuscript should be detailed in this section. We
recommend using author initials and then stating briefly how they contributed.

\section*{Competing interests}
No competing interests were disclosed

\section*{Grant information}
This work was supported by the RNA Bioscience Initiative (funded by a Transformational Research Award from the University of Colorado School of Medicine) and a grant from the National Institutes of Health (XXX to J.H.)

\section*{Acknowledgments}
This work was in part completed during an NIH sponsored Hackathon hosted by the Biofrontiers Department at the University of Colorado at Boulder.

{\small\bibliographystyle{unsrtnat}
\bibliography{valr}}

\section*{Figures}

\begin{figure}
\centering
    \begin{subfigure}[b]{0.45\textwidth}
\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{x <-}\StringTok{ }\KeywordTok{trbl_interval}\NormalTok{(}
  \NormalTok{~chrom, ~start, ~end,}
  \StringTok{'chr1'}\NormalTok{, }\DecValTok{25}\NormalTok{,     }\DecValTok{50}\NormalTok{,}
  \StringTok{'chr1'}\NormalTok{, }\DecValTok{100}\NormalTok{,    }\DecValTok{125}
\NormalTok{)}

\NormalTok{y <-}\StringTok{ }\KeywordTok{trbl_interval}\NormalTok{(}
  \NormalTok{~chrom, ~start, ~end,}
  \StringTok{'chr1'}\NormalTok{, }\DecValTok{30}\NormalTok{,     }\DecValTok{75}
\NormalTok{)}

\KeywordTok{bed_glyph}\NormalTok{(}\KeywordTok{bed_intersect}\NormalTok{(x, y))}
\end{Highlighting}
\end{Shaded}
        \label{fig:bed_glyph_code}
    \end{subfigure}
    ~
\begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{intersect_glyph-1.pdf}
        \label{fig:bed_glyph_out}
    \end{subfigure}
\caption{\label{fig:your-figure}Visualizing interval intersections with valr.	 }
\end{figure}

\begin{figure}
\centering
    \begin{subfigure}[b]{\textwidth}
\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# `valr_example()` identifies the path of example files}
\NormalTok{bedfile <-}\StringTok{ }\KeywordTok{valr_example}\NormalTok{(}\StringTok{'genes.hg19.chr22.bed.gz'}\NormalTok{)}
\NormalTok{genomefile <-}\StringTok{ }\KeywordTok{valr_example}\NormalTok{(}\StringTok{'hg19.chrom.sizes.gz'}\NormalTok{)}
\NormalTok{bgfile  <-}\StringTok{ }\KeywordTok{valr_example}\NormalTok{(}\StringTok{'hela.h3k4.chip.bg.gz'}\NormalTok{)}

\NormalTok{genes <-}\StringTok{ }\KeywordTok{read_bed}\NormalTok{(bedfile, }\DataTypeTok{n_fields =} \DecValTok{6}\NormalTok{)}
\NormalTok{genome <-}\StringTok{ }\KeywordTok{read_genome}\NormalTok{(genomefile)}
\NormalTok{y <-}\StringTok{ }\KeywordTok{read_bedgraph}\NormalTok{(bgfile)}

\CommentTok{# generate 1 bp TSS intervals, `+` strand only}
\NormalTok{tss <-}\StringTok{ }\NormalTok{genes %>%}\StringTok{ }\KeywordTok{filter}\NormalTok{(strand ==}\StringTok{ '+'}\NormalTok{) %>%}\StringTok{ }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{end =} \NormalTok{start +}\StringTok{ }\DecValTok{1}\NormalTok{)}

\CommentTok{# 1000 bp up and downstream}
\NormalTok{region_size <-}\StringTok{ }\DecValTok{1000}

\CommentTok{# 50 bp windows}
\NormalTok{win_size <-}\StringTok{ }\DecValTok{50}

\CommentTok{# add slop to the TSS, break into windows and add a group}
\NormalTok{x <-}\StringTok{ }\NormalTok{tss %>%}\StringTok{ }\KeywordTok{bed_slop}\NormalTok{(genome, }\DataTypeTok{both =} \NormalTok{region_size) %>%}
\StringTok{  }\KeywordTok{bed_makewindows}\NormalTok{(genome, win_size)}

\CommentTok{# map signals to TSS regions and calculate summary statistics.}
\NormalTok{res <-}\StringTok{ }\KeywordTok{bed_map}\NormalTok{(x, y, }\DataTypeTok{win_sum =} \KeywordTok{sum}\NormalTok{(value, }\DataTypeTok{na.rm =} \OtherTok{TRUE}\NormalTok{)) %>%}
\StringTok{  }\KeywordTok{group_by}\NormalTok{(.win_id) %>%}
\StringTok{  }\KeywordTok{summarize}\NormalTok{(}\DataTypeTok{win_mean =} \KeywordTok{mean}\NormalTok{(win_sum, }\DataTypeTok{na.rm =} \OtherTok{TRUE}\NormalTok{),}
            \DataTypeTok{win_sd =} \KeywordTok{sd}\NormalTok{(win_sum, }\DataTypeTok{na.rm =} \OtherTok{TRUE}\NormalTok{))}

\NormalTok{x_labels <-}\StringTok{ }\KeywordTok{seq}\NormalTok{(-region_size, region_size, }\DataTypeTok{by =} \NormalTok{win_size *}\StringTok{ }\DecValTok{5}\NormalTok{)}
\NormalTok{x_breaks <-}\StringTok{ }\KeywordTok{seq}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{41}\NormalTok{, }\DataTypeTok{by =} \DecValTok{5}\NormalTok{)}

\NormalTok{sd_limits <-}\StringTok{ }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{ymax =} \NormalTok{win_mean +}\StringTok{ }\NormalTok{win_sd, }\DataTypeTok{ymin =} \NormalTok{win_mean -}\StringTok{ }\NormalTok{win_sd)}

\KeywordTok{ggplot}\NormalTok{(res, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x =} \NormalTok{.win_id, }\DataTypeTok{y =} \NormalTok{win_mean)) +}\StringTok{ }\KeywordTok{geom_point}\NormalTok{() +}\StringTok{ }\KeywordTok{ylab}\NormalTok{(}\StringTok{'Signal'}\NormalTok{) +}
\StringTok{  }\KeywordTok{scale_x_continuous}\NormalTok{(}\DataTypeTok{labels =} \NormalTok{x_labels, }\DataTypeTok{breaks =} \NormalTok{x_breaks) +}\StringTok{ }
\StringTok{  }\KeywordTok{ggtitle}\NormalTok{(}\StringTok{'Human H3K4me3 signal near transcription start sites'}\NormalTok{) +}
\StringTok{  }\KeywordTok{xlab}\NormalTok{(}\StringTok{'Position (bp from TSS)'}\NormalTok{) +}\StringTok{ }\KeywordTok{geom_pointrange}\NormalTok{(sd_limits) +}
\StringTok{  }\KeywordTok{theme_classic}\NormalTok{()}
\end{Highlighting}
\end{Shaded}
        \label{fig:meta_gene_code}
    \end{subfigure}

\begin{subfigure}[b]{0.5\textwidth}
        \includegraphics[width=\textwidth]{demo-tss-1.pdf}
        \label{fig:meta_gene_out}
    \end{subfigure}
\caption{\label{fig:your-figure}Meta-analysis of signals relative to genomic features with valr.}
\end{figure}

% See this guide for more information on BibTeX:
% http://libguides.mit.edu/content.php?pid=55482&sid=406343

% For more author guidance please see:
% https://f1000research.com/for-authors/article-guidelines/software-tool-articles

% When all authors are happy with the paper, use the 
% ‘Submit to F1000RESEARCH' button from the menu above
% to submit directly to the open life science journal F1000Research.

% Please note that this template results in a draft pre-submission PDF document.
% Articles will be professionally typeset when accepted for publication.

% We hope you find the F1000Research Overleaf template useful,
% please let us know if yo\begin{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(dplyr)}
\KeywordTok{library}\NormalTok{(valr)}
\KeywordTok{library}\NormalTok{(R.utils)}
\KeywordTok{library}\NormalTok{(curl)}
\KeywordTok{library}\NormalTok{(ggplot2)}
\KeywordTok{library}\NormalTok{(gridExtra)}
\end{Highlighting}
\end{Shaded}

\subsection{Get the public datasets from
UCSC}\label{get-the-public-datasets-from-ucsc}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{DMSO_treated<-}\KeywordTok{read_bed}\NormalTok{(}\StringTok{'http://hgdownload.soe.ucsc.edu/goldenPath/hg19/encodeDCC/wgEncodeHaibMethylRrbs/wgEncodeHaibMethylRrbsA549Dm002p7dHaibSitesRep1.bed.gz'}\NormalTok{, }\DataTypeTok{n_fields=}\DecValTok{6}\NormalTok{, }\DataTypeTok{skip=}\DecValTok{1}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{untreated<-}\KeywordTok{read_bed}\NormalTok{(}\StringTok{'http://hgdownload.soe.ucsc.edu/goldenPath/hg19/encodeDCC/wgEncodeHaibMethylRrbs/wgEncodeHaibMethylRrbsAg04449UwSitesRep1.bed.gz'}\NormalTok{, }\DataTypeTok{n_fields=}\DecValTok{6}\NormalTok{, }\DataTypeTok{skip=}\DecValTok{1}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

To view the data

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{head}\NormalTok{(untreated)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
#> # A tibble: 6 x 6
#>   chrom start   end           name score strand
#>   <chr> <int> <int>          <chr> <chr>  <chr>
#> 1  chr1 10785 10786 AG04449_0__JS_     1      +
#> 2  chr1 10788 10789 AG04449_0__JS_     1      +
#> 3  chr1 10794 10795 AG04449_0__JS_     1      +
#> 4  chr1 10810 10811 AG04449_0__JS_     1      +
#> 5  chr1 10812 10813 AG04449_0__JS_     1      +
#> 6  chr1 10815 10816 AG04449_0__JS_     1      +
\end{verbatim}

\subsection{Examine regions that are different and regions that
intersect}\label{examine-regions-that-are-different-and-regions-that-intersect}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{different_region<-}\KeywordTok{bed_subtract}\NormalTok{(DMSO_treated,untreated, }\DataTypeTok{any=}\OtherTok{FALSE}\NormalTok{)}
\KeywordTok{dim}\NormalTok{(different_region)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
#> [1] 278241      6
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{intersect_region<-}\KeywordTok{bed_intersect}\NormalTok{(DMSO_treated,untreated, }\DataTypeTok{suffix=}\KeywordTok{c}\NormalTok{(}\StringTok{'.treated'}\NormalTok{,}\StringTok{'.untreated'}\NormalTok{))}
\KeywordTok{dim}\NormalTok{(intersect_region)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
#> [1] 1576400      12
\end{verbatim}

\subsection{Plot the distribution of different regions and intersected
regions across
chromosomes}\label{plot-the-distribution-of-different-regions-and-intersected-regions-across-chromosomes}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{p1<-}\KeywordTok{ggplot}\NormalTok{(}\DataTypeTok{data=}\NormalTok{different_region, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x=}\NormalTok{chrom))+}\KeywordTok{geom_histogram}\NormalTok{(}\DataTypeTok{stat=}\StringTok{'count'}\NormalTok{, }\KeywordTok{aes}\NormalTok{(}\KeywordTok{factor}\NormalTok{(chrom), }\DataTypeTok{fill=}\StringTok{'differential methylation'}\NormalTok{))+}\KeywordTok{scale_fill_brewer}\NormalTok{(}\DataTypeTok{palette =} \StringTok{'Set2'}\NormalTok{)+}\KeywordTok{theme_minimal}\NormalTok{()+}\KeywordTok{ggtitle}\NormalTok{(}\StringTok{'Distribution of differential methylation across chromosomes'}\NormalTok{)+}\KeywordTok{theme}\NormalTok{(}\DataTypeTok{axis.text.x =} \KeywordTok{element_text}\NormalTok{(}\DataTypeTok{angle =} \DecValTok{45}\NormalTok{, }\DataTypeTok{hjust =} \DecValTok{1}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
#> Warning: Ignoring unknown parameters: binwidth, bins, pad
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{p2<-}\KeywordTok{ggplot}\NormalTok{(}\DataTypeTok{data=}\NormalTok{intersect_region, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x=}\NormalTok{chrom))+}\KeywordTok{geom_histogram}\NormalTok{(}\DataTypeTok{stat=}\StringTok{'count'}\NormalTok{, }\KeywordTok{aes}\NormalTok{(}\KeywordTok{factor}\NormalTok{(chrom), }\DataTypeTok{fill=}\StringTok{'overlap methylation'}\NormalTok{))+}\KeywordTok{scale_fill_brewer}\NormalTok{(}\DataTypeTok{palette =} \StringTok{'Spectral'}\NormalTok{)+}\KeywordTok{theme_minimal}\NormalTok{()+}\KeywordTok{ggtitle}\NormalTok{(}\StringTok{'Distribution of overlapping methylation across chromosomes'}\NormalTok{)+}\KeywordTok{theme}\NormalTok{(}\DataTypeTok{axis.text.x =} \KeywordTok{element_text}\NormalTok{(}\DataTypeTok{angle =} \DecValTok{45}\NormalTok{, }\DataTypeTok{hjust =} \DecValTok{1}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
#> Warning: Ignoring unknown parameters: binwidth, bins, pad
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{grid.arrange}\NormalTok{(p1, p2, }\DataTypeTok{ncol =} \DecValTok{1}\NormalTok{, }\DataTypeTok{nrow =}\DecValTok{2}\NormalTok{)}
\end{Highlighting}
\end{Shaded}
\begin{figure}
\begin{center}\includegraphics{unnamed-chunk-8-1} \end{center}
\end{figure}

\end{document}

### 2017-06-06 inst/paper/figs.Rmd
update filenames, change fig3 color to greys

@@
 knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(cache = TRUE)
 ```
@@
 ```
+
 ## Figure 1
@@
 plot_grid(a, b, align = 'h', nrow = 1, labels="AUTO", scale = c(.9, .9))
-ggsave("glyphs.pdf", width = 8, height = 4)
+ggsave("figure1.pdf", width = 8, height = 4)
 
@@
 plot_grid(p, align = 'h', nrow = 1, labels="AUTO")
-ggsave("tss.pdf", width = 3.5, height = 2.25)
+ggsave("figure2.pdf", width = 3.5, height = 2.25)
 ```
@@
 bench_tests "bedtools makewindows -b x1.bed -w 100" "time.txt" "makewindows"
+
+# cleanup data
+rm -f x.bed y.bed x1.bed y1.bed genome1.txt genome.txt
 ```
@@
 
-mean_dat$func <- ifelse(mean_dat$package == "bedtools", 
+mean_dat$func <- ifelse(mean_dat$package == "BEDTools", 
                         mean_dat$expr, 
@@
 
-ggplot(mean_dat, aes(x=reorder(func, mean), y = mean, fill = package)) +
+p <- ggplot(mean_dat, aes(x=reorder(func, mean), y = mean, fill = package)) +
   geom_bar(stat = "identity", position = "dodge") +
@@
                 width = .1, position = position_dodge(1)) +
-  scale_fill_brewer(palette = "Paired") + 
+  scale_fill_grey() + 
   coord_flip() +
@@
     legend.text = element_text(size = 14),
-    legend.title = element_text(size = 14)
+    legend.title = element_blank(),
+    plot.subtitle = element_text(size = 10)
   )
 
+plot_grid(p, align = 'h', nrow = 1, labels="AUTO")
+ggsave("figure3.pdf")
+
 ```

### 2017-06-06 inst/paper/figs.Rmd
update fig color, size, names, and update manuscript

@@
 plot_grid(a, b, align = 'h', nrow = 1, labels="AUTO", scale = c(.9, .9))
-ggsave("figure1.pdf", width = 8, height = 4)
+ggsave("figure1.pdf", width = 11.2, height = 5.6)
 
@@
   xlab('Position (bp from TSS)') + ylab('Signal') + 
-  pub_theme +
-  theme_classic()
+  theme_classic() +
+  pub_theme
 
 plot_grid(p, align = 'h', nrow = 1, labels="AUTO")
-ggsave("figure2.pdf", width = 3.5, height = 2.25)
+ggsave("figure2.pdf", width = 5.6, height = 4)
 ```
@@
 plot_grid(p, align = 'h', nrow = 1, labels="AUTO")
-ggsave("figure3.pdf")
+ggsave("figure3.pdf", width = 5.6, height = 4)
 

### 2017-06-06 inst/paper/main.tex
update fig color, size, names, and update manuscript

@@
 
-R packages developed for interval analysis include IRanges \cite{lawrence_software_2013}, GenometriCorr [@Favorov12], and bedr \cite{haider_bedr_2016}.
+R packages developed for interval analysis include IRanges \cite{lawrence_software_2013}, GenometriCorr \cite{favorov_exploring_2012}, and bedr \cite{haider_bedr_2016}.
 IRanges is a Bioconductor package that provides specialized interval classes and methods to conduct interval arthimetic and is utilized extensively by Bioconductor packages.
@@
 
-\begin{figure}[!htb]
-\centering
-\includegraphics[width=\textwidth]{glyphs.pdf}
-\caption{\label{fig:Figure 1}Visualizing interval operations in valr with \texttt{bed\_glyph()}.} 
-\end{figure}
-
 \subsubsection{Reproducible reports}\label{reproducible-reports}
@@
 
-\begin{Shaded}
 \begin{Highlighting}[]
@@
 \end{Highlighting}
-\end{Shaded}
 
@@
 The major functions available in valr are shown in Table \ref{tab:tools}.
-
-\begin{table}[h!]
-\hrule \vspace{0.1cm}
-\caption{\label{tab:tools} An overview of major functions available in valr.}
-\centering
-\begin{tabular}{lr}
- \bfseries \large Function Name & \bfseries  \large Purpose \\ 
-  \hline
-\bfseries Reading Data & \\ 
-  read\_bed & Read BED files \\ 
-  read\_bedgraph & Read bedGraph files \\
-  read\_narrowpeak & Read narrowPeak files \\
-  read\_broadpeak & Read broadPeak files \\
-\bfseries Interval Transformation \\ 
-  bed\_slop & Expand interval coordinates \\
-  bed\_shift & Shift interval coordinates \\
-  bed\_flank & Create flanking intervals \\
-  bed\_merge & Merge overlapping intervals \\
-  bed\_cluster & Identify (but not merge) overlapping intervals \\
-  bed\_complement & Create intervals not covered by a query \\
-\bfseries Interval Comparison & \\ 
-  bed\_intersect & Report intersecting intervals from x and y tbls \\ 
-  bed\_cluster & Cluster neighboring intervals \\
-  bed\_map & Apply functions to selected columns for overlapping intervals \\
-  bed\_subtract & Remove intervals based on overlaps \\
-  bed\_window & Find overlapping intervals within a window \\
-  bed\_closest & Find the closest intervals independent of overlaps \\
-\bfseries Randomizing intervals & \\
-  bed\_random & Generate random intervals from an input genome \\
-  bed\_shuffle & Shuffle the coordinates of input intervals \\
-\bfseries Interval statistics & \\
-  bed\_fisher, bed\_projection & Calculate significance of overlaps between two sets of intervals \\
-  bed\_reldist & Quantify relative distances between sets of intervals \\
-  bed\_absdist & Quantify absolute distances between sets of intervals \\
-  bed\_jaccard &Quantify extent of overlap between two sets of intervals \\
-\bfseries Utilities &  \\ 
-  bed\_glyph & Visualize the actions of valr functions \\ 
-  bound\_intervals & Constrain intervals to a genome reference \\
-  bed\_makewindows & Subdivide intervals \\
-  bed12\_to\_exons & Convert BED12 to BED6 format \\
-  interval\_spacing & Calculate spacing between intervals \\
-  db\_ucsc, db\_ensembl & Access remote databases \\
-  \end{tabular}
-\end{table}
 
@@
 
-\begin{figure}[!htb]
-\centering
-\includegraphics[width=\textwidth]{tss.pdf}
-\caption{\label{fig:Figure 2}Meta-analysis of signals relative to genomic features with valr.
-\textnormal{Summarized coverage of human H3K4Me3 Chip-Seq coverage across positive strand transcription start sites.
-Data presented +/- SEM.}}
-\end{figure}
-
 \subsection*{Interval statistics}\label{interval-statistics}
@@
 
-\begin{figure}[!htb]
-\centering
-\includegraphics[width=\textwidth]{benchmarks.pdf}
-\caption{\label{fig:Figure 3} Performance of major valr functionality benchmarked against BEDTools \textnormal{Timings calculated using microbenchmark and GNU time commands for valr and BEDTools respectively (A).}}
-\end{figure}
-
 \section*{Summary} % Optional - only if NO new datasets are included
@@
 
+%%%% Figure 1 %%%%
+
+\begin{figure}[!htb]
+\centering
+\includegraphics[width=1\textwidth]{figure1.pdf}
+\caption{\label{fig:Figure 1}Visualizing interval operations in valr with \texttt{bed\_glyph()}.} 
+\end{figure}
+
+%%%% Figure 2 %%%%
+
+\begin{figure}[!htb]
+\centering
+\includegraphics[width=0.5\textwidth]{figure2.pdf}
+\caption{\label{fig:Figure 2}Meta-analysis of signals relative to genomic features with valr.
+\textnormal{Summarized coverage of human H3K4Me3 Chip-Seq coverage across positive strand transcription start sites.
+Data presented +/- SEM.}}
+\end{figure}
+
+%%%% Figure 3 %%%%
+\begin{figure}[!htb]
+\centering
+\includegraphics[width=0.5\textwidth]{figure3.pdf}
+\caption{\label{fig:Figure 3} Performance of major valr functionality benchmarked against BEDTools \textnormal{Timings calculated using microbenchmark and GNU time commands for valr and BEDTools respectively (A).}}
+\end{figure}
+
+%%%% Table 1 %%%%
+
+\begin{table}[h!]
+\hrule \vspace{0.1cm}
+\caption{\label{tab:tools} An overview of major functions available in valr.}
+\centering
+\begin{tabular}{lr}
+ \bfseries \large Function Name & \bfseries  \large Purpose \\ 
+  \hline
+\bfseries Reading Data & \\ 
+  read\_bed & Read BED files \\ 
+  read\_bedgraph & Read bedGraph files \\
+  read\_narrowpeak & Read narrowPeak files \\
+  read\_broadpeak & Read broadPeak files \\
+\bfseries Interval Transformation \\ 
+  bed\_slop & Expand interval coordinates \\
+  bed\_shift & Shift interval coordinates \\
+  bed\_flank & Create flanking intervals \\
+  bed\_merge & Merge overlapping intervals \\
+  bed\_cluster & Identify (but not merge) overlapping intervals \\
+  bed\_complement & Create intervals not covered by a query \\
+\bfseries Interval Comparison & \\ 
+  bed\_intersect & Report intersecting intervals from x and y tbls \\ 
+  bed\_cluster & Cluster neighboring intervals \\
+  bed\_map & Apply functions to selected columns for overlapping intervals \\
+  bed\_subtract & Remove intervals based on overlaps \\
+  bed\_window & Find overlapping intervals within a window \\
+  bed\_closest & Find the closest intervals independent of overlaps \\
+\bfseries Randomizing intervals & \\
+  bed\_random & Generate random intervals from an input genome \\
+  bed\_shuffle & Shuffle the coordinates of input intervals \\
+\bfseries Interval statistics & \\
+  bed\_fisher, bed\_projection & Calculate significance of overlaps between two sets of intervals \\
+  bed\_reldist & Quantify relative distances between sets of intervals \\
+  bed\_absdist & Quantify absolute distances between sets of intervals \\
+  bed\_jaccard &Quantify extent of overlap between two sets of intervals \\
+\bfseries Utilities &  \\ 
+  bed\_glyph & Visualize the actions of valr functions \\ 
+  bound\_intervals & Constrain intervals to a genome reference \\
+  bed\_makewindows & Subdivide intervals \\
+  bed12\_to\_exons & Convert BED12 to BED6 format \\
+  interval\_spacing & Calculate spacing between intervals \\
+  db\_ucsc, db\_ensembl & Access remote databases \\
+  \end{tabular}
+\end{table}
+
+%%%%%%%%%%%% Previous Figures %%%%%%%%%%%
+
+%%%% TSS H3K4me3 example %%%%
 \begin{figure}
@@
 
+%%%% DNA methylation example %%%%
+
+\subsection{Get the public datasets from
+UCSC}\label{get-the-public-datasets-from-ucsc}
+
 \begin{Shaded}
@@
 \end{Shaded}
-
-\subsection{Get the public datasets from
-UCSC}\label{get-the-public-datasets-from-ucsc}
 

### 2017-06-09 inst/paper/figs.Rmd
update figs and add tutorial

@@
 date: "June 6, 2017"
-output: html_document
+output: pdf_document
 ---
@@
 
-pub_theme <- theme(axis.text = element_text(size = 14),
-                   plot.title = element_text(size = 14),
-                   strip.text = element_text(size = 14))
+pub_theme <- theme(axis.text = element_text(size = 20),
+                   axis.title = element_text(size = 20),
+                   plot.title = element_text(size = 20),
+                   strip.text = element_text(size = 20))
 
@@
 
-```{bash}
+```{bash, eval = F}
 # Function to benchmark any bash command
@@
 
-mean_dat <- group_by(dat, expr, package) %>% 
-             summarize(mean = mean(time), se = sd(time))
+#mean_dat <- group_by(dat, expr, package) %>% 
+#             summarize(mean = mean(time), se = sd(time))
 
-mean_dat$func <- ifelse(mean_dat$package == "BEDTools", 
-                        mean_dat$expr, 
-                        str_split(mean_dat$expr, "\\(",  simplify = T)[, 1] %>% 
+dat$func <- ifelse(dat$package == "BEDTools", 
+                   dat$expr, 
+                   str_split(dat$expr, "\\(",  simplify = T)[, 1] %>% 
   str_split(., "_", simplify = T) %>% 
@@
 
-p <- ggplot(mean_dat, aes(x=reorder(func, mean), y = mean, fill = package)) +
-  geom_bar(stat = "identity", position = "dodge") +
-  geom_errorbar(aes(ymin = mean - se, ymax = mean + se), 
-                width = .1, position = position_dodge(1)) +
-  scale_fill_grey() + 
+# remove makewindows, really throws off proportion of graph
+dat <- filter(dat, func != "makewindows")
+valr_data <- filter(dat, package == "valr")
+
+plot_a <- ggplot(res, aes(x=reorder(expr, time), y = time)) +
+  geom_boxplot(width = 0.75, size = 0.33, outlier.shape = NA, fill = RColorBrewer::brewer.pal(3, "Set1")[2]) +
   coord_flip() +
@@
     x='',
-    subtitle = paste(comma(n), "random 1 kilobase intervals,", 
-                     comma(nrep), "repetitions")) +
+    subtitle=paste0(comma(n), " random x/y intervals,\n", comma(nrep), " repetitions")) +
   theme_classic() +
@@
     legend.title = element_blank(),
-    plot.subtitle = element_text(size = 10)
+    legend.position = "none",
+    plot.subtitle = element_text(size = 18),
+    axis.text.x = element_text(angle = 90)
   )
 
-plot_grid(p, align = 'h', nrow = 1, labels="AUTO")
-ggsave("figure3.pdf", width = 5.6, height = 4)
+plot_b <- ggplot(dat, aes(x=reorder(func, time), y = time, fill = package)) +
+  geom_boxplot(width = 0.75, size = 0.33, outlier.shape = NA) +
+  scale_fill_brewer(palette = "Set1") + 
+  coord_flip() +
+  theme_bw() +
+  labs(
+    y='execution time (seconds)',
+    x='') +
+  theme_classic() +
+  pub_theme +
+  theme(
+    legend.text = element_text(size = 18),
+    legend.title = element_blank(),
+    legend.position = "top",
+    axis.text.x = element_text(angle = 90)
+  )
+
+plot_grid(plot_a, plot_b, align = 'h', nrow = 1, labels="AUTO", label_size = 28)
+ggsave("figure3.pdf", width = 20, height = 10)
 
 ```
+
+## Code demonstrations
+
+```{r}
+library(valr)
+# function to retrieve path to example data
+bed_filepath <- valr_example("3fields.bed.gz") 
+read_bed(bed_filepath)
+
+#using URL
+read_bed("https://github.com/rnabioco/valr/raw/master/inst/extdata/3fields.bed.gz")
+```
+
+Using mysql
+```{r, warnings = F}
+library(dplyr, warn.conflicts = F)
+hg19 <- db_ucsc("hg19")
+tbl(hg19, "refGene")
+```
+
+### Interval statistics
+
+Estimates of significance for interval overlaps can be obtained by combining `bed_shuffle()`, `bed_random()` and the `sample_` functions from `dplyr` with interval statistics in `valr`.
+
+Here we examine the overlap of repeat classes in the human genome (on `chr22` only, for simplicity) using `bed_jaccard()`.
+
+```{r reldist_shuffle, message = FALSE, warning = FALSE}
+
+repeats <- read_bed(valr_example('hg19.rmsk.chr22.bed.gz'), n_fields = 6) 
+genome <- read_genome(valr_example('hg19.chrom.sizes.gz'))
+genes <- read_bed12(valr_example('hg19.refGene.chr22.bed.gz'))
+exons <- bed12_to_exons(genes)
+
+shuffle_intervals <- function(n, .data, genome) {
+  replicate(n, bed_shuffle(.data, genome), simplify = FALSE) %>%
+    bind_rows(.id = 'rep') %>%
+    group_by(rep) %>% nest()
+}
+
+shuffled <- shuffle_intervals(n = 100, repeats, genome) %>%
+  mutate(jaccard = data %>%
+           map(bed_jaccard, repeats) %>%
+           map_dbl("jaccard"))
+
+shuffled
+
+bed_jaccard(repeats, exons)
+```
+
+## Correlations among DNase I hypersensitive sites
+
+Here we use `bed_jaccard` for a large-scale comparison of related datasets.
+As shown in the [`BEDtools` tutorial](http://quinlanlab.org/tutorials/bedtools/bedtools.html), we can measure the similarity of DNaseI hypersensitivity sites for 20 fetal tissue samples.
+
+This data was taken from [Maurano *et al.* Systematic Localization of Common Disease-Associated Variation in Regulatory DNA. (2012) *Science*](www.sciencemag.org/content/337/6099/1190.short).
+
+```{r init, warning = FALSE, message = FALSE}
+library(valr)
+library(dplyr)
+library(tidyr)
+library(purrr)
+library(broom)
+library(stringr)
+library(ggplot2)
+library(ggrepel)
+library(ComplexHeatmap)
+```
+
+First read all 20 BED files containing DNase I hypersensitivity sites from 20 fetal tissues.
+
+```{r read_files, eval = F}
+dnase_files <- list.files('data/dnasei',pattern = 'merge.bed.gz', full.names = TRUE)
+data <- dnase_files %>% map(read_bed, n_fields = 4)
+```
+
+Then generate a 20x20 table containing a Jaccard statistic for each of the 400 pairwise comparisons.
+
+```{r jaccard-mat, eval = F}
+res <- data %>%
+  cross2(.,.) %>%
+  map(lift(bed_jaccard)) %>% 
+  map("jaccard") %>%
+  flatten_dbl() %>%
+  matrix(nrow = 20, ncol = 20)
+```
+
+We also need to generate labels for the table from the file names.
+
+```{r jaccard-names, eval = F}
+# names are tissue + sample_num
+col_names <- dnase_files %>%
+  str_split('/') %>% map(`[[`, 3) %>%
+  str_split('\\.') %>% map(`[[`, 1) %>%
+  str_split('-') %>% map(`[[`, 1) %>%
+  flatten_chr() %>%
+  str_replace('^f', '') %>%
+  str_c(str_c('-', seq(length(.))))
+
+colnames(res) <- col_names
+rownames(res) <- col_names
+```
+
+Now the Jaccard coefficients can be visualized in heatmap form.
+
+```{r jaccard-heatmap, fig.cap='Heatmap of Jaccard scores', fig.align='center', fig.width=8, eval = F}
+Heatmap(res, color_space = 'Blues')
+```
+
+Finally we can do some PCA analysis on the Jaccard coefficients to identify clusters of related cell types.
+
+```{r jaccard-PCA, fig.cap='PCA analysis of Jaccard scores', fig.align='center', fig.width=8, eval = F}
+pca <- broom::tidy(prcomp(res)) %>% as_data_frame()
+
+pca_comps <- filter(pca, PC <= 2) %>%
+  tidyr::spread(PC, value) %>%
+  setNames(c('label','PC1','PC2'))
+
+ggplot(pca_comps, aes(PC1, PC2)) +
+  geom_point(size = 3, color = 'red') +
+  geom_text_repel(aes(label = label))
+```

### 2017-06-14 inst/paper/figs.tex
first draft

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(valr)}
\KeywordTok{library}\NormalTok{(tidyverse)}
\KeywordTok{library}\NormalTok{(cowplot)}
\end{Highlighting}
\end{Shaded}

\subsection{Figure 1}\label{figure-1}

\begin{Shaded}
\begin{Highlighting}[]

\CommentTok{# theme to make pretty for publication}

\NormalTok{pub_theme <-}\StringTok{ }\KeywordTok{theme}\NormalTok{(}\DataTypeTok{axis.text =} \KeywordTok{element_text}\NormalTok{(}\DataTypeTok{size =} \DecValTok{20}\NormalTok{),}
                   \DataTypeTok{axis.title =} \KeywordTok{element_text}\NormalTok{(}\DataTypeTok{size =} \DecValTok{20}\NormalTok{),}
                   \DataTypeTok{plot.title =} \KeywordTok{element_text}\NormalTok{(}\DataTypeTok{size =} \DecValTok{20}\NormalTok{),}
                   \DataTypeTok{strip.text =} \KeywordTok{element_text}\NormalTok{(}\DataTypeTok{size =} \DecValTok{20}\NormalTok{))}

\NormalTok{x <-}\StringTok{ }\NormalTok{tibble::}\KeywordTok{tribble}\NormalTok{(}
  \NormalTok{~chrom, ~start, ~end,}
  \StringTok{'chr1'}\NormalTok{, }\DecValTok{25}\NormalTok{,     }\DecValTok{50}\NormalTok{,}
  \StringTok{'chr1'}\NormalTok{, }\DecValTok{100}\NormalTok{,    }\DecValTok{125}
\NormalTok{)}

\NormalTok{y <-}\StringTok{ }\NormalTok{tibble::}\KeywordTok{tribble}\NormalTok{(}
  \NormalTok{~chrom, ~start, ~end,}
  \StringTok{'chr1'}\NormalTok{, }\DecValTok{30}\NormalTok{,     }\DecValTok{75}
\NormalTok{)}

\NormalTok{a <-}\StringTok{ }\KeywordTok{bed_glyph}\NormalTok{(}\KeywordTok{bed_intersect}\NormalTok{(x, y)) +}\StringTok{ }\NormalTok{pub_theme}

\NormalTok{x <-}\StringTok{ }\NormalTok{tibble::}\KeywordTok{tribble}\NormalTok{(}
  \NormalTok{~chrom, ~start, ~end,}
  \StringTok{'chr1'}\NormalTok{,      }\DecValTok{1}\NormalTok{,      }\DecValTok{50}\NormalTok{,}
  \StringTok{'chr1'}\NormalTok{,      }\DecValTok{10}\NormalTok{,     }\DecValTok{75}\NormalTok{,}
  \StringTok{'chr1'}\NormalTok{,      }\DecValTok{100}\NormalTok{,    }\DecValTok{120}
\NormalTok{)}

\NormalTok{b <-}\StringTok{ }\KeywordTok{bed_glyph}\NormalTok{(}\KeywordTok{bed_merge}\NormalTok{(x)) +}\StringTok{ }\NormalTok{pub_theme}
\KeywordTok{plot_grid}\NormalTok{(a, b, }\DataTypeTok{align =} \StringTok{'h'}\NormalTok{, }\DataTypeTok{nrow =} \DecValTok{1}\NormalTok{, }\DataTypeTok{labels=}\StringTok{"AUTO"}\NormalTok{, }\DataTypeTok{scale =} \KeywordTok{c}\NormalTok{(.}\DecValTok{9}\NormalTok{, .}\DecValTok{9}\NormalTok{))}
\CommentTok{#> Warning in align_plots(plotlist = plots, align = align): Graphs cannot be}
\CommentTok{#> horizontally aligned. Placing graphs unaligned.}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{ggsave}\NormalTok{(}\StringTok{"figure1.pdf"}\NormalTok{, }\DataTypeTok{width =} \FloatTok{11.2}\NormalTok{, }\DataTypeTok{height =} \FloatTok{5.6}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\subsection{Figure 2}\label{figure-2}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# `valr_example()` identifies the path of example files}
\NormalTok{bedfile <-}\StringTok{ }\KeywordTok{valr_example}\NormalTok{(}\StringTok{'genes.hg19.chr22.bed.gz'}\NormalTok{)}
\NormalTok{genomefile <-}\StringTok{ }\KeywordTok{valr_example}\NormalTok{(}\StringTok{'hg19.chrom.sizes.gz'}\NormalTok{)}
\NormalTok{bgfile  <-}\StringTok{ }\KeywordTok{valr_example}\NormalTok{(}\StringTok{'hela.h3k4.chip.bg.gz'}\NormalTok{)}

\NormalTok{genes <-}\StringTok{ }\KeywordTok{read_bed}\NormalTok{(bedfile, }\DataTypeTok{n_fields =} \DecValTok{6}\NormalTok{)}
\NormalTok{genome <-}\StringTok{ }\KeywordTok{read_genome}\NormalTok{(genomefile)}
\NormalTok{y <-}\StringTok{ }\KeywordTok{read_bedgraph}\NormalTok{(bgfile)}

\CommentTok{# generate 1 bp TSS intervals, `+` strand only}
\NormalTok{tss <-}\StringTok{ }\NormalTok{genes %>%}
\StringTok{  }\KeywordTok{filter}\NormalTok{(strand ==}\StringTok{ '+'}\NormalTok{) %>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{end =} \NormalTok{start +}\StringTok{ }\DecValTok{1}\NormalTok{)}

\CommentTok{# 1000 bp up and downstream}
\NormalTok{region_size <-}\StringTok{ }\DecValTok{1000}
\CommentTok{# 50 bp windows}
\NormalTok{win_size <-}\StringTok{ }\DecValTok{50}

\CommentTok{# add slop to the TSS, break into windows and add a group}
\NormalTok{x <-}\StringTok{ }\NormalTok{tss %>%}
\StringTok{  }\KeywordTok{bed_slop}\NormalTok{(genome, }\DataTypeTok{both =} \NormalTok{region_size) %>%}
\StringTok{  }\KeywordTok{bed_makewindows}\NormalTok{(genome, win_size)}

\NormalTok{x}
\CommentTok{#> # A tibble: 13,530 x 7}
\CommentTok{#>    chrom    start      end      name score strand .win_id}
\CommentTok{#>    <chr>    <int>    <int>     <chr> <chr>  <chr>   <int>}
\CommentTok{#>  1 chr22 16161065 16161115 LINC00516     3      +       1}
\CommentTok{#>  2 chr22 16161115 16161165 LINC00516     3      +       2}
\CommentTok{#>  3 chr22 16161165 16161215 LINC00516     3      +       3}
\CommentTok{#>  4 chr22 16161215 16161265 LINC00516     3      +       4}
\CommentTok{#>  5 chr22 16161265 16161315 LINC00516     3      +       5}
\CommentTok{#>  6 chr22 16161315 16161365 LINC00516     3      +       6}
\CommentTok{#>  7 chr22 16161365 16161415 LINC00516     3      +       7}
\CommentTok{#>  8 chr22 16161415 16161465 LINC00516     3      +       8}
\CommentTok{#>  9 chr22 16161465 16161515 LINC00516     3      +       9}
\CommentTok{#> 10 chr22 16161515 16161565 LINC00516     3      +      10}
\CommentTok{#> # ... with 13,520 more rows}

\CommentTok{# map signals to TSS regions and calculate summary statistics.}
\NormalTok{res <-}\StringTok{ }\KeywordTok{bed_map}\NormalTok{(x, y, }\DataTypeTok{win_sum =} \KeywordTok{sum}\NormalTok{(value, }\DataTypeTok{na.rm =} \OtherTok{TRUE}\NormalTok{)) %>%}
\StringTok{  }\KeywordTok{group_by}\NormalTok{(.win_id) %>%}
\StringTok{  }\KeywordTok{summarize}\NormalTok{(}\DataTypeTok{win_mean =} \KeywordTok{mean}\NormalTok{(win_sum, }\DataTypeTok{na.rm =} \OtherTok{TRUE}\NormalTok{),}
            \DataTypeTok{win_sd =} \KeywordTok{sd}\NormalTok{(win_sum, }\DataTypeTok{na.rm =} \OtherTok{TRUE}\NormalTok{))}

\NormalTok{res}
\CommentTok{#> # A tibble: 41 x 3}
\CommentTok{#>    .win_id win_mean    win_sd}
\CommentTok{#>      <int>    <dbl>     <dbl>}
\CommentTok{#>  1       1 100.8974  85.83423}
\CommentTok{#>  2       2 110.6829  81.13521}
\CommentTok{#>  3       3 122.9070  99.09635}
\CommentTok{#>  4       4 116.2800  96.30098}
\CommentTok{#>  5       5 116.3500 102.33773}
\CommentTok{#>  6       6 124.9048  95.08887}
\CommentTok{#>  7       7 122.9437  94.39792}
\CommentTok{#>  8       8 127.5946  91.47407}
\CommentTok{#>  9       9 130.2051  95.71309}
\CommentTok{#> 10      10 130.1220  88.82809}
\CommentTok{#> # ... with 31 more rows}

\NormalTok{x_labels <-}\StringTok{ }\KeywordTok{seq}\NormalTok{(-region_size, region_size, }\DataTypeTok{by =} \NormalTok{win_size *}\StringTok{ }\DecValTok{5}\NormalTok{)}
\NormalTok{x_breaks <-}\StringTok{ }\KeywordTok{seq}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{41}\NormalTok{, }\DataTypeTok{by =} \DecValTok{5}\NormalTok{)}

\NormalTok{sd_limits <-}\StringTok{ }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{ymax =} \NormalTok{win_mean +}\StringTok{ }\NormalTok{win_sd, }\DataTypeTok{ymin =} \NormalTok{win_mean -}\StringTok{ }\NormalTok{win_sd)}

\NormalTok{p <-}\StringTok{ }\KeywordTok{ggplot}\NormalTok{(res, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x =} \NormalTok{.win_id, }\DataTypeTok{y =} \NormalTok{win_mean)) +}
\StringTok{  }\KeywordTok{geom_point}\NormalTok{(}\DataTypeTok{size =} \FloatTok{0.25}\NormalTok{) +}\StringTok{ }\KeywordTok{geom_pointrange}\NormalTok{(sd_limits, }\DataTypeTok{size =} \FloatTok{0.1}\NormalTok{) +}\StringTok{ }
\StringTok{  }\KeywordTok{scale_x_continuous}\NormalTok{(}\DataTypeTok{labels =} \NormalTok{x_labels, }\DataTypeTok{breaks =} \NormalTok{x_breaks) +}\StringTok{ }
\StringTok{  }\KeywordTok{xlab}\NormalTok{(}\StringTok{'Position (bp from TSS)'}\NormalTok{) +}\StringTok{ }\KeywordTok{ylab}\NormalTok{(}\StringTok{'Signal'}\NormalTok{) +}\StringTok{ }
\StringTok{  }\KeywordTok{theme_classic}\NormalTok{() +}
\StringTok{  }\NormalTok{pub_theme}

\KeywordTok{plot_grid}\NormalTok{(p, }\DataTypeTok{align =} \StringTok{'h'}\NormalTok{, }\DataTypeTok{nrow =} \DecValTok{1}\NormalTok{, }\DataTypeTok{labels=}\StringTok{"AUTO"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{ggsave}\NormalTok{(}\StringTok{"figure2.pdf"}\NormalTok{, }\DataTypeTok{width =} \FloatTok{5.6}\NormalTok{, }\DataTypeTok{height =} \DecValTok{4}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\subsection{Figure 3}\label{figure-3}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Load required packages}
\KeywordTok{library}\NormalTok{(valr)}
\KeywordTok{library}\NormalTok{(tidyverse)}
\KeywordTok{library}\NormalTok{(scales)}
\KeywordTok{library}\NormalTok{(microbenchmark)}
\KeywordTok{library}\NormalTok{(stringr)}
\KeywordTok{library}\NormalTok{(cowplot)}

\CommentTok{# Set-up test genome}
\NormalTok{genome <-}\StringTok{ }\KeywordTok{read_genome}\NormalTok{(}\KeywordTok{valr_example}\NormalTok{(}\StringTok{'hg19.chrom.sizes.gz'}\NormalTok{))}
\KeywordTok{write_tsv}\NormalTok{(genome, }\StringTok{"genome.txt"}\NormalTok{, }\DataTypeTok{col_names=}\OtherTok{FALSE}\NormalTok{)}

\CommentTok{# number of intervals}
\NormalTok{n <-}\StringTok{ }\FloatTok{1e6}
\CommentTok{# number of timing reps}
\NormalTok{nrep <-}\StringTok{ }\DecValTok{10}

\NormalTok{seed_x <-}\StringTok{ }\DecValTok{1010486}
\NormalTok{x <-}\StringTok{ }\KeywordTok{bed_random}\NormalTok{(genome, }\DataTypeTok{n =} \NormalTok{n, }\DataTypeTok{seed =} \NormalTok{seed_x)}
\KeywordTok{write_tsv}\NormalTok{(x, }\StringTok{"x.bed"}\NormalTok{, }\DataTypeTok{col_names=}\OtherTok{FALSE}\NormalTok{)}
\NormalTok{seed_y <-}\StringTok{ }\DecValTok{9283019}
\NormalTok{y <-}\StringTok{ }\KeywordTok{bed_random}\NormalTok{(genome, }\DataTypeTok{n =} \NormalTok{n, }\DataTypeTok{seed =} \NormalTok{seed_y)}
\KeywordTok{write_tsv}\NormalTok{(y, }\StringTok{"y.bed"}\NormalTok{, }\DataTypeTok{col_names=}\OtherTok{FALSE}\NormalTok{)}

\NormalTok{res <-}\StringTok{ }\KeywordTok{microbenchmark}\NormalTok{(}
  \CommentTok{# randomizing functions}
  \KeywordTok{bed_random}\NormalTok{(genome, }\DataTypeTok{n =} \NormalTok{n, }\DataTypeTok{seed =} \NormalTok{seed_x),}
  \KeywordTok{bed_shuffle}\NormalTok{(x, genome, }\DataTypeTok{seed =} \NormalTok{seed_x),}
  \CommentTok{# # single tbl functions}
  \KeywordTok{bed_slop}\NormalTok{(x, genome, }\DataTypeTok{both =} \DecValTok{1000}\NormalTok{),}
  \KeywordTok{bed_flank}\NormalTok{(x, genome, }\DataTypeTok{both =} \DecValTok{1000}\NormalTok{),}
  \KeywordTok{bed_merge}\NormalTok{(x),}
  \KeywordTok{bed_cluster}\NormalTok{(x),}
  \KeywordTok{bed_complement}\NormalTok{(x, genome),}
  \CommentTok{# multi tbl functions}
  \KeywordTok{bed_closest}\NormalTok{(x, y),}
  \KeywordTok{bed_intersect}\NormalTok{(x, y),}
  \KeywordTok{bed_map}\NormalTok{(x, y, }\DataTypeTok{.n =} \KeywordTok{n}\NormalTok{()),}
  \KeywordTok{bed_subtract}\NormalTok{(x, y),}
  \CommentTok{# stats}
  \KeywordTok{bed_reldist}\NormalTok{(x, y),}
  \KeywordTok{bed_jaccard}\NormalTok{(x, y),}
  \KeywordTok{bed_fisher}\NormalTok{(x, y, genome),}
  \CommentTok{#utils }
  \KeywordTok{bed_makewindows}\NormalTok{(x, genome, }\DataTypeTok{win_size =} \DecValTok{100}\NormalTok{),}
  \DataTypeTok{times =} \NormalTok{nrep,}
  \DataTypeTok{unit =} \StringTok{'s'}\NormalTok{)}

\NormalTok{load_file <-}\StringTok{ }\NormalTok{function(filename) \{}
  \NormalTok{if(}\KeywordTok{endsWith}\NormalTok{(filename, }\StringTok{".txt"}\NormalTok{))\{}
    \KeywordTok{read_genome}\NormalTok{(filename)}
  \NormalTok{\} else if (}\KeywordTok{endsWith}\NormalTok{(filename, }\StringTok{".bed"}\NormalTok{)) \{}
    \KeywordTok{read_bed}\NormalTok{(filename)}
  \NormalTok{\}}
\NormalTok{\}}

\NormalTok{genome_file <-}\StringTok{ "genome.txt"}
\NormalTok{x_bed <-}\StringTok{ "x.bed"}
\NormalTok{y_bed  <-}\StringTok{ "y.bed"}

\NormalTok{res_hard <-}\StringTok{ }\KeywordTok{microbenchmark}\NormalTok{(}
  \CommentTok{# randomizing functions}
  \KeywordTok{bed_random}\NormalTok{(}\KeywordTok{load_file}\NormalTok{(genome_file), }\DataTypeTok{n =} \NormalTok{n, }\DataTypeTok{seed =} \NormalTok{seed_x),}
  \KeywordTok{bed_shuffle}\NormalTok{(}\KeywordTok{load_file}\NormalTok{(x_bed), }\KeywordTok{load_file}\NormalTok{(genome_file), }\DataTypeTok{seed =} \NormalTok{seed_x),}
  \CommentTok{# # single tbl functions}
  \KeywordTok{bed_slop}\NormalTok{(}\KeywordTok{load_file}\NormalTok{(x_bed), }\KeywordTok{load_file}\NormalTok{(genome_file), }\DataTypeTok{both =} \DecValTok{1000}\NormalTok{),}
  \KeywordTok{bed_flank}\NormalTok{(}\KeywordTok{load_file}\NormalTok{(x_bed), }\KeywordTok{load_file}\NormalTok{(genome_file), }\DataTypeTok{both =} \DecValTok{1000}\NormalTok{),}
  \KeywordTok{bed_merge}\NormalTok{(}\KeywordTok{load_file}\NormalTok{(x_bed)),}
  \KeywordTok{bed_cluster}\NormalTok{(}\KeywordTok{load_file}\NormalTok{(x_bed)),}
  \KeywordTok{bed_complement}\NormalTok{(}\KeywordTok{load_file}\NormalTok{(x_bed), }\KeywordTok{load_file}\NormalTok{(genome_file)),}
  \CommentTok{# multi tbl functions}
  \KeywordTok{bed_closest}\NormalTok{(}\KeywordTok{load_file}\NormalTok{(x_bed), }\KeywordTok{load_file}\NormalTok{(y_bed)),}
  \KeywordTok{bed_intersect}\NormalTok{(}\KeywordTok{load_file}\NormalTok{(x_bed), }\KeywordTok{load_file}\NormalTok{(y_bed)),}
  \KeywordTok{bed_map}\NormalTok{(}\KeywordTok{load_file}\NormalTok{(x_bed), }\KeywordTok{load_file}\NormalTok{(y_bed), }\DataTypeTok{.n =} \KeywordTok{n}\NormalTok{()),}
  \KeywordTok{bed_subtract}\NormalTok{(}\KeywordTok{load_file}\NormalTok{(x_bed), }\KeywordTok{load_file}\NormalTok{(y_bed)),}
  \CommentTok{# stats}
  \KeywordTok{bed_reldist}\NormalTok{(}\KeywordTok{load_file}\NormalTok{(x_bed), }\KeywordTok{load_file}\NormalTok{(y_bed)),}
  \KeywordTok{bed_jaccard}\NormalTok{(}\KeywordTok{load_file}\NormalTok{(x_bed), }\KeywordTok{load_file}\NormalTok{(y_bed)),}
  \KeywordTok{bed_fisher}\NormalTok{(}\KeywordTok{load_file}\NormalTok{(x_bed), }\KeywordTok{load_file}\NormalTok{(y_bed), }\KeywordTok{load_file}\NormalTok{(genome_file)),}
  \CommentTok{#utils }
  \KeywordTok{bed_makewindows}\NormalTok{(}\KeywordTok{load_file}\NormalTok{(x_bed), }\KeywordTok{load_file}\NormalTok{(genome_file), }\DataTypeTok{win_size =} \DecValTok{100}\NormalTok{),}
  \DataTypeTok{times =} \NormalTok{nrep,}
  \DataTypeTok{unit =} \StringTok{'s'}\NormalTok{)}

\CommentTok{# covert nanoseconds to seconds}
\NormalTok{res <-}\StringTok{ }\NormalTok{res %>%}
\StringTok{  }\KeywordTok{as_tibble}\NormalTok{() %>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{time =} \NormalTok{time /}\StringTok{ }\FloatTok{1e9}\NormalTok{)}

\NormalTok{res_hard <-}\StringTok{ }\NormalTok{res_hard %>%}
\StringTok{  }\KeywordTok{as_tibble}\NormalTok{() %>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{time =} \NormalTok{time /}\StringTok{ }\FloatTok{1e9}\NormalTok{)}

\CommentTok{# futz with the x-axis}
\NormalTok{sts <-}\StringTok{ }\KeywordTok{boxplot.stats}\NormalTok{(res$time)$stats}
\CommentTok{# filter out outliers (currently filtering fisher and map)}
\CommentTok{# res <- filter(res, time <= max(sts) * 1.05)}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Function to benchmark any bash command}
\OtherTok{repeats=}\NormalTok{10}

\FunctionTok{bench_tests()} \KeywordTok{\{}
    \CommentTok{# --------------------------------------------------------------------------}
    \CommentTok{# Benchmark loop}
    \CommentTok{# --------------------------------------------------------------------------}
    \KeywordTok{echo} \StringTok{'Benchmarking '} \OtherTok{$1}\StringTok{'...'}\KeywordTok{;}

    \CommentTok{# Run the given command [repeats] times}
    \KeywordTok{for} \KeywordTok{((} \NormalTok{i = 1; i <= }\OtherTok{$repeats} \NormalTok{; i++ }\KeywordTok{))}
    \KeywordTok{do}
        \CommentTok{# Indicate the command we just ran in the csv file}
        \KeywordTok{echo} \NormalTok{-ne }\OtherTok{$3}\StringTok{"\textbackslash{}t"} \KeywordTok{>>} \OtherTok{$2}
        \CommentTok{# runs time function for the called script, output in a comma seperated}
        \CommentTok{# format output file specified with -o command and -a specifies append}
        \CommentTok{# requires gnu-time not BSD}
        \KeywordTok{gtime} \NormalTok{-f %e -o }\OtherTok{$\{2\}} \NormalTok{-a }\OtherTok{$\{1\}} \KeywordTok{>} \NormalTok{/dev/null }\KeywordTok{2>&1}
    \KeywordTok{done}\NormalTok{;}

\KeywordTok{\}}

\CommentTok{# bedtools time}
\CommentTok{# set-up files for bedtools function}
\KeywordTok{bedtools} \NormalTok{sort -i y.bed }\KeywordTok{>} \NormalTok{y1.bed}
\KeywordTok{bedtools} \NormalTok{sort -i x.bed }\KeywordTok{>} \NormalTok{x1.bed}
\KeywordTok{sort} \NormalTok{-k1,1 genome.txt }\KeywordTok{>} \NormalTok{genome1.txt}

\CommentTok{## Run timing test}
\KeywordTok{rm} \NormalTok{-f time.txt}

\KeywordTok{bench_tests} \StringTok{"bedtools random -g genome1.txt -n 1000000 -seed 1010486"} \StringTok{"time.txt"} \StringTok{"random"}
\KeywordTok{bench_tests} \StringTok{"bedtools shuffle -i x.bed -g genome1.txt -seed 1010486"} \StringTok{"time.txt"} \StringTok{"shuffle"}
\CommentTok{# single tbl function}
\KeywordTok{bench_tests} \StringTok{"bedtools slop -i x1.bed -g genome1.txt -b 1000"} \StringTok{"time.txt"} \StringTok{"slop"}
\KeywordTok{bench_tests} \StringTok{"bedtools flank -i x1.bed -g genome1.txt -b 1000"} \StringTok{"time.txt"} \StringTok{"flank"}
\KeywordTok{bench_tests} \StringTok{"bedtools merge -i x1.bed"} \StringTok{"time.txt"} \StringTok{"merge"}
\KeywordTok{bench_tests} \StringTok{"bedtools cluster -i x1.bed"} \StringTok{"time.txt"} \StringTok{"cluster"}
\KeywordTok{bench_tests} \StringTok{"bedtools complement -i x1.bed -g genome1.txt"} \StringTok{"time.txt"} \StringTok{"complement"}
\CommentTok{# multi tbl functions}
\KeywordTok{bench_tests} \StringTok{"bedtools closest -a x1.bed -b y1.bed"} \StringTok{"time.txt"} \StringTok{"closest"}
\KeywordTok{bench_tests} \StringTok{"bedtools intersect -a x1.bed -b y1.bed"} \StringTok{"time.txt"} \StringTok{"intersect"}
\KeywordTok{bench_tests} \StringTok{"bedtools map -a x1.bed -b y1.bed -c 1 -o count"} \StringTok{"time.txt"} \StringTok{"map"}
\KeywordTok{bench_tests} \StringTok{"bedtools subtract -a x1.bed -b y1.bed"} \StringTok{"time.txt"} \StringTok{"subtract"}
\CommentTok{# stats}
\KeywordTok{bench_tests} \StringTok{"bedtools reldist -a x1.bed -b y1.bed"} \StringTok{"time.txt"} \StringTok{"reldist"}
\KeywordTok{bench_tests} \StringTok{"bedtools jaccard -a x1.bed -b y1.bed"} \StringTok{"time.txt"} \StringTok{"jaccard"}
\KeywordTok{bench_tests} \StringTok{"bedtools fisher -a x1.bed -b y1.bed -g genome1.txt"} \StringTok{"time.txt"} \StringTok{"fisher"}
\CommentTok{#utils }
\KeywordTok{bench_tests} \StringTok{"bedtools makewindows -b x1.bed -w 100"} \StringTok{"time.txt"} \StringTok{"makewindows"}

\CommentTok{# cleanup data}
\KeywordTok{rm} \NormalTok{-f x.bed y.bed x1.bed y1.bed genome1.txt genome.txt}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Code to process data from previous chunks (above)}

\CommentTok{# This is to import the data from bedtools (bash) and graph it}
\NormalTok{times <-}\StringTok{ }\KeywordTok{read_tsv}\NormalTok{(}\StringTok{"time.txt"}\NormalTok{, }\DataTypeTok{col_names =} \KeywordTok{c}\NormalTok{(}\StringTok{"expr"}\NormalTok{, }\StringTok{"time"}\NormalTok{))}
\CommentTok{#> Parsed with column specification:}
\CommentTok{#> cols(}
\CommentTok{#>   expr = col_character(),}
\CommentTok{#>   time = col_double()}
\CommentTok{#> )}
\NormalTok{dat <-}\StringTok{ }\KeywordTok{list}\NormalTok{(times, res_hard)}
\KeywordTok{names}\NormalTok{(dat) <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"BEDTools"}\NormalTok{, }\StringTok{"valr"}\NormalTok{)}
\NormalTok{dat <-}\StringTok{ }\KeywordTok{bind_rows}\NormalTok{(dat, }\DataTypeTok{.id =} \StringTok{"package"}\NormalTok{)}
\CommentTok{#> Warning in bind_rows_(x, .id): binding character and factor vector,}
\CommentTok{#> coercing into character vector}

\CommentTok{#mean_dat <- group_by(dat, expr, package) %>% }
\CommentTok{#             summarize(mean = mean(time), se = sd(time))}

\NormalTok{dat$func <-}\StringTok{ }\KeywordTok{ifelse}\NormalTok{(dat$package ==}\StringTok{ "BEDTools"}\NormalTok{, }
                   \NormalTok{dat$expr, }
                   \KeywordTok{str_split}\NormalTok{(dat$expr, }\StringTok{"}\CharTok{\textbackslash{}\textbackslash{}}\StringTok{("}\NormalTok{,  }\DataTypeTok{simplify =} \NormalTok{T)[, }\DecValTok{1}\NormalTok{] %>%}\StringTok{ }
\StringTok{  }\KeywordTok{str_split}\NormalTok{(., }\StringTok{"_"}\NormalTok{, }\DataTypeTok{simplify =} \NormalTok{T) %>%}\StringTok{ }
\StringTok{  }\NormalTok{.[,}\DecValTok{2}\NormalTok{])}

\CommentTok{# remove makewindows, really throws off proportion of graph}
\NormalTok{dat <-}\StringTok{ }\KeywordTok{filter}\NormalTok{(dat, func !=}\StringTok{ "makewindows"}\NormalTok{)}
\NormalTok{valr_data <-}\StringTok{ }\KeywordTok{filter}\NormalTok{(dat, package ==}\StringTok{ "valr"}\NormalTok{)}

\NormalTok{plot_a <-}\StringTok{ }\KeywordTok{ggplot}\NormalTok{(res, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x=}\KeywordTok{reorder}\NormalTok{(expr, time), }\DataTypeTok{y =} \NormalTok{time)) +}
\StringTok{  }\KeywordTok{geom_boxplot}\NormalTok{(}\DataTypeTok{width =} \FloatTok{0.75}\NormalTok{, }\DataTypeTok{size =} \FloatTok{0.33}\NormalTok{, }\DataTypeTok{outlier.shape =} \OtherTok{NA}\NormalTok{, }\DataTypeTok{fill =} \NormalTok{RColorBrewer::}\KeywordTok{brewer.pal}\NormalTok{(}\DecValTok{3}\NormalTok{, }\StringTok{"Set1"}\NormalTok{)[}\DecValTok{2}\NormalTok{]) +}
\StringTok{  }\KeywordTok{coord_flip}\NormalTok{() +}
\StringTok{  }\KeywordTok{theme_bw}\NormalTok{() +}
\StringTok{  }\KeywordTok{labs}\NormalTok{(}
    \DataTypeTok{y=}\StringTok{'execution time (seconds)'}\NormalTok{,}
    \DataTypeTok{x=}\StringTok{''}\NormalTok{,}
    \DataTypeTok{subtitle=}\KeywordTok{paste0}\NormalTok{(}\KeywordTok{comma}\NormalTok{(n), }\StringTok{" random x/y intervals,}\CharTok{\textbackslash{}n}\StringTok{"}\NormalTok{, }\KeywordTok{comma}\NormalTok{(nrep), }\StringTok{" repetitions"}\NormalTok{)) +}
\StringTok{  }\KeywordTok{theme_classic}\NormalTok{() +}
\StringTok{  }\NormalTok{pub_theme +}
\StringTok{  }\KeywordTok{theme}\NormalTok{(}
    \DataTypeTok{legend.text =} \KeywordTok{element_text}\NormalTok{(}\DataTypeTok{size =} \DecValTok{14}\NormalTok{),}
    \DataTypeTok{legend.title =} \KeywordTok{element_blank}\NormalTok{(),}
    \DataTypeTok{legend.position =} \StringTok{"none"}\NormalTok{,}
    \DataTypeTok{plot.subtitle =} \KeywordTok{element_text}\NormalTok{(}\DataTypeTok{size =} \DecValTok{18}\NormalTok{),}
    \DataTypeTok{axis.text.x =} \KeywordTok{element_text}\NormalTok{(}\DataTypeTok{angle =} \DecValTok{90}\NormalTok{)}
  \NormalTok{)}

\NormalTok{plot_b <-}\StringTok{ }\KeywordTok{ggplot}\NormalTok{(dat, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x=}\KeywordTok{reorder}\NormalTok{(func, time), }\DataTypeTok{y =} \NormalTok{time, }\DataTypeTok{fill =} \NormalTok{package)) +}
\StringTok{  }\KeywordTok{geom_boxplot}\NormalTok{(}\DataTypeTok{width =} \FloatTok{0.75}\NormalTok{, }\DataTypeTok{size =} \FloatTok{0.33}\NormalTok{, }\DataTypeTok{outlier.shape =} \OtherTok{NA}\NormalTok{) +}
\StringTok{  }\KeywordTok{scale_fill_brewer}\NormalTok{(}\DataTypeTok{palette =} \StringTok{"Set1"}\NormalTok{) +}\StringTok{ }
\StringTok{  }\KeywordTok{coord_flip}\NormalTok{() +}
\StringTok{  }\KeywordTok{theme_bw}\NormalTok{() +}
\StringTok{  }\KeywordTok{labs}\NormalTok{(}
    \DataTypeTok{y=}\StringTok{'execution time (seconds)'}\NormalTok{,}
    \DataTypeTok{x=}\StringTok{''}\NormalTok{) +}
\StringTok{  }\KeywordTok{theme_classic}\NormalTok{() +}
\StringTok{  }\NormalTok{pub_theme +}
\StringTok{  }\KeywordTok{theme}\NormalTok{(}
    \DataTypeTok{legend.text =} \KeywordTok{element_text}\NormalTok{(}\DataTypeTok{size =} \DecValTok{18}\NormalTok{),}
    \DataTypeTok{legend.title =} \KeywordTok{element_blank}\NormalTok{(),}
    \DataTypeTok{legend.position =} \StringTok{"top"}\NormalTok{,}
    \DataTypeTok{axis.text.x =} \KeywordTok{element_text}\NormalTok{(}\DataTypeTok{angle =} \DecValTok{90}\NormalTok{)}
  \NormalTok{)}

\KeywordTok{plot_grid}\NormalTok{(plot_a, plot_b, }\DataTypeTok{align =} \StringTok{'h'}\NormalTok{, }\DataTypeTok{nrow =} \DecValTok{1}\NormalTok{, }\DataTypeTok{labels=}\StringTok{"AUTO"}\NormalTok{, }\DataTypeTok{label_size =} \DecValTok{28}\NormalTok{)}
\CommentTok{#> Warning in align_plots(plotlist = plots, align = align): Graphs cannot be}
\CommentTok{#> horizontally aligned. Placing graphs unaligned.}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{ggsave}\NormalTok{(}\StringTok{"figure3.pdf"}\NormalTok{, }\DataTypeTok{width =} \DecValTok{20}\NormalTok{, }\DataTypeTok{height =} \DecValTok{10}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\subsection{Code demonstrations}\label{code-demonstrations}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(valr)}
\CommentTok{# function to retrieve path to example data}
\NormalTok{bed_filepath <-}\StringTok{ }\KeywordTok{valr_example}\NormalTok{(}\StringTok{"3fields.bed.gz"}\NormalTok{) }
\KeywordTok{read_bed}\NormalTok{(bed_filepath) }
\CommentTok{#> # A tibble: 10 x 3}
\CommentTok{#>    chrom  start    end}
\CommentTok{#>    <chr>  <int>  <int>}
\CommentTok{#>  1  chr1  11873  14409}
\CommentTok{#>  2  chr1  14361  19759}
\CommentTok{#>  3  chr1  14406  29370}
\CommentTok{#>  4  chr1  34610  36081}
\CommentTok{#>  5  chr1  69090  70008}
\CommentTok{#>  6  chr1 134772 140566}
\CommentTok{#>  7  chr1 321083 321115}
\CommentTok{#>  8  chr1 321145 321207}
\CommentTok{#>  9  chr1 322036 326938}
\CommentTok{#> 10  chr1 327545 328439}

\CommentTok{#using URL}
\KeywordTok{read_bed}\NormalTok{(}\StringTok{"https://github.com/rnabioco/valr/raw/master/inst/extdata/3fields.bed.gz"}\NormalTok{)}
\CommentTok{#> # A tibble: 10 x 3}
\CommentTok{#>    chrom  start    end}
\CommentTok{#>    <chr>  <int>  <int>}
\CommentTok{#>  1  chr1  11873  14409}
\CommentTok{#>  2  chr1  14361  19759}
\CommentTok{#>  3  chr1  14406  29370}
\CommentTok{#>  4  chr1  34610  36081}
\CommentTok{#>  5  chr1  69090  70008}
\CommentTok{#>  6  chr1 134772 140566}
\CommentTok{#>  7  chr1 321083 321115}
\CommentTok{#>  8  chr1 321145 321207}
\CommentTok{#>  9  chr1 322036 326938}
\CommentTok{#> 10  chr1 327545 328439}
\end{Highlighting}
\end{Shaded}

Using mysql

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(dplyr, }\DataTypeTok{warn.conflicts =} \NormalTok{F)}
\NormalTok{hg19 <-}\StringTok{ }\KeywordTok{db_ucsc}\NormalTok{(}\StringTok{"hg19"}\NormalTok{)}
\KeywordTok{tbl}\NormalTok{(hg19, }\StringTok{"refGene"}\NormalTok{)}
\CommentTok{#> Warning in .local(conn, statement, ...): Unsigned INTEGER in col 0 imported}
\CommentTok{#> as numeric}
\CommentTok{#> Warning in .local(conn, statement, ...): Unsigned INTEGER in col 4 imported}
\CommentTok{#> as numeric}
\CommentTok{#> Warning in .local(conn, statement, ...): Unsigned INTEGER in col 5 imported}
\CommentTok{#> as numeric}
\CommentTok{#> Warning in .local(conn, statement, ...): Unsigned INTEGER in col 6 imported}
\CommentTok{#> as numeric}
\CommentTok{#> Warning in .local(conn, statement, ...): Unsigned INTEGER in col 7 imported}
\CommentTok{#> as numeric}
\CommentTok{#> Warning in .local(conn, statement, ...): Unsigned INTEGER in col 8 imported}
\CommentTok{#> as numeric}
\CommentTok{#> Warning in .local(conn, statement, ...): Unsigned INTEGER in col 0 imported}
\CommentTok{#> as numeric}
\CommentTok{#> Warning in .local(conn, statement, ...): Unsigned INTEGER in col 4 imported}
\CommentTok{#> as numeric}
\CommentTok{#> Warning in .local(conn, statement, ...): Unsigned INTEGER in col 5 imported}
\CommentTok{#> as numeric}
\CommentTok{#> Warning in .local(conn, statement, ...): Unsigned INTEGER in col 6 imported}
\CommentTok{#> as numeric}
\CommentTok{#> Warning in .local(conn, statement, ...): Unsigned INTEGER in col 7 imported}
\CommentTok{#> as numeric}
\CommentTok{#> Warning in .local(conn, statement, ...): Unsigned INTEGER in col 8 imported}
\CommentTok{#> as numeric}
\CommentTok{#> # Source:   table<refGene> [?? x 16]}
\CommentTok{#> # Database: mysql 5.6.26-log [genomep@genome-mysql.cse.ucsc.edu:/hg19]}
\CommentTok{#>      bin         name chrom strand   txStart     txEnd  cdsStart    cdsEnd}
\CommentTok{#>    <dbl>        <chr> <chr>  <chr>     <dbl>     <dbl>     <dbl>     <dbl>}
\CommentTok{#>  1   585    NR_024540  chr1      -     14361     29370     29370     29370}
\CommentTok{#>  2  1759    NR_110022  chrX      + 153991016 154005964 154005964 154005964}
\CommentTok{#>  3   604    NR_033380  chrX      +   2527305   2575270   2575270   2575270}
\CommentTok{#>  4    75    NR_033381  chrY      +   2477305   2506370   2506370   2506370}
\CommentTok{#>  5   915    NM_182508 chr13      +  43355685  43365685  43358203  43362926}
\CommentTok{#>  6   923    NM_181845 chr19      +  44331472  44353050  44337656  44352793}
\CommentTok{#>  7   586    NR_039983  chr1      -    134772    140566    140566    140566}
\CommentTok{#>  8  1964    NR_028327  chr5      + 180750506 180755196 180755196 180755196}
\CommentTok{#>  9   126 NM_001271875 chr17      -  55938603  56032684  55943837  55962925}
\CommentTok{#> 10  2278    NR_125989  chr1      - 222001007 222014008 222014008 222014008}
\CommentTok{#> # ... with more rows, and 8 more variables: exonCount <dbl>,}
\CommentTok{#> #   exonStarts <chr>, exonEnds <chr>, score <int>, name2 <chr>,}
\CommentTok{#> #   cdsStartStat <chr>, cdsEndStat <chr>, exonFrames <chr>}
\end{Highlighting}
\end{Shaded}

\subsubsection{Interval statistics}\label{interval-statistics}

Estimates of significance for interval overlaps can be obtained by
combining \texttt{bed\_shuffle()}, \texttt{bed\_random()} and the
\texttt{sample\_} functions from \texttt{dplyr} with interval statistics
in \texttt{valr}.

Here we examine the extent of overlap of repeat classes with exons in
the human genome (on \texttt{chr22} only, for simplicity) using the
jaccard similarity index. \texttt{bed\_jaccard()} implements the jaccard
test to examine the similarity between two sets of genomic intervals.
Using \texttt{bed\_shuffle()} and \texttt{replicate()} we generate a
\texttt{data\_frame} containing 100 sets of randomly placed intervals
then calculate the jaccard index for each set to generate a
null-distribution of jaccard scores. Finally an empirical p-value is
then calculated from the null-distribution.

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(tidyverse, }\DataTypeTok{warn.conflicts =} \NormalTok{F)}

\NormalTok{repeats <-}\StringTok{ }\KeywordTok{read_bed}\NormalTok{(}\KeywordTok{valr_example}\NormalTok{(}\StringTok{'hg19.rmsk.chr22.bed.gz'}\NormalTok{), }\DataTypeTok{n_fields =} \DecValTok{6}\NormalTok{) }
\NormalTok{genome <-}\StringTok{ }\KeywordTok{read_genome}\NormalTok{(}\KeywordTok{valr_example}\NormalTok{(}\StringTok{'hg19.chrom.sizes.gz'}\NormalTok{))}
\NormalTok{genes <-}\StringTok{ }\KeywordTok{read_bed12}\NormalTok{(}\KeywordTok{valr_example}\NormalTok{(}\StringTok{'hg19.refGene.chr22.bed.gz'}\NormalTok{))}
\CommentTok{# convert bed12 to bed with exons}
\NormalTok{exons <-}\StringTok{ }\KeywordTok{bed12_to_exons}\NormalTok{(genes)}

\CommentTok{# function to repeat interval shuffling}
\NormalTok{shuffle_intervals <-}\StringTok{ }\NormalTok{function(n, .data, genome) \{}
  \KeywordTok{replicate}\NormalTok{(n, }\KeywordTok{bed_shuffle}\NormalTok{(.data, genome), }\DataTypeTok{simplify =} \OtherTok{FALSE}\NormalTok{) %>%}
\StringTok{    }\KeywordTok{bind_rows}\NormalTok{(}\DataTypeTok{.id =} \StringTok{'rep'}\NormalTok{) %>%}
\StringTok{    }\KeywordTok{group_by}\NormalTok{(rep) %>%}\StringTok{ }\KeywordTok{nest}\NormalTok{()}
\NormalTok{\}}
\NormalTok{nreps <-}\StringTok{ }\DecValTok{100}
\NormalTok{shuffled <-}\StringTok{ }\KeywordTok{shuffle_intervals}\NormalTok{(}\DataTypeTok{n =} \NormalTok{nreps, repeats, genome) %>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{jaccard =} \NormalTok{data %>%}
\StringTok{           }\KeywordTok{map}\NormalTok{(bed_jaccard, repeats) %>%}
\StringTok{           }\KeywordTok{map_dbl}\NormalTok{(}\StringTok{"jaccard"}\NormalTok{))}
\NormalTok{shuffled}
\CommentTok{#> # A tibble: 100 x 3}
\CommentTok{#>      rep                  data      jaccard}
\CommentTok{#>    <chr>                <list>        <dbl>}
\CommentTok{#>  1     1 <tibble [10,000 x 6]> 0.0004337515}
\CommentTok{#>  2     2 <tibble [10,000 x 6]> 0.0003501136}
\CommentTok{#>  3     3 <tibble [10,000 x 6]> 0.0006080318}
\CommentTok{#>  4     4 <tibble [10,000 x 6]> 0.0005768465}
\CommentTok{#>  5     5 <tibble [10,000 x 6]> 0.0002537739}
\CommentTok{#>  6     6 <tibble [10,000 x 6]> 0.0004480599}
\CommentTok{#>  7     7 <tibble [10,000 x 6]> 0.0001777942}
\CommentTok{#>  8     8 <tibble [10,000 x 6]> 0.0003448734}
\CommentTok{#>  9     9 <tibble [10,000 x 6]> 0.0003303327}
\CommentTok{#> 10    10 <tibble [10,000 x 6]> 0.0003312376}
\CommentTok{#> # ... with 90 more rows}

\NormalTok{obs <-}\StringTok{ }\KeywordTok{bed_jaccard}\NormalTok{(repeats, exons)}
\NormalTok{obs}
\CommentTok{#> # A tibble: 1 x 4}
\CommentTok{#>    len_i   len_u    jaccard     n}
\CommentTok{#>    <dbl>   <dbl>      <dbl> <dbl>}
\CommentTok{#> 1 112123 4132109 0.02789139   805}

\NormalTok{pvalue <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(shuffled$jaccard >=}\StringTok{ }\NormalTok{obs$jaccard) +}\StringTok{ }\DecValTok{1} \NormalTok{/(nreps +}\StringTok{ }\DecValTok{1}\NormalTok{)}
\NormalTok{pvalue}
\CommentTok{#> [1] 0.00990099}
\end{Highlighting}
\end{Shaded}

\subsection{Correlations among DNase I hypersensitive
sites}\label{correlations-among-dnase-i-hypersensitive-sites}

Here we use \texttt{bed\_jaccard} for a large-scale comparison of
related datasets. As shown in the
\href{http://quinlanlab.org/tutorials/bedtools/bedtools.html}{\texttt{BEDtools}
tutorial}, we can measure the similarity of DNaseI hypersensitivity
sites for 20 fetal tissue samples.

This data was taken from
\href{www.sciencemag.org/content/337/6099/1190.short}{Maurano \emph{et
al.} Systematic Localization of Common Disease-Associated Variation in
Regulatory DNA. (2012) \emph{Science}}.

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(valr)}
\KeywordTok{library}\NormalTok{(dplyr)}
\KeywordTok{library}\NormalTok{(tidyr)}
\KeywordTok{library}\NormalTok{(purrr)}
\KeywordTok{library}\NormalTok{(broom)}
\KeywordTok{library}\NormalTok{(stringr)}
\KeywordTok{library}\NormalTok{(ggplot2)}
\KeywordTok{library}\NormalTok{(ggrepel)}
\KeywordTok{library}\NormalTok{(ComplexHeatmap)}
\end{Highlighting}
\end{Shaded}

First read all 20 BED files containing DNase I hypersensitivity sites
from 20 fetal tissues.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{dnase_files <-}\StringTok{ }\KeywordTok{list.files}\NormalTok{(}\StringTok{'data/dnasei'}\NormalTok{,}\DataTypeTok{pattern =} \StringTok{'merge.bed.gz'}\NormalTok{, }\DataTypeTok{full.names =} \OtherTok{TRUE}\NormalTok{)}
\NormalTok{data <-}\StringTok{ }\NormalTok{dnase_files %>%}\StringTok{ }\KeywordTok{map}\NormalTok{(read_bed, }\DataTypeTok{n_fields =} \DecValTok{4}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

Then generate a 20x20 table containing a Jaccard statistic for each of
the 400 pairwise comparisons.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{res <-}\StringTok{ }\NormalTok{data %>%}
\StringTok{  }\KeywordTok{cross2}\NormalTok{(.,.) %>%}
\StringTok{  }\KeywordTok{map}\NormalTok{(}\KeywordTok{lift}\NormalTok{(bed_jaccard)) %>%}\StringTok{ }
\StringTok{  }\KeywordTok{map}\NormalTok{(}\StringTok{"jaccard"}\NormalTok{) %>%}
\StringTok{  }\KeywordTok{flatten_dbl}\NormalTok{() %>%}
\StringTok{  }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{nrow =} \DecValTok{20}\NormalTok{, }\DataTypeTok{ncol =} \DecValTok{20}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

We also need to generate labels for the table from the file names.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# names are tissue + sample_num}
\NormalTok{col_names <-}\StringTok{ }\NormalTok{dnase_files %>%}
\StringTok{  }\KeywordTok{str_split}\NormalTok{(}\StringTok{'/'}\NormalTok{) %>%}\StringTok{ }\KeywordTok{map}\NormalTok{(}\StringTok{`}\DataTypeTok{[[}\StringTok{`}\NormalTok{, }\DecValTok{3}\NormalTok{) %>%}
\StringTok{  }\KeywordTok{str_split}\NormalTok{(}\StringTok{'}\CharTok{\textbackslash{}\textbackslash{}}\StringTok{.'}\NormalTok{) %>%}\StringTok{ }\KeywordTok{map}\NormalTok{(}\StringTok{`}\DataTypeTok{[[}\StringTok{`}\NormalTok{, }\DecValTok{1}\NormalTok{) %>%}
\StringTok{  }\KeywordTok{str_split}\NormalTok{(}\StringTok{'-'}\NormalTok{) %>%}\StringTok{ }\KeywordTok{map}\NormalTok{(}\StringTok{`}\DataTypeTok{[[}\StringTok{`}\NormalTok{, }\DecValTok{1}\NormalTok{) %>%}
\StringTok{  }\KeywordTok{flatten_chr}\NormalTok{() %>%}
\StringTok{  }\KeywordTok{str_replace}\NormalTok{(}\StringTok{'^f'}\NormalTok{, }\StringTok{''}\NormalTok{) %>%}
\StringTok{  }\KeywordTok{str_c}\NormalTok{(}\KeywordTok{str_c}\NormalTok{(}\StringTok{'-'}\NormalTok{, }\KeywordTok{seq}\NormalTok{(}\KeywordTok{length}\NormalTok{(.))))}

\KeywordTok{colnames}\NormalTok{(res) <-}\StringTok{ }\NormalTok{col_names}
\KeywordTok{rownames}\NormalTok{(res) <-}\StringTok{ }\NormalTok{col_names}
\end{Highlighting}
\end{Shaded}

Now the Jaccard coefficients can be visualized in heatmap form.

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{Heatmap}\NormalTok{(res, }\DataTypeTok{color_space =} \StringTok{'Blues'}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

Finally we can do some PCA analysis on the Jaccard coefficients to
identify clusters of related cell types.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{pca <-}\StringTok{ }\NormalTok{broom::}\KeywordTok{tidy}\NormalTok{(}\KeywordTok{prcomp}\NormalTok{(res)) %>%}\StringTok{ }\KeywordTok{as_data_frame}\NormalTok{()}

\NormalTok{pca_comps <-}\StringTok{ }\KeywordTok{filter}\NormalTok{(pca, PC <=}\StringTok{ }\DecValTok{2}\NormalTok{) %>%}
\StringTok{  }\NormalTok{tidyr::}\KeywordTok{spread}\NormalTok{(PC, value) %>%}
\StringTok{  }\KeywordTok{setNames}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\StringTok{'label'}\NormalTok{,}\StringTok{'PC1'}\NormalTok{,}\StringTok{'PC2'}\NormalTok{))}

\KeywordTok{ggplot}\NormalTok{(pca_comps, }\KeywordTok{aes}\NormalTok{(PC1, PC2)) +}
\StringTok{  }\KeywordTok{geom_point}\NormalTok{(}\DataTypeTok{size =} \DecValTok{3}\NormalTok{, }\DataTypeTok{color =} \StringTok{'red'}\NormalTok{) +}
\StringTok{  }\KeywordTok{geom_text_repel}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{label =} \NormalTok{label))}
\end{Highlighting}
\end{Shaded}

### 2017-06-14 inst/paper/main.tex
first draft

@@
 \begin{abstract}
-New tools for genome internal analysis enable analysis of large genomic data sets
-Availability of large genomic data sets. Need for new tools for exploratoratory data analysis. valr enables genomic interval analysis in the RStudio environment.
+New tools for genome interval analysis enable analysis of large genomic data sets.
+The size and complexity of genomic data continues to increase highlighting the need for novels tools for exploratory data analysis.
+\texttt{valr} enables flexible and efficient genomic interval analysis in the RStudio environment.
+\texttt{valr} can be used for interactive exploratory analyses or incorporated into existing analysis pipelines.
+
 \end{abstract}
@@
 
-Determining the correlation between sets of genomic features is a routine task in bioinformatics that requires sophisticated interval arithmetic.
-Command-line tools for interval analysis such as BEDtools \cite{quinlan_bedtools:_2010} and BEDOPS \cite{neph_bedops:_2012} are powerful tools that are central to many genomic analysis pipelines.
+A routine task in bioinformatics is determining the relationships between sets of genomic intervals.
+Frequent tasks include for example, identifying SNPs located within geneic regions, annotating regions enriched for specific nucleic acid binding proteins, or computing read density across a set of exons.
+Command-line tools for interval analysis such as BEDtools \cite{quinlan_bedtools:_2010} and BEDOPS \cite{neph_bedops:_2012} enable analyses of genome-wide datasets and are central components of analysis pipelines.
 A common workflow for genomic analysis begins with preprocessing interval analyses via command-line tools, followed by importing data into R for visualization and further statistical analyses.
@@
 
-R packages developed for interval analysis include IRanges \cite{lawrence_software_2013}, GenometriCorr \cite{favorov_exploring_2012}, and bedr \cite{haider_bedr_2016}.
-IRanges is a Bioconductor package that provides specialized interval classes and methods to conduct interval arthimetic and is utilized extensively by Bioconductor packages.
-GenometriCorr provides a set of statistical tests to determine the relationships between interval sets and builds upon IRanges.
-bedr is a CRAN distributed package that provides wrapper functions to call BEDTools, BedOPS, and tabix functions providing out-of-memory support for interval analysis in R.
-These packages provide important functionality for interval analysis in R however they do not easily integrate with interactive data visualization features developed for R, such as Shiny\cite{chang_shiny_2017}, and the RStudio IDE.
-Additionally, the programmatic advances provided by the recent tidyverse suite of data processing and visualization tools (dplyr, purrr, broom, and ggplot2) do not easily interface with the R S4 objects implemented in many Bioconductor packages for data organization \cite{wickham_tidyverse_2017}.
-
-We therefore sought to develop an R package that allows seamless and extensible interval arithmetic built to incorporate new R programming, visualization, and interactivity features.
+Previous R packages developed for interval analysis include IRanges \cite{lawrence_software_2013}, bedr \cite{haider_bedr_2016}, and GenometriCorr \cite{favorov_exploring_2012}.
+IRanges is a Bioconductor package that provides specialized interval classes and methods to perform interval arithmetic in R and is used by many Bioconductor packages.
+In contrast, bedr is a CRAN distributed package that provides wrapper R functions to call BEDTools, BedOPS, and tabix command-line utilities providing out-of-memory support for interval analysis in R.
+GenometriCorr provides a set of statistical tests to determine the relationships between interval sets using IRanges data structures.
+These packages provide important functionality for processing and statistical inference of interval data, but they do not easily integrate with interactive data visualization features developed for R, such as Shiny\cite{chang_shiny_2017}, and the RStudio IDE.
+Additionally, the R S4 data structures implemented by IRanges do not easily interface with the recent advances provided by the popular tidyverse suite of data processing and visualization tools (dplyr, purrr, broom, and ggplot2) \cite{wickham_tidyverse_2017}.
+We therefore sought to develop an flexible R package for genomic interval arithmetic built to incorporate new R programming, visualization, and interactivity features.
 
 \section*{Methods}
+
 \subsection*{Implementation}
-valr is implemented as an R package that makes extensive use of dplyr, a popular package that provides a flexible and high-performance framework for data manipulation \cite{wickham_dplyr_2016}.
-Computationally intensive functions in valr are written in C++ using Rcpp to enable fluid and quick interactive analysis of large datasets\cite{eddelbuettel_rcpp_2011}.
+valr is an R package that makes extensive use of dplyr, a flexible and high-performance framework for data manipulation in R \cite{wickham_dplyr_2016}.
+Compute intensive functions in valr are written in C++ using Rcpp to enable fluid interactive analysis of large datasets\cite{eddelbuettel_rcpp_2011}.
+Interval intersections and related operations utilize an interval tree algorithm to efficently process large datasets\cite{cormen_2001}.
+BED files are imported and handled in R as data\_frame objects, requiring minimal pre or post-processing to integrate with additional R packages or command line tools.
 
@@
 
-\section*{Use Cases} % Optional - only if NO new datasets are included
-To demonstrate the functionality and utility of \textbf{valr} presented below are some common use cases for genomic interval analysis.
+\section*{Use Cases}
+To demonstrate the functionality and utility of \textbf{valr}, presented below is a basic tutorial for using valr and additional common use cases for genomic interval analysis.
 
-\subsection{Basic Usage}
+\subsection*{Basic Usage}
 
-\subsubsection{Input data}\label{input-data}
+\subsubsection*{Input data}\label{input-data}
 
-\texttt{valr} provides a set of functions to read BED, BEDgraph, and VCF into R as convenient \texttt{tbl\_dfs()} dataframe objects.
-All tbls will have \texttt{chrom}, \texttt{start}, and\texttt{end} columns, and tbls from multi-column formats will have additional pre-determined column names.
+\texttt{valr} provides a set of functions to read BED, BEDgraph, and VCF formats into R as convenient \texttt{tbl\_dfs()} dataframe objects.
+All tbls will have \texttt{chrom}, \texttt{start}, and \texttt{end} columns, and tbls from multi-column formats will have additional pre-determined column names.
+Standards methods for importing data (e.g.
+\texttt{read.table}, \texttt{readr::read\_tsv}) are also supported provided the contructed dataframes contain the requiste column names (\texttt{chrom}, \texttt{start}, \texttt{end}).
+Additionally \texttt{valr} supports connections to remote databases to programatically access the UCSC and Ensembl databases via the \texttt{db\_ucsc} and \texttt{db\_ensembl} functions.
 
-Additionally \texttt{valr} supports connections to remote databases to programatically access the UCSC and Ensembl databases.
+\begin{Highlighting}[]
+\KeywordTok{library}\NormalTok{(valr)}
+\CommentTok{# function to retrieve path to example data}
+\NormalTok{bed_filepath <-}\StringTok{ }\KeywordTok{valr_example}\NormalTok{(}\StringTok{"3fields.bed.gz"}\NormalTok{) }
+\KeywordTok{read_bed}\NormalTok{(bed_filepath) }
+\CommentTok{#> # A tibble: 10 x 3}
+\CommentTok{#>    chrom  start    end}
+\CommentTok{#>    <chr>  <int>  <int>}
+\CommentTok{#>  1  chr1  11873  14409}
+\CommentTok{#>  2  chr1  14361  19759}
+\CommentTok{#>  3  chr1  14406  29370}
+\CommentTok{#>  4  chr1  34610  36081}
+\CommentTok{#>  5  chr1  69090  70008}
+\CommentTok{#>  6  chr1 134772 140566}
+\CommentTok{#>  7  chr1 321083 321115}
+\CommentTok{#>  8  chr1 321145 321207}
+\CommentTok{#>  9  chr1 322036 326938}
+\CommentTok{#> 10  chr1 327545 328439}
 
-Show examples of read bed with files, url, and readucsc.
+\CommentTok{#using URL}
+\KeywordTok{read_bed}\NormalTok{(}\StringTok{"https://github.com/rnabioco/valr/raw/master/inst/extdata/3fields.bed.gz"}\NormalTok{)}
+\CommentTok{#> # A tibble: 10 x 3}
+\CommentTok{#>    chrom  start    end}
+\CommentTok{#>    <chr>  <int>  <int>}
+\CommentTok{#>  1  chr1  11873  14409}
+\CommentTok{#>  2  chr1  14361  19759}
+\CommentTok{#>  3  chr1  14406  29370}
+\CommentTok{#>  4  chr1  34610  36081}
+\CommentTok{#>  5  chr1  69090  70008}
+\CommentTok{#>  6  chr1 134772 140566}
+\CommentTok{#>  7  chr1 321083 321115}
+\CommentTok{#>  8  chr1 321145 321207}
+\CommentTok{#>  9  chr1 322036 326938}
+\CommentTok{#> 10  chr1 327545 328439}
+\end{Highlighting}
 
-\subsection{Terse Syntax}
+\subsubsection*{Simple example}
 
@@
 the \texttt{BEDtools} suite.
-
-Similar to pybedtools\cite{dale_pybedtools:_2011}, a python wrapper for BEDTools, valr has a terse syntax. Shown below is an example of how to find all SNPs within 1 Kbp of genes using valr.
+Also, similar to pybedtools\cite{dale_pybedtools:_2011}, a python wrapper for \texttt{BEDTools}, \texttt{valr} has a terse syntax.
+For example, shown below is an example of how to find all intergenic SNPs within 1 kilobase of genes using \texttt{valr}.
 
 \begin{Highlighting}[]
-\KeywordTok{library}\NormalTok{(valr)}
 \KeywordTok{library}\NormalTok{(dplyr)}
@@
 
-\subsubsection{Visual documentation}\label{visual-documentation}
-
-The \texttt{bed\_glyph()} tool illustrates the results of operations in
-\texttt{valr}, similar to those found in the \texttt{BEDtools}
-documentation. Shown below is the code to produce glyphs that illustrate the result of intersecting \texttt{x}
-and \texttt{y} intervals with \texttt{bed\_intersect()} , and the result of merging \texttt{x} intervals with \texttt{bed\_merge()} (Figure 1 \ref{fig:Figure 1}).
+\subsubsection*{Visual documentation}\label{visual-documentation}
+By conducting interval arthimetic entirely in R, \texttt{valr} is also an effective teaching tool for introducing interval analysis to early-stage analysts without requiring familiarity with both command-line tools and R.
+To aid in demonstrating the interval operations available in \texttt{valr}, we developed the \texttt{bed\_glyph()} tool which produces plots demonstrating the input and output of operations in \texttt{valr} in a manner similar to those found in the \texttt{BEDtools} documentation.
+Shown below is the code required to produce glyphs displaying the results of intersecting \texttt{x} and \texttt{y} intervals with \texttt{bed\_intersect()} , and the result of merging \texttt{x} intervals with \texttt{bed\_merge()} (Figure 1 \ref{fig:Figure 1}).
 
@@
 \end{Highlighting}
+% * <jay.hesselberth@gmail.com> 2017-06-10T17:17:56.304Z:
+%
+% Is it possible to insert the figure inline in the document?
+% * KR
+% Yes, but they request that figures are listed at the end of the document.
+% Once the document is formatted by F1000 I think that the document will look similar to this article:
+% https://f1000research.com/articles/5-2122/v2
+% ^.
 
-\subsubsection{Reproducible reports}\label{reproducible-reports}
-
-\texttt{valr} can be used in RMarkdown documents to generate
-reproducible work-flows for data processing. Because \texttt{valr} is
-reasonably fast (see the \protect\hyperlink{benchmarks}{benchmarks}), we
-now use it in lieu of other tools for exploratory analysis of genomic
-data sets in R.
-
-Command-line tools like \texttt{BEDtools} and \texttt{bedops} can be
-used in reproducible workflows (e.g., with
-\href{https://bitbucket.org/snakemake/snakemake/wiki/Home}{\texttt{snakemake}}),
-but it is cumbersome to move from command-line tools to exploratory
-analysis and plotting software.
-\href{https://pythonhosted.org/pybedtools/}{\texttt{pybedtools}} can be
-used within \texttt{ipython\ notebooks} to accomplish a similar goal,
-but others have pointed out
-\href{https://www.r-bloggers.com/why-i-dont-like-jupyter-fka-ipython-notebook/}{issues
-with this approach}, including clunky version control. Because RMarkdown
-files are text files, they are readily kept under version control.
-Moreover, new features in RStudio, such as notebook viewing, and multiple language support enable similar functionality to \texttt{ipython}.
-
-\subsubsection{Grouping data}\label{grouping-data}
+\subsubsection*{Grouping data}\label{grouping-data}
 
@@
 
-\subsubsection{Column specification}\label{column-specification}
+\subsubsection*{Column specification}\label{column-specification}
 
@@
 
-\subsection{API}\label{api}
+\subsubsection*{API}\label{api}
 The major functions available in valr are shown in Table \ref{tab:tools}.
@@
 analyze the distribution of histone marks surrounding transcription
-start sites.
+start sites, using H3K4Me3 Chip-Seq data downloaded from ENCODE (Link to data here?).
 
@@
 \begin{Highlighting}[]
-\CommentTok{# `valr_example()` identifies the path of example files}
+\CommentTok{# `valr_example()` identifies the path of example files in the valr package}
 \NormalTok{bedfile <-}\StringTok{ }\KeywordTok{valr_example}\NormalTok{(}\StringTok{'genes.hg19.chr22.bed.gz'}\NormalTok{)}
@@
 Now we use the \texttt{.win\_id} group with \texttt{bed\_map()} to
-caluclate a sum by mapping \texttt{y} signals onto the intervals in
+calculate a sum by mapping \texttt{y} signals onto the intervals in
 \texttt{x}. These data are regrouped by \texttt{.win\_id} and a summary
@@
 
-Here we examine the overlap of repeat classes in the human genome (on
-\texttt{chr22} only, for simplicity) using \texttt{bed\_jaccard()}.
+Here we examine the extent of overlap of repeat classes with exons in
+the human genome (on \texttt{chr22} only, for simplicity) using the
+jaccard similarity index. \texttt{bed\_jaccard()} implements the jaccard
+test to examine the similarity between two sets of genomic intervals.
+Using \texttt{bed\_shuffle()} and \texttt{replicate()} we generate a
+\texttt{data\_frame} containing 100 sets of randomly placed intervals
+then calculate the jaccard index for each set to generate a
+null-distribution of jaccard scores. Finally an empirical p-value is
+then calculated from the null-distribution.
 
 \begin{Highlighting}[]
-\KeywordTok{library}\NormalTok{(purrr)}
-\KeywordTok{library}\NormalTok{(tidyr)}
+\KeywordTok{library}\NormalTok{(tidyverse, }\DataTypeTok{warn.conflicts =} \NormalTok{F)}
 
@@
 \NormalTok{genome <-}\StringTok{ }\KeywordTok{read_genome}\NormalTok{(}\KeywordTok{valr_example}\NormalTok{(}\StringTok{'hg19.chrom.sizes.gz'}\NormalTok{))}
+\NormalTok{genes <-}\StringTok{ }\KeywordTok{read_bed12}\NormalTok{(}\KeywordTok{valr_example}\NormalTok{(}\StringTok{'hg19.refGene.chr22.bed.gz'}\NormalTok{))}
+\CommentTok{# convert bed12 to bed with exons}
+\NormalTok{exons <-}\StringTok{ }\KeywordTok{bed12_to_exons}\NormalTok{(genes)}
 
+\CommentTok{# function to repeat interval shuffling}
 \NormalTok{shuffle_intervals <-}\StringTok{ }\NormalTok{function(n, .data, genome) \{}
-  \KeywordTok{replicate}\NormalTok{(n, }\KeywordTok{bed_shuffle}\NormalTok{(.data, genome, }\DataTypeTok{seed =} \DecValTok{1010486}\NormalTok{), }\DataTypeTok{simplify =} \OtherTok{FALSE}\NormalTok{) %>%}
+  \KeywordTok{replicate}\NormalTok{(n, }\KeywordTok{bed_shuffle}\NormalTok{(.data, genome), }\DataTypeTok{simplify =} \OtherTok{FALSE}\NormalTok{) %>%}
 \StringTok{    }\KeywordTok{bind_rows}\NormalTok{(}\DataTypeTok{.id =} \StringTok{'rep'}\NormalTok{) %>%}
@@
 \NormalTok{\}}
-
-\NormalTok{shuffled <-}\StringTok{ }\KeywordTok{shuffle_intervals}\NormalTok{(}\DataTypeTok{n =} \DecValTok{100}\NormalTok{, repeats, genome) %>%}
+\NormalTok{nreps <-}\StringTok{ }\DecValTok{100}
+\NormalTok{shuffled <-}\StringTok{ }\KeywordTok{shuffle_intervals}\NormalTok{(}\DataTypeTok{n =} \NormalTok{nreps, repeats, genome) %>%}
 \StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{jaccard =} \NormalTok{data %>%}
@@
 \StringTok{           }\KeywordTok{map_dbl}\NormalTok{(}\StringTok{"jaccard"}\NormalTok{))}
-
 \NormalTok{shuffled}
@@
 \CommentTok{#>    <chr>                <list>        <dbl>}
-\CommentTok{#>  1     1 <tibble [10,000 x 6]> 0.0004355887}
-\CommentTok{#>  2     2 <tibble [10,000 x 6]> 0.0004355887}
-\CommentTok{#>  3     3 <tibble [10,000 x 6]> 0.0004355887}
-\CommentTok{#>  4     4 <tibble [10,000 x 6]> 0.0004355887}
-\CommentTok{#>  5     5 <tibble [10,000 x 6]> 0.0004355887}
-\CommentTok{#>  6     6 <tibble [10,000 x 6]> 0.0004355887}
-\CommentTok{#>  7     7 <tibble [10,000 x 6]> 0.0004355887}
-\CommentTok{#>  8     8 <tibble [10,000 x 6]> 0.0004355887}
-\CommentTok{#>  9     9 <tibble [10,000 x 6]> 0.0004355887}
-\CommentTok{#> 10    10 <tibble [10,000 x 6]> 0.0004355887}
+\CommentTok{#>  1     1 <tibble [10,000 x 6]> 0.0004337515}
+\CommentTok{#>  2     2 <tibble [10,000 x 6]> 0.0003501136}
+\CommentTok{#>  3     3 <tibble [10,000 x 6]> 0.0006080318}
+\CommentTok{#>  4     4 <tibble [10,000 x 6]> 0.0005768465}
+\CommentTok{#>  5     5 <tibble [10,000 x 6]> 0.0002537739}
+\CommentTok{#>  6     6 <tibble [10,000 x 6]> 0.0004480599}
+\CommentTok{#>  7     7 <tibble [10,000 x 6]> 0.0001777942}
+\CommentTok{#>  8     8 <tibble [10,000 x 6]> 0.0003448734}
+\CommentTok{#>  9     9 <tibble [10,000 x 6]> 0.0003303327}
+\CommentTok{#> 10    10 <tibble [10,000 x 6]> 0.0003312376}
 \CommentTok{#> # ... with 90 more rows}
+
+\NormalTok{obs <-}\StringTok{ }\KeywordTok{bed_jaccard}\NormalTok{(repeats, exons)}
+\NormalTok{obs}
+\CommentTok{#> # A tibble: 1 x 4}
+\CommentTok{#>    len_i   len_u    jaccard     n}
+\CommentTok{#>    <dbl>   <dbl>      <dbl> <dbl>}
+\CommentTok{#> 1 112123 4132109 0.02789139   805}
+
+\NormalTok{pvalue <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(shuffled$jaccard >=}\StringTok{ }\NormalTok{obs$jaccard) +}\StringTok{ }\DecValTok{1} \NormalTok{/(nreps +}\StringTok{ }\DecValTok{1}\NormalTok{)}
+\NormalTok{pvalue}
+\CommentTok{#> [1] 0.00990099}
 \end{Highlighting}
-
-\subsection*{Analysis of DNA methylation}
 
@@
 In order to ensure that valr performs fast enough to enable interactive analysis, key functionality is implemented in C++.
-We benchmarked major valr functions against corresponding commands in BEDTools.
-valr operates on tables already preloaded into RAM, whereas BEDTools performs file-reading, processing, and writing.
-To compare valr against bedtools we generated two files containing 1 million random 1 Kbp intervals derived from the human genome.
+To test the speed of major valr functions we generated two dataframes containing 1 million random 1 kilobase intervals derived from the human genome.
+Most of the major valr functions complete execution in less than 1 second, demonstrating that valr can process large interval datasets efficiently (Figure 3A\ref{fig:Figure 3}).
+
+We also benchmarked major valr functions against corresponding commands in BEDTools.
+valr operates on data\_frames already loaded into RAM, whereas BEDTools performs file-reading, processing, and writing.
+To compare valr against BEDTools we generated two BED files containing 1 million random 1 kilobase intervals derived from the human genome.
 For valr functions, we timed reading the table into R (e.g.
@@
 For BEDTools commands we timed executing the command with the output written to /dev/null.
-valr functions performed similarly or slightly faster than BEDTools commands with the exception of the bed\_map and bed\_fisher (Figure 3\ref{fig:Figure 3}) .
+valr functions performed similarly or faster than BEDTools commands with the exception of the \texttt{bed\_map} and \texttt{bed\_fisher} (Figure 3B\ref{fig:Figure 3}).
+
+\subsection{Reproducible Reports and Interactive Visualizations}
+Command-line tools like \texttt{BEDtools} and \texttt{bedops} can be incorporated into reproducible workflows (e.g., with
+\href{https://bitbucket.org/snakemake/snakemake/wiki/Home}{\texttt{snakemake}}\cite{koster_snakemakescalable_2012}),
+but it is cumbersome to transistion from command-line tools to exploratory analysis and plotting software.
+RMarkdown documents are plain text files, amenable to version control, which provide an interface to generate feature rich PDF and HTML reports that combine text, executable code, and figures in a single document.
+\texttt{valr} can be used in RMarkdown documents to provide rapid documentation of exploratory data analyses and generate reproducible work-flows for data processing.
+Moreover, new features in RStudio, such as notebook viewing, and multiple language support enable similar functionality to another popular notebook platform \texttt{jupyter notebooks}.
+
+Additionally, \texttt{valr} seamlessly integrates into R shiny applications allowing for complex interactive visualizations relating to genomic interval analyses.
+We have developed a shiny application available on github \href{https://github.com/rnabioco/valrdata} that explores CHiP-Seq signal density surrounding transcription start sites and demonstrates the ease of implementing \texttt{valr} to power dynamic visualizations, (e.g ).
 
@@
 Additionally by providing an easy-to-use interface for interval arithmetic in R, valr is also a useful teaching tool to introduce the analyses necessary to investigate correlations between genomic intervals, without requiring familiarity with the command-line.
-
-bedr provides a complementary set of tools for working with datasets where RAM limits loading all data into R.
-IRanges, is a mature package, but is a different philsophy.
-valr provides some of the functionality of Genometricorr, nbut is extensible to additional statistical test.
+We envison that \texttt{valr} will help researchers quickly and reproducibly analyze genome interval datasets.
 
@@
 \section*{Author contributions}
-In order to give appropriate credit to each author of an article, the individual
-contributions of each author to the manuscript should be detailed in this section. We
-recommend using author initials and then stating briefly how they contributed.
+
+JR conceptualized the package and provided direction for software development.
+JR, KR, AG, and RS developed and tested the package.
+YY and CB contributed workflows and benchmarks for the package.
+All authors wrote the manuscript.
 
@@
 \section*{Grant information}
-This work was supported by the RNA Bioscience Initiative (funded by a Transformational Research Award from the University of Colorado School of Medicine) and a grant from the National Institutes of Health (XXX to J.H.)
+This work was supported by the RNA Bioscience Initiative (funded by a Transformational Research Award from the University of Colorado School of Medicine),  a grant from the National Institutes of Health (R35 GM119550 to J.H.), the Colorado Office of Economic Development and International Trade (CTGGI 2016-2096), the BioFrontiers Computing Core at the BioFrontiers Institute, University of Colorado at Boulder and the Intramural Research Program of the National Library of Medicine.
 
@@
 \caption{\label{fig:Figure 2}Meta-analysis of signals relative to genomic features with valr.
-\textnormal{Summarized coverage of human H3K4Me3 Chip-Seq coverage across positive strand transcription start sites.
+\textnormal{(A) Summarized coverage of human H3K4Me3 Chip-Seq coverage across positive strand transcription start sites.
 Data presented +/- SEM.}}
@@
 \centering
-\includegraphics[width=0.5\textwidth]{figure3.pdf}
-\caption{\label{fig:Figure 3} Performance of major valr functionality benchmarked against BEDTools \textnormal{Timings calculated using microbenchmark and GNU time commands for valr and BEDTools respectively (A).}}
+\includegraphics[width=1\textwidth]{figure3.pdf}
+\caption{\label{fig:Figure 3} Performance of major valr functionality.
+\textnormal{(A)  Timings calculated by performing 10 repetitions of indicated functions on dataframes containing 1 million random 1 kilobase x/y intervals generated using \texttt{bed\_random()}.
+(B) Timings for executing functions in BEDTools or equivalent functions in valr using the same interval sets as in (A).
+All BEDTools functions were written to /dev/null, and were timed using GNU \texttt{time}.
+Timings for valr functions in (B) include times for reading files using \texttt{read\_bed()} and were timed using the \texttt{microbenchmark} package.}}
+% * <jay.hesselberth@gmail.com> 2017-06-07T13:17:50.374Z:
+%
+% No bar charts! Make this a boxplot ideally.
+%
+% ^.
 \end{figure}
@@
 
-%%%%%%%%%%%% Previous Figures %%%%%%%%%%%
-
-%%%% TSS H3K4me3 example %%%%
-\begin{figure}
-\centering
-    \begin{subfigure}[b]{0.45\textwidth}
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{x <-}\StringTok{ }\KeywordTok{trbl_interval}\NormalTok{(}
-  \NormalTok{~chrom, ~start, ~end,}
-  \StringTok{'chr1'}\NormalTok{, }\DecValTok{25}\NormalTok{,     }\DecValTok{50}\NormalTok{,}
-  \StringTok{'chr1'}\NormalTok{, }\DecValTok{100}\NormalTok{,    }\DecValTok{125}
-\NormalTok{)}
-
-\NormalTok{y <-}\StringTok{ }\KeywordTok{trbl_interval}\NormalTok{(}
-  \NormalTok{~chrom, ~start, ~end,}
-  \StringTok{'chr1'}\NormalTok{, }\DecValTok{30}\NormalTok{,     }\DecValTok{75}
-\NormalTok{)}
-
-\KeywordTok{bed_glyph}\NormalTok{(}\KeywordTok{bed_intersect}\NormalTok{(x, y))}
-\end{Highlighting}
-\end{Shaded}
-        \label{fig:bed_glyph_code}
-    \end{subfigure}
-    ~
-\begin{subfigure}[b]{0.45\textwidth}
-        \includegraphics[width=\textwidth]{intersect_glyph-1.pdf}
-        \label{fig:bed_glyph_out}
-    \end{subfigure}
-\caption{\label{fig:your-figure}Visualizing interval intersections with valr.	 }
-\end{figure}
-
-\begin{figure}
-\centering
-    \begin{subfigure}[b]{\textwidth}
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{# `valr_example()` identifies the path of example files}
-\NormalTok{bedfile <-}\StringTok{ }\KeywordTok{valr_example}\NormalTok{(}\StringTok{'genes.hg19.chr22.bed.gz'}\NormalTok{)}
-\NormalTok{genomefile <-}\StringTok{ }\KeywordTok{valr_example}\NormalTok{(}\StringTok{'hg19.chrom.sizes.gz'}\NormalTok{)}
-\NormalTok{bgfile  <-}\StringTok{ }\KeywordTok{valr_example}\NormalTok{(}\StringTok{'hela.h3k4.chip.bg.gz'}\NormalTok{)}
-
-\NormalTok{genes <-}\StringTok{ }\KeywordTok{read_bed}\NormalTok{(bedfile, }\DataTypeTok{n_fields =} \DecValTok{6}\NormalTok{)}
-\NormalTok{genome <-}\StringTok{ }\KeywordTok{read_genome}\NormalTok{(genomefile)}
-\NormalTok{y <-}\StringTok{ }\KeywordTok{read_bedgraph}\NormalTok{(bgfile)}
-
-\CommentTok{# generate 1 bp TSS intervals, `+` strand only}
-\NormalTok{tss <-}\StringTok{ }\NormalTok{genes %>%}\StringTok{ }\KeywordTok{filter}\NormalTok{(strand ==}\StringTok{ '+'}\NormalTok{) %>%}\StringTok{ }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{end =} \NormalTok{start +}\StringTok{ }\DecValTok{1}\NormalTok{)}
-
-\CommentTok{# 1000 bp up and downstream}
-\NormalTok{region_size <-}\StringTok{ }\DecValTok{1000}
-
-\CommentTok{# 50 bp windows}
-\NormalTok{win_size <-}\StringTok{ }\DecValTok{50}
-
-\CommentTok{# add slop to the TSS, break into windows and add a group}
-\NormalTok{x <-}\StringTok{ }\NormalTok{tss %>%}\StringTok{ }\KeywordTok{bed_slop}\NormalTok{(genome, }\DataTypeTok{both =} \NormalTok{region_size) %>%}
-\StringTok{  }\KeywordTok{bed_makewindows}\NormalTok{(genome, win_size)}
-
-\CommentTok{# map signals to TSS regions and calculate summary statistics.}
-\NormalTok{res <-}\StringTok{ }\KeywordTok{bed_map}\NormalTok{(x, y, }\DataTypeTok{win_sum =} \KeywordTok{sum}\NormalTok{(value, }\DataTypeTok{na.rm =} \OtherTok{TRUE}\NormalTok{)) %>%}
-\StringTok{  }\KeywordTok{group_by}\NormalTok{(.win_id) %>%}
-\StringTok{  }\KeywordTok{summarize}\NormalTok{(}\DataTypeTok{win_mean =} \KeywordTok{mean}\NormalTok{(win_sum, }\DataTypeTok{na.rm =} \OtherTok{TRUE}\NormalTok{),}
-            \DataTypeTok{win_sd =} \KeywordTok{sd}\NormalTok{(win_sum, }\DataTypeTok{na.rm =} \OtherTok{TRUE}\NormalTok{))}
-
-\NormalTok{x_labels <-}\StringTok{ }\KeywordTok{seq}\NormalTok{(-region_size, region_size, }\DataTypeTok{by =} \NormalTok{win_size *}\StringTok{ }\DecValTok{5}\NormalTok{)}
-\NormalTok{x_breaks <-}\StringTok{ }\KeywordTok{seq}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{41}\NormalTok{, }\DataTypeTok{by =} \DecValTok{5}\NormalTok{)}
-
-\NormalTok{sd_limits <-}\StringTok{ }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{ymax =} \NormalTok{win_mean +}\StringTok{ }\NormalTok{win_sd, }\DataTypeTok{ymin =} \NormalTok{win_mean -}\StringTok{ }\NormalTok{win_sd)}
-
-\KeywordTok{ggplot}\NormalTok{(res, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x =} \NormalTok{.win_id, }\DataTypeTok{y =} \NormalTok{win_mean)) +}\StringTok{ }\KeywordTok{geom_point}\NormalTok{() +}\StringTok{ }\KeywordTok{ylab}\NormalTok{(}\StringTok{'Signal'}\NormalTok{) +}
-\StringTok{  }\KeywordTok{scale_x_continuous}\NormalTok{(}\DataTypeTok{labels =} \NormalTok{x_labels, }\DataTypeTok{breaks =} \NormalTok{x_breaks) +}\StringTok{ }
-\StringTok{  }\KeywordTok{ggtitle}\NormalTok{(}\StringTok{'Human H3K4me3 signal near transcription start sites'}\NormalTok{) +}
-\StringTok{  }\KeywordTok{xlab}\NormalTok{(}\StringTok{'Position (bp from TSS)'}\NormalTok{) +}\StringTok{ }\KeywordTok{geom_pointrange}\NormalTok{(sd_limits) +}
-\StringTok{  }\KeywordTok{theme_classic}\NormalTok{()}
-\end{Highlighting}
-\end{Shaded}
-        \label{fig:meta_gene_code}
-    \end{subfigure}
-
-\begin{subfigure}[b]{0.5\textwidth}
-        \includegraphics[width=\textwidth]{demo-tss-1.pdf}
-        \label{fig:meta_gene_out}
-    \end{subfigure}
-\caption{\label{fig:your-figure}Meta-analysis of signals relative to genomic features with valr.}
-\end{figure}
-
 % See this guide for more information on BibTeX:
@@
 
-%%%% DNA methylation example %%%%
-
-\subsection{Get the public datasets from
-UCSC}\label{get-the-public-datasets-from-ucsc}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{library}\NormalTok{(dplyr)}
-\KeywordTok{library}\NormalTok{(valr)}
-\KeywordTok{library}\NormalTok{(R.utils)}
-\KeywordTok{library}\NormalTok{(curl)}
-\KeywordTok{library}\NormalTok{(ggplot2)}
-\KeywordTok{library}\NormalTok{(gridExtra)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{DMSO_treated<-}\KeywordTok{read_bed}\NormalTok{(}\StringTok{'http://hgdownload.soe.ucsc.edu/goldenPath/hg19/encodeDCC/wgEncodeHaibMethylRrbs/wgEncodeHaibMethylRrbsA549Dm002p7dHaibSitesRep1.bed.gz'}\NormalTok{, }\DataTypeTok{n_fields=}\DecValTok{6}\NormalTok{, }\DataTypeTok{skip=}\DecValTok{1}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{untreated<-}\KeywordTok{read_bed}\NormalTok{(}\StringTok{'http://hgdownload.soe.ucsc.edu/goldenPath/hg19/encodeDCC/wgEncodeHaibMethylRrbs/wgEncodeHaibMethylRrbsAg04449UwSitesRep1.bed.gz'}\NormalTok{, }\DataTypeTok{n_fields=}\DecValTok{6}\NormalTok{, }\DataTypeTok{skip=}\DecValTok{1}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-To view the data
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{head}\NormalTok{(untreated)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-#> # A tibble: 6 x 6
-#>   chrom start   end           name score strand
-#>   <chr> <int> <int>          <chr> <chr>  <chr>
-#> 1  chr1 10785 10786 AG04449_0__JS_     1      +
-#> 2  chr1 10788 10789 AG04449_0__JS_     1      +
-#> 3  chr1 10794 10795 AG04449_0__JS_     1      +
-#> 4  chr1 10810 10811 AG04449_0__JS_     1      +
-#> 5  chr1 10812 10813 AG04449_0__JS_     1      +
-#> 6  chr1 10815 10816 AG04449_0__JS_     1      +
-\end{verbatim}
-
-\subsection{Examine regions that are different and regions that
-intersect}\label{examine-regions-that-are-different-and-regions-that-intersect}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{different_region<-}\KeywordTok{bed_subtract}\NormalTok{(DMSO_treated,untreated, }\DataTypeTok{any=}\OtherTok{FALSE}\NormalTok{)}
-\KeywordTok{dim}\NormalTok{(different_region)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-#> [1] 278241      6
-\end{verbatim}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{intersect_region<-}\KeywordTok{bed_intersect}\NormalTok{(DMSO_treated,untreated, }\DataTypeTok{suffix=}\KeywordTok{c}\NormalTok{(}\StringTok{'.treated'}\NormalTok{,}\StringTok{'.untreated'}\NormalTok{))}
-\KeywordTok{dim}\NormalTok{(intersect_region)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-#> [1] 1576400      12
-\end{verbatim}
-
-\subsection{Plot the distribution of different regions and intersected
-regions across
-chromosomes}\label{plot-the-distribution-of-different-regions-and-intersected-regions-across-chromosomes}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{p1<-}\KeywordTok{ggplot}\NormalTok{(}\DataTypeTok{data=}\NormalTok{different_region, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x=}\NormalTok{chrom))+}\KeywordTok{geom_histogram}\NormalTok{(}\DataTypeTok{stat=}\StringTok{'count'}\NormalTok{, }\KeywordTok{aes}\NormalTok{(}\KeywordTok{factor}\NormalTok{(chrom), }\DataTypeTok{fill=}\StringTok{'differential methylation'}\NormalTok{))+}\KeywordTok{scale_fill_brewer}\NormalTok{(}\DataTypeTok{palette =} \StringTok{'Set2'}\NormalTok{)+}\KeywordTok{theme_minimal}\NormalTok{()+}\KeywordTok{ggtitle}\NormalTok{(}\StringTok{'Distribution of differential methylation across chromosomes'}\NormalTok{)+}\KeywordTok{theme}\NormalTok{(}\DataTypeTok{axis.text.x =} \KeywordTok{element_text}\NormalTok{(}\DataTypeTok{angle =} \DecValTok{45}\NormalTok{, }\DataTypeTok{hjust =} \DecValTok{1}\NormalTok{))}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-#> Warning: Ignoring unknown parameters: binwidth, bins, pad
-\end{verbatim}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{p2<-}\KeywordTok{ggplot}\NormalTok{(}\DataTypeTok{data=}\NormalTok{intersect_region, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x=}\NormalTok{chrom))+}\KeywordTok{geom_histogram}\NormalTok{(}\DataTypeTok{stat=}\StringTok{'count'}\NormalTok{, }\KeywordTok{aes}\NormalTok{(}\KeywordTok{factor}\NormalTok{(chrom), }\DataTypeTok{fill=}\StringTok{'overlap methylation'}\NormalTok{))+}\KeywordTok{scale_fill_brewer}\NormalTok{(}\DataTypeTok{palette =} \StringTok{'Spectral'}\NormalTok{)+}\KeywordTok{theme_minimal}\NormalTok{()+}\KeywordTok{ggtitle}\NormalTok{(}\StringTok{'Distribution of overlapping methylation across chromosomes'}\NormalTok{)+}\KeywordTok{theme}\NormalTok{(}\DataTypeTok{axis.text.x =} \KeywordTok{element_text}\NormalTok{(}\DataTypeTok{angle =} \DecValTok{45}\NormalTok{, }\DataTypeTok{hjust =} \DecValTok{1}\NormalTok{))}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-#> Warning: Ignoring unknown parameters: binwidth, bins, pad
-\end{verbatim}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{grid.arrange}\NormalTok{(p1, p2, }\DataTypeTok{ncol =} \DecValTok{1}\NormalTok{, }\DataTypeTok{nrow =}\DecValTok{2}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-\begin{figure}
-\begin{center}\includegraphics{unnamed-chunk-8-1} \end{center}
-\end{figure}
-
 \end{document}

### 2017-06-14 inst/paper/old_figs.tex
first draft

%%%% old figures for valr manuscript %%%%

%%%%%%%%%%%% Previous Figures %%%%%%%%%%%

%%%% TSS H3K4me3 example %%%%
\begin{figure}
\centering
    \begin{subfigure}[b]{0.45\textwidth}
\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{x <-}\StringTok{ }\KeywordTok{trbl_interval}\NormalTok{(}
  \NormalTok{~chrom, ~start, ~end,}
  \StringTok{'chr1'}\NormalTok{, }\DecValTok{25}\NormalTok{,     }\DecValTok{50}\NormalTok{,}
  \StringTok{'chr1'}\NormalTok{, }\DecValTok{100}\NormalTok{,    }\DecValTok{125}
\NormalTok{)}

\NormalTok{y <-}\StringTok{ }\KeywordTok{trbl_interval}\NormalTok{(}
  \NormalTok{~chrom, ~start, ~end,}
  \StringTok{'chr1'}\NormalTok{, }\DecValTok{30}\NormalTok{,     }\DecValTok{75}
\NormalTok{)}

\KeywordTok{bed_glyph}\NormalTok{(}\KeywordTok{bed_intersect}\NormalTok{(x, y))}
\end{Highlighting}
\end{Shaded}
        \label{fig:bed_glyph_code}
    \end{subfigure}
    ~
\begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{intersect_glyph-1.pdf}
        \label{fig:bed_glyph_out}
    \end{subfigure}
\caption{\label{fig:your-figure}Visualizing interval intersections with valr.	 }
\end{figure}

\begin{figure}
\centering
    \begin{subfigure}[b]{\textwidth}
\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# `valr_example()` identifies the path of example files}
\NormalTok{bedfile <-}\StringTok{ }\KeywordTok{valr_example}\NormalTok{(}\StringTok{'genes.hg19.chr22.bed.gz'}\NormalTok{)}
\NormalTok{genomefile <-}\StringTok{ }\KeywordTok{valr_example}\NormalTok{(}\StringTok{'hg19.chrom.sizes.gz'}\NormalTok{)}
\NormalTok{bgfile  <-}\StringTok{ }\KeywordTok{valr_example}\NormalTok{(}\StringTok{'hela.h3k4.chip.bg.gz'}\NormalTok{)}

\NormalTok{genes <-}\StringTok{ }\KeywordTok{read_bed}\NormalTok{(bedfile, }\DataTypeTok{n_fields =} \DecValTok{6}\NormalTok{)}
\NormalTok{genome <-}\StringTok{ }\KeywordTok{read_genome}\NormalTok{(genomefile)}
\NormalTok{y <-}\StringTok{ }\KeywordTok{read_bedgraph}\NormalTok{(bgfile)}

\CommentTok{# generate 1 bp TSS intervals, `+` strand only}
\NormalTok{tss <-}\StringTok{ }\NormalTok{genes %>%}\StringTok{ }\KeywordTok{filter}\NormalTok{(strand ==}\StringTok{ '+'}\NormalTok{) %>%}\StringTok{ }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{end =} \NormalTok{start +}\StringTok{ }\DecValTok{1}\NormalTok{)}

\CommentTok{# 1000 bp up and downstream}
\NormalTok{region_size <-}\StringTok{ }\DecValTok{1000}

\CommentTok{# 50 bp windows}
\NormalTok{win_size <-}\StringTok{ }\DecValTok{50}

\CommentTok{# add slop to the TSS, break into windows and add a group}
\NormalTok{x <-}\StringTok{ }\NormalTok{tss %>%}\StringTok{ }\KeywordTok{bed_slop}\NormalTok{(genome, }\DataTypeTok{both =} \NormalTok{region_size) %>%}
\StringTok{  }\KeywordTok{bed_makewindows}\NormalTok{(genome, win_size)}

\CommentTok{# map signals to TSS regions and calculate summary statistics.}
\NormalTok{res <-}\StringTok{ }\KeywordTok{bed_map}\NormalTok{(x, y, }\DataTypeTok{win_sum =} \KeywordTok{sum}\NormalTok{(value, }\DataTypeTok{na.rm =} \OtherTok{TRUE}\NormalTok{)) %>%}
\StringTok{  }\KeywordTok{group_by}\NormalTok{(.win_id) %>%}
\StringTok{  }\KeywordTok{summarize}\NormalTok{(}\DataTypeTok{win_mean =} \KeywordTok{mean}\NormalTok{(win_sum, }\DataTypeTok{na.rm =} \OtherTok{TRUE}\NormalTok{),}
            \DataTypeTok{win_sd =} \KeywordTok{sd}\NormalTok{(win_sum, }\DataTypeTok{na.rm =} \OtherTok{TRUE}\NormalTok{))}

\NormalTok{x_labels <-}\StringTok{ }\KeywordTok{seq}\NormalTok{(-region_size, region_size, }\DataTypeTok{by =} \NormalTok{win_size *}\StringTok{ }\DecValTok{5}\NormalTok{)}
\NormalTok{x_breaks <-}\StringTok{ }\KeywordTok{seq}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{41}\NormalTok{, }\DataTypeTok{by =} \DecValTok{5}\NormalTok{)}

\NormalTok{sd_limits <-}\StringTok{ }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{ymax =} \NormalTok{win_mean +}\StringTok{ }\NormalTok{win_sd, }\DataTypeTok{ymin =} \NormalTok{win_mean -}\StringTok{ }\NormalTok{win_sd)}

\KeywordTok{ggplot}\NormalTok{(res, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x =} \NormalTok{.win_id, }\DataTypeTok{y =} \NormalTok{win_mean)) +}\StringTok{ }\KeywordTok{geom_point}\NormalTok{() +}\StringTok{ }\KeywordTok{ylab}\NormalTok{(}\StringTok{'Signal'}\NormalTok{) +}
\StringTok{  }\KeywordTok{scale_x_continuous}\NormalTok{(}\DataTypeTok{labels =} \NormalTok{x_labels, }\DataTypeTok{breaks =} \NormalTok{x_breaks) +}\StringTok{ }
\StringTok{  }\KeywordTok{ggtitle}\NormalTok{(}\StringTok{'Human H3K4me3 signal near transcription start sites'}\NormalTok{) +}
\StringTok{  }\KeywordTok{xlab}\NormalTok{(}\StringTok{'Position (bp from TSS)'}\NormalTok{) +}\StringTok{ }\KeywordTok{geom_pointrange}\NormalTok{(sd_limits) +}
\StringTok{  }\KeywordTok{theme_classic}\NormalTok{()}
\end{Highlighting}
\end{Shaded}
        \label{fig:meta_gene_code}
    \end{subfigure}

\begin{subfigure}[b]{0.5\textwidth}
        \includegraphics[width=\textwidth]{demo-tss-1.pdf}
        \label{fig:meta_gene_out}
    \end{subfigure}
\caption{\label{fig:your-figure}Meta-analysis of signals relative to genomic features with valr.}
\end{figure}

%%%% DNA methylation example %%%%

\subsection{Get the public datasets from
UCSC}\label{get-the-public-datasets-from-ucsc}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(dplyr)}
\KeywordTok{library}\NormalTok{(valr)}
\KeywordTok{library}\NormalTok{(R.utils)}
\KeywordTok{library}\NormalTok{(curl)}
\KeywordTok{library}\NormalTok{(ggplot2)}
\KeywordTok{library}\NormalTok{(gridExtra)}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{DMSO_treated<-}\KeywordTok{read_bed}\NormalTok{(}\StringTok{'http://hgdownload.soe.ucsc.edu/goldenPath/hg19/encodeDCC/wgEncodeHaibMethylRrbs/wgEncodeHaibMethylRrbsA549Dm002p7dHaibSitesRep1.bed.gz'}\NormalTok{, }\DataTypeTok{n_fields=}\DecValTok{6}\NormalTok{, }\DataTypeTok{skip=}\DecValTok{1}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{untreated<-}\KeywordTok{read_bed}\NormalTok{(}\StringTok{'http://hgdownload.soe.ucsc.edu/goldenPath/hg19/encodeDCC/wgEncodeHaibMethylRrbs/wgEncodeHaibMethylRrbsAg04449UwSitesRep1.bed.gz'}\NormalTok{, }\DataTypeTok{n_fields=}\DecValTok{6}\NormalTok{, }\DataTypeTok{skip=}\DecValTok{1}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

To view the data

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{head}\NormalTok{(untreated)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
#> # A tibble: 6 x 6
#>   chrom start   end           name score strand
#>   <chr> <int> <int>          <chr> <chr>  <chr>
#> 1  chr1 10785 10786 AG04449_0__JS_     1      +
#> 2  chr1 10788 10789 AG04449_0__JS_     1      +
#> 3  chr1 10794 10795 AG04449_0__JS_     1      +
#> 4  chr1 10810 10811 AG04449_0__JS_     1      +
#> 5  chr1 10812 10813 AG04449_0__JS_     1      +
#> 6  chr1 10815 10816 AG04449_0__JS_     1      +
\end{verbatim}

\subsection{Examine regions that are different and regions that
intersect}\label{examine-regions-that-are-different-and-regions-that-intersect}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{different_region<-}\KeywordTok{bed_subtract}\NormalTok{(DMSO_treated,untreated, }\DataTypeTok{any=}\OtherTok{FALSE}\NormalTok{)}
\KeywordTok{dim}\NormalTok{(different_region)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
#> [1] 278241      6
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{intersect_region<-}\KeywordTok{bed_intersect}\NormalTok{(DMSO_treated,untreated, }\DataTypeTok{suffix=}\KeywordTok{c}\NormalTok{(}\StringTok{'.treated'}\NormalTok{,}\StringTok{'.untreated'}\NormalTok{))}
\KeywordTok{dim}\NormalTok{(intersect_region)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
#> [1] 1576400      12
\end{verbatim}

\subsection{Plot the distribution of different regions and intersected
regions across
chromosomes}\label{plot-the-distribution-of-different-regions-and-intersected-regions-across-chromosomes}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{p1<-}\KeywordTok{ggplot}\NormalTok{(}\DataTypeTok{data=}\NormalTok{different_region, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x=}\NormalTok{chrom))+}\KeywordTok{geom_histogram}\NormalTok{(}\DataTypeTok{stat=}\StringTok{'count'}\NormalTok{, }\KeywordTok{aes}\NormalTok{(}\KeywordTok{factor}\NormalTok{(chrom), }\DataTypeTok{fill=}\StringTok{'differential methylation'}\NormalTok{))+}\KeywordTok{scale_fill_brewer}\NormalTok{(}\DataTypeTok{palette =} \StringTok{'Set2'}\NormalTok{)+}\KeywordTok{theme_minimal}\NormalTok{()+}\KeywordTok{ggtitle}\NormalTok{(}\StringTok{'Distribution of differential methylation across chromosomes'}\NormalTok{)+}\KeywordTok{theme}\NormalTok{(}\DataTypeTok{axis.text.x =} \KeywordTok{element_text}\NormalTok{(}\DataTypeTok{angle =} \DecValTok{45}\NormalTok{, }\DataTypeTok{hjust =} \DecValTok{1}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
#> Warning: Ignoring unknown parameters: binwidth, bins, pad
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{p2<-}\KeywordTok{ggplot}\NormalTok{(}\DataTypeTok{data=}\NormalTok{intersect_region, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x=}\NormalTok{chrom))+}\KeywordTok{geom_histogram}\NormalTok{(}\DataTypeTok{stat=}\StringTok{'count'}\NormalTok{, }\KeywordTok{aes}\NormalTok{(}\KeywordTok{factor}\NormalTok{(chrom), }\DataTypeTok{fill=}\StringTok{'overlap methylation'}\NormalTok{))+}\KeywordTok{scale_fill_brewer}\NormalTok{(}\DataTypeTok{palette =} \StringTok{'Spectral'}\NormalTok{)+}\KeywordTok{theme_minimal}\NormalTok{()+}\KeywordTok{ggtitle}\NormalTok{(}\StringTok{'Distribution of overlapping methylation across chromosomes'}\NormalTok{)+}\KeywordTok{theme}\NormalTok{(}\DataTypeTok{axis.text.x =} \KeywordTok{element_text}\NormalTok{(}\DataTypeTok{angle =} \DecValTok{45}\NormalTok{, }\DataTypeTok{hjust =} \DecValTok{1}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
#> Warning: Ignoring unknown parameters: binwidth, bins, pad
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{grid.arrange}\NormalTok{(p1, p2, }\DataTypeTok{ncol =} \DecValTok{1}\NormalTok{, }\DataTypeTok{nrow =}\DecValTok{2}\NormalTok{)}
\end{Highlighting}
\end{Shaded}
\begin{figure}
\begin{center}\includegraphics{unnamed-chunk-8-1} \end{center}
\end{figure}

### 2017-06-14 inst/paper/maketex.R

@@ -1,0 +1,8 @@
+
+#render to markdown
+rmarkdown::render("~/src/rnabioco/valr/inst/paper/figs.Rmd",
+                  output_format = "html_document",
+                  clean = F)
+
+# render to latex
+rmarkdown::pandoc_convert("~/src/rnabioco/valr/inst/paper/figs.knit.md", to = "latex", output = "figs.tex")

### 2017-06-15 inst/paper/main.tex
version to send to all authors

@@
 \begin{abstract}
-New tools for genome interval analysis enable analysis of large genomic data sets.
-The size and complexity of genomic data continues to increase highlighting the need for novels tools for exploratory data analysis.
-\texttt{valr} enables flexible and efficient genomic interval analysis in the RStudio environment.
-\texttt{valr} can be used for interactive exploratory analyses or incorporated into existing analysis pipelines.
+New tools for reproducible exploratory data analysis of large data sets are important to address the rising size and complexity of genomic data.
+We developed the \texttt{valr} R package to enable flexible and efficient genomic interval analysis.
+\texttt{valr} leverages new tools available in the "tidyverse" including \texttt{dplyr}.
+Benchmarks of \texttt{valr} show it performs similar to \texttt{BEDtools} and can be used for interactive  analyses and incorporated into existing analysis pipelines.
 
@@
 
-A routine task in bioinformatics is determining the relationships between sets of genomic intervals.
-Frequent tasks include for example, identifying SNPs located within geneic regions, annotating regions enriched for specific nucleic acid binding proteins, or computing read density across a set of exons.
-Command-line tools for interval analysis such as BEDtools \cite{quinlan_bedtools:_2010} and BEDOPS \cite{neph_bedops:_2012} enable analyses of genome-wide datasets and are central components of analysis pipelines.
-A common workflow for genomic analysis begins with preprocessing interval analyses via command-line tools, followed by importing data into R for visualization and further statistical analyses.
-However, this workflow separation hinders exploratory data analysis, and the development reproducible research workflows built in the RMarkdown framework.
+	A routine bioinformatic task is the analysis of the relationships between sets of genomic intervals including the identification of DNA variants within genetic regions, annotation of regions enriched for nucleic acid binding proteins, and computation of read density within a set of exons.
+Command-line tools for interval analysis such as BEDtools \cite{quinlan_bedtools:_2010} and BEDOPS \cite{neph_bedops:_2012} enable analyses of genome-wide datasets and are keys components of analysis pipelines.
+Analyses with these tools commonly combine processing intervals on the command-line with visualization and statistical analysis in R.
+However, the need to master both the command-line and R hinders exploratory data analysis, and the development reproducible research workflows built in the RMarkdown framework.
 
-Previous R packages developed for interval analysis include IRanges \cite{lawrence_software_2013}, bedr \cite{haider_bedr_2016}, and GenometriCorr \cite{favorov_exploring_2012}.
-IRanges is a Bioconductor package that provides specialized interval classes and methods to perform interval arithmetic in R and is used by many Bioconductor packages.
-In contrast, bedr is a CRAN distributed package that provides wrapper R functions to call BEDTools, BedOPS, and tabix command-line utilities providing out-of-memory support for interval analysis in R.
-GenometriCorr provides a set of statistical tests to determine the relationships between interval sets using IRanges data structures.
-These packages provide important functionality for processing and statistical inference of interval data, but they do not easily integrate with interactive data visualization features developed for R, such as Shiny\cite{chang_shiny_2017}, and the RStudio IDE.
-Additionally, the R S4 data structures implemented by IRanges do not easily interface with the recent advances provided by the popular tidyverse suite of data processing and visualization tools (dplyr, purrr, broom, and ggplot2) \cite{wickham_tidyverse_2017}.
-We therefore sought to develop an flexible R package for genomic interval arithmetic built to incorporate new R programming, visualization, and interactivity features.
+	Existing R packages developed for interval analysis include IRanges \cite{lawrence_software_2013}, bedr \cite{haider_bedr_2016}, and GenometriCorr \cite{favorov_exploring_2012}.
+IRanges is a Bioconductor package that provides  interval classes and methods to perform interval arithmetic and is used by many Bioconductor packages.
+bedr is a CRAN-distributed package that provides wrapper R functions to call the BEDTools, BedOPS, and tabix command-line utilities, providing out-of-memory support for interval analysis.
+Finally,  GenometriCorr provides a set of statistical tests to determine the relationships between interval sets using IRanges data structures.
+These packages provide  functionality for processing and statistical inference of interval data, but they do not easily integrate with interactive data visualization features developed for R including Shiny\cite{chang_shiny_2017}, and the RStudio IDE.
+Moreover, the S4 data structures implemented by IRanges do not easily interface with the recent advances provided by the popular tidyverse suite of data processing and visualization tools (e.g., dplyr, purrr, broom, and ggplot2) \cite{wickham_tidyverse_2017}.
+We sought to develop an flexible R package for genomic interval arithmetic built to incorporate new R programming, visualization, and interactivity features.
 
@@
 Compute intensive functions in valr are written in C++ using Rcpp to enable fluid interactive analysis of large datasets\cite{eddelbuettel_rcpp_2011}.
-Interval intersections and related operations utilize an interval tree algorithm to efficently process large datasets\cite{cormen_2001}.
+Interval intersections and related operations use an interval tree algorithm to efficiently search for overlapping intervals\cite{cormen_2001}.
 BED files are imported and handled in R as data\_frame objects, requiring minimal pre or post-processing to integrate with additional R packages or command line tools.
@@
 valr is distributed as part of the CRAN R package repository and is compatible with Mac OS X, Windows, and major Linux operating systems.
-Package dependencies and system requirements are documented on the valr CRAN repository (\url{https://cran.r-project.org/web/packages/valr/index.html}).
+Package dependencies and system requirements are documented in the valr CRAN repository (\url{https://cran.r-project.org/web/packages/valr/index.html}).
 
 \section*{Use Cases}
-To demonstrate the functionality and utility of \textbf{valr}, presented below is a basic tutorial for using valr and additional common use cases for genomic interval analysis.
+To demonstrate the functionality and utility of \texttt{valr}, presented below is a basic tutorial for using valr and additional common use cases for genomic interval analysis.
 
@@
 
-\texttt{valr} provides a set of functions to read BED, BEDgraph, and VCF formats into R as convenient \texttt{tbl\_dfs()} dataframe objects.
-All tbls will have \texttt{chrom}, \texttt{start}, and \texttt{end} columns, and tbls from multi-column formats will have additional pre-determined column names.
+\texttt{valr} provides a set of functions to read BED, BEDgraph, and VCF formats into R as convenient \texttt{tibble} data\_frame objects.
+All tbls have \texttt{chrom}, \texttt{start}, and \texttt{end} columns, and tbls from multi-column formats have additional pre-determined column names.
 Standards methods for importing data (e.g.
-\texttt{read.table}, \texttt{readr::read\_tsv}) are also supported provided the contructed dataframes contain the requiste column names (\texttt{chrom}, \texttt{start}, \texttt{end}).
+\texttt{read.table}, \texttt{readr::read\_tsv}) are also supported provided the constructed dataframes contain the requisite column names (\texttt{chrom}, \texttt{start}, \texttt{end}).
 Additionally \texttt{valr} supports connections to remote databases to programatically access the UCSC and Ensembl databases via the \texttt{db\_ucsc} and \texttt{db\_ensembl} functions.
@@
 
-\subsubsection*{Simple example}
+\subsubsection*{Example of combining valr tools}
 
-The functions in \texttt{valr} have similar names to their
-\texttt{BEDtools} counterparts, and so will be familiar to users of
-the \texttt{BEDtools} suite.
+The functions in \texttt{valr} have similar names to their \texttt{BEDtools} counterparts, and so will be familiar to users of the \texttt{BEDtools} suite.
 Also, similar to pybedtools\cite{dale_pybedtools:_2011}, a python wrapper for \texttt{BEDTools}, \texttt{valr} has a terse syntax.
-For example, shown below is an example of how to find all intergenic SNPs within 1 kilobase of genes using \texttt{valr}.
+For example, shown below is a demonstration of how to find all intergenic SNPs within 1 kilobase of genes using \texttt{valr}.
 
@@
 \subsubsection*{Visual documentation}\label{visual-documentation}
-By conducting interval arthimetic entirely in R, \texttt{valr} is also an effective teaching tool for introducing interval analysis to early-stage analysts without requiring familiarity with both command-line tools and R.
+By conducting interval arithmetic entirely in R, \texttt{valr} is also an effective teaching tool for introducing interval analysis to early-stage analysts without requiring familiarity with both command-line tools and R.
 To aid in demonstrating the interval operations available in \texttt{valr}, we developed the \texttt{bed\_glyph()} tool which produces plots demonstrating the input and output of operations in \texttt{valr} in a manner similar to those found in the \texttt{BEDtools} documentation.
@@
 This demonstration illustrates how to use \texttt{valr} tools to perform
-a ``meta-analysis'' of signals relative to genomic features. Here we to
+a ``meta-analysis'' of signals relative to genomic features. Here we
 analyze the distribution of histone marks surrounding transcription
-start sites, using H3K4Me3 Chip-Seq data downloaded from ENCODE (Link to data here?).
+start sites, using H3K4Me3 Chip-Seq data downloaded from ENCODE.
 
-First we load libraries and relevant data.
+First we load packages and relevant data.
 
@@
 (TSSs). We focus on \texttt{+} strand genes, but \texttt{-} genes are
-easily accomodated by filtering them and using
-\texttt{bed\_makewindows()} with \texttt{reverse}d window numbers.
+easily accommodated by filtering them and using
+\texttt{bed\_makewindows()} with \texttt{reversed} window numbers.
 
@@
 Using \texttt{bed\_shuffle()} and \texttt{replicate()} we generate a
-\texttt{data\_frame} containing 100 sets of randomly placed intervals
-then calculate the jaccard index for each set to generate a
-null-distribution of jaccard scores. Finally an empirical p-value is
-then calculated from the null-distribution.
+\texttt{data\_frame} containing 100 sets of randomly selected intervals
+then calculate the jaccard index for each set against the repeat intervals
+to generate a null-distribution of jaccard scores. Finally, an empirical p-value is
+calculated from the null-distribution.
 
@@
 In order to ensure that valr performs fast enough to enable interactive analysis, key functionality is implemented in C++.
-To test the speed of major valr functions we generated two dataframes containing 1 million random 1 kilobase intervals derived from the human genome.
+To test the speed of major valr functions we generated two data\_frames containing 1 million random 1 kilobase intervals derived from the human genome.
 Most of the major valr functions complete execution in less than 1 second, demonstrating that valr can process large interval datasets efficiently (Figure 3A\ref{fig:Figure 3}).
@@
 \href{https://bitbucket.org/snakemake/snakemake/wiki/Home}{\texttt{snakemake}}\cite{koster_snakemakescalable_2012}),
-but it is cumbersome to transistion from command-line tools to exploratory analysis and plotting software.
+but it is cumbersome to transition from command-line tools to exploratory analysis and plotting software.
 RMarkdown documents are plain text files, amenable to version control, which provide an interface to generate feature rich PDF and HTML reports that combine text, executable code, and figures in a single document.
@@
 Additionally, \texttt{valr} seamlessly integrates into R shiny applications allowing for complex interactive visualizations relating to genomic interval analyses.
-We have developed a shiny application available on github \href{https://github.com/rnabioco/valrdata} that explores CHiP-Seq signal density surrounding transcription start sites and demonstrates the ease of implementing \texttt{valr} to power dynamic visualizations, (e.g ).
+We have developed a shiny application available on github \href{https://github.com/rnabioco/valrdata} that explores ChiP-Seq signal density surrounding transcription start sites and demonstrates the ease of implementing \texttt{valr} to power dynamic visualizations.
 
@@
 Additionally by providing an easy-to-use interface for interval arithmetic in R, valr is also a useful teaching tool to introduce the analyses necessary to investigate correlations between genomic intervals, without requiring familiarity with the command-line.
-We envison that \texttt{valr} will help researchers quickly and reproducibly analyze genome interval datasets.
+We envision that \texttt{valr} will help researchers quickly and reproducibly analyze genome interval datasets.
 
@@
 \item valr source code is available at http://github.com/rnabioco/valr
-\item Latest stable version of source code is at: https://github.com/rnabioco/valr/archive/v0.2.0.tar.gz TO UPDATE
+\item Latest stable version of source code is at: https://github.com/rnabioco/valr/archive/v0.3.0.tar.gz
 \item Software license: MIT
@@
 
-JR conceptualized the package and provided direction for software development.
+JR conceptualized the package and directed software development.
 JR, KR, AG, and RS developed and tested the package.
@@
 \caption{\label{fig:Figure 2}Meta-analysis of signals relative to genomic features with valr.
-\textnormal{(A) Summarized coverage of human H3K4Me3 Chip-Seq coverage across positive strand transcription start sites.
+\textnormal{(A) Summarized coverage of human H3K4Me3 Chip-Seq coverage across positive strand transcription start sites on chromosome 22.
 Data presented +/- SEM.}}
@@
 \caption{\label{fig:Figure 3} Performance of major valr functionality.
-\textnormal{(A)  Timings calculated by performing 10 repetitions of indicated functions on dataframes containing 1 million random 1 kilobase x/y intervals generated using \texttt{bed\_random()}.
-(B) Timings for executing functions in BEDTools or equivalent functions in valr using the same interval sets as in (A).
-All BEDTools functions were written to /dev/null, and were timed using GNU \texttt{time}.
-Timings for valr functions in (B) include times for reading files using \texttt{read\_bed()} and were timed using the \texttt{microbenchmark} package.}}
-% * <jay.hesselberth@gmail.com> 2017-06-07T13:17:50.374Z:
-%
-% No bar charts! Make this a boxplot ideally.
-%
-% ^.
+\textnormal{(A)  Timings calculated by performing 10 repetitions of indicated functions on dataframes preloaded in R containing 1 million random 1 kilobase x/y intervals generated using \texttt{bed\_random()}.
+(B) Timings for executing functions in BEDTools or equivalent functions in valr using the same interval sets as in (A) written to files.
+All BEDTools function outputs were written to /dev/null/, and were timed using GNU \texttt{time}.
+Timings for valr functions in (B) include times for reading files using \texttt{read\_bed()} functions and were timed using the \texttt{microbenchmark} package.}}
 \end{figure}

### 2017-06-19 inst/paper/figs.Rmd
grey scale colors and add lines per expression

@@
 date: "June 6, 2017"
-output: pdf_document
 ---
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
-knitr::opts_chunk$set(cache = TRUE)
+knitr::opts_chunk$set( 
+  collapse = TRUE,
+  comment = "#>",
+  cache = TRUE,
+  echo = TRUE)
 ```
@@
 plot_a <- ggplot(res, aes(x=reorder(expr, time), y = time)) +
-  geom_boxplot(width = 0.75, size = 0.33, outlier.shape = NA, fill = RColorBrewer::brewer.pal(3, "Set1")[2]) +
+  geom_boxplot(width = 0.75, size = 0.33, outlier.shape = NA, fill = grDevices::grey.colors(2)[2]) +
   coord_flip() +
@@
   geom_boxplot(width = 0.75, size = 0.33, outlier.shape = NA) +
-  scale_fill_brewer(palette = "Set1") + 
+  scale_fill_grey() + 
   coord_flip() +
+  geom_vline(xintercept = seq_along(unique(dat$func)) + 0.5,
+             color = "grey") +
   theme_bw() +
@@
 
-Here we examine the overlap of repeat classes in the human genome (on `chr22` only, for simplicity) using `bed_jaccard()`.
+Here we examine the extent of overlap of repeat classes with exons in the human genome (on `chr22` only, for simplicity) using the jaccard similarity index.
+`bed_jaccard()` implements the jaccard test to examine the similarity between two sets of genomic intervals.
+Using `bed_shuffle()` and `replicate()` we generate a `data_frame` containing 100 sets of randomly placed intervals then calculate the jaccard index for each set to generate a null-distribution of jaccard scores.
+Finally an empirical p-value is then calculated from the null-distribution.
 
 ```{r reldist_shuffle, message = FALSE, warning = FALSE}
+library(tidyverse, warn.conflicts = F)
 
@@
 genes <- read_bed12(valr_example('hg19.refGene.chr22.bed.gz'))
+# convert bed12 to bed with exons
 exons <- bed12_to_exons(genes)
 
+# function to repeat interval shuffling
 shuffle_intervals <- function(n, .data, genome) {
@@
 }
-
-shuffled <- shuffle_intervals(n = 100, repeats, genome) %>%
+nreps <- 100
+shuffled <- shuffle_intervals(n = nreps, repeats, genome) %>%
   mutate(jaccard = data %>%
@@
            map_dbl("jaccard"))
-
 shuffled
 
-bed_jaccard(repeats, exons)
+obs <- bed_jaccard(repeats, exons)
+obs
+
+pvalue <- sum(shuffled$jaccard >= obs$jaccard) + 1 /(nreps + 1)
+pvalue
 ```

### 2017-06-20 inst/paper/figs.Rmd
fixed quotes and a few minor changes

@@
 ```{r fig2}
-# `valr_example()` identifies the path of example files
-bedfile <- valr_example('genes.hg19.chr22.bed.gz')
-genomefile <- valr_example('hg19.chrom.sizes.gz')
-bgfile  <- valr_example('hela.h3k4.chip.bg.gz')
+bedfile <- valr_example("genes.hg19.chr22.bed.gz")
+genomefile <- valr_example("hg19.chrom.sizes.gz")
+bgfile  <- valr_example("hela.h3k4.chip.bg.gz")
 
@@
   scale_x_continuous(labels = x_labels, breaks = x_breaks) + 
-  xlab('Position (bp from TSS)') + ylab('Signal') + 
+  xlab("Position (bp from TSS)") + ylab("Signal") + 
   theme_classic() +
-  pub_theme
+  pub_theme +
+  theme(axis.text.x = element_text(angle = 90))
 
-plot_grid(p, align = 'h', nrow = 1, labels="AUTO")
-ggsave("figure2.pdf", width = 5.6, height = 4)
+plot_grid(p, align = "h", nrow = 1, labels="AUTO")
+ggsave("figure2.pdf", width = 5.6, height = 5.6)
 ```
@@
 # Set-up test genome
-genome <- read_genome(valr_example('hg19.chrom.sizes.gz'))
+genome <- read_genome(valr_example("hg19.chrom.sizes.gz"))
 write_tsv(genome, "genome.txt", col_names=FALSE)
@@
 
+### file I/O
 ```{r}
@@
 
-Using mysql
-```{r, warnings = F}
-library(dplyr, warn.conflicts = F)
-hg19 <- db_ucsc("hg19")
-tbl(hg19, "refGene")
+### syntax demo
+
+```{r syntax_demo, message = FALSE}
+library(valr)
+library(dplyr)
+
+snps <- read_bed(valr_example("hg19.snps147.chr22.bed.gz"), n_fields = 6)
+genes <- read_bed(valr_example("genes.hg19.chr22.bed.gz"), n_fields = 6)
+
+# find snps in intergenic regions
+intergenic <- bed_subtract(snps, genes)
+# distance from intergenic snps to nearest gene
+nearby <- bed_closest(intergenic, genes)
+
+nearby %>%
+  select(starts_with("name"), .overlap, .dist) %>%
+  filter(abs(.dist) < 1000)
+```
+
+### Figure 1 code demo
+
+```{r}
+x <- tibble::tribble(
+  ~chrom, ~start, ~end,
+  "chr1", 25,     50,
+  "chr1", 100,    125
+)
+
+y <- tibble::tribble(
+  ~chrom, ~start, ~end,
+  "chr1", 30,     75
+)
+
+bed_glyph(bed_intersect(x, y))
+```
+
+```{r}
+x <- tibble::tribble(
+  ~chrom, ~start, ~end,
+  "chr1",      1,      50,
+  "chr1",      10,     75,
+  "chr1",      100,    120
+)
+
+bed_glyph(bed_merge(x))
+```
+
+### grouping data demo
+
+```{r}
+x <- tibble::tribble(
+  ~chrom, ~start, ~end, ~strand,
+  "chr1", 1,      100,  "+",
+  "chr1", 50,     150,  "+",
+  "chr2", 100,    200,  "-"
+)
+
+y <- tibble::tribble(
+  ~chrom, ~start, ~end, ~strand,
+  "chr1", 50,     125,  "+",
+  "chr1", 50,     150,  "-",
+  "chr2", 50,     150,  "+"
+)
+
+# intersect tbls by strand
+x <- group_by(x, strand)
+y <- group_by(y, strand)
+
+bed_intersect(x, y)
+```
+
+Comparisons between intervals on opposite strands are done using the `flip_strands()` function:
+
+```{r}
+x <- group_by(x, strand)
+
+y <- flip_strands(y)
+y <- group_by(y, strand)
+
+bed_intersect(x, y)
+```
+
+### Figure 2 code demo
+```{r }
+bedfile <- valr_example("genes.hg19.chr22.bed.gz")
+genomefile <- valr_example("hg19.chrom.sizes.gz")
+bgfile  <- valr_example("hela.h3k4.chip.bg.gz")
+
+genes <- read_bed(bedfile, n_fields = 6)
+genome <- read_genome(genomefile)
+y <- read_bedgraph(bgfile)
+```
+
+Then we generate 1 bp intervals to represent transcription start sites (TSSs).
+We focus on `+` strand genes, but `-` genes are easily accomodated by filtering them and using `bed_makewindows()` with `reversed` window numbers.
+
+```{r}
+# generate 1 bp TSS intervals, "+" strand only
+tss <- genes %>%
+  filter(strand == "+") %>%
+  mutate(end = start + 1)
+
+# 1000 bp up and downstream
+region_size <- 1000
+# 50 bp windows
+win_size <- 50
+
+# add slop to the TSS, break into windows and add a group
+x <- tss %>%
+  bed_slop(genome, both = region_size) %>%
+  bed_makewindows(genome, win_size)
+
+x
+```
+
+Now we use the `.win_id` group with `bed_map()` to caluclate a sum by mapping `y` signals onto the intervals in `x`.
+These data are regrouped by `.win_id` and a summary with `mean` and `sd` values is calculated.
+
+```{r}
+# map signals to TSS regions and calculate summary statistics.
+res <- bed_map(x, y, win_sum = sum(value, na.rm = TRUE)) %>%
+  group_by(.win_id) %>%
+  summarize(win_mean = mean(win_sum, na.rm = TRUE),
+            win_sd = sd(win_sum, na.rm = TRUE))
+
+res
+```
+
+Finally, these summary statistics are used to construct a plot that illustrates histone density surrounding TSSs.
+
+```{r}
+library(ggplot2)
+
+x_labels <- seq(-region_size, region_size, by = win_size * 5)
+x_breaks <- seq(1, 41, by = 5)
+
+sd_limits <- aes(ymax = win_mean + win_sd, ymin = win_mean - win_sd)
+
+p <- ggplot(res, aes(x = .win_id, y = win_mean)) +
+  geom_point(size = 0.25) + geom_pointrange(sd_limits, size = 0.1) + 
+  scale_x_continuous(labels = x_labels, breaks = x_breaks) + 
+  xlab("Position (bp from TSS)") + ylab("Signal") + 
+  theme_classic()
 ```
@@
 
-repeats <- read_bed(valr_example('hg19.rmsk.chr22.bed.gz'), n_fields = 6) 
-genome <- read_genome(valr_example('hg19.chrom.sizes.gz'))
-genes <- read_bed12(valr_example('hg19.refGene.chr22.bed.gz'))
+repeats <- read_bed(valr_example("hg19.rmsk.chr22.bed.gz"), n_fields = 6) 
+genome <- read_genome(valr_example("hg19.chrom.sizes.gz"))
+genes <- read_bed12(valr_example("hg19.refGene.chr22.bed.gz"))
 # convert bed12 to bed with exons
@@
   replicate(n, bed_shuffle(.data, genome), simplify = FALSE) %>%
-    bind_rows(.id = 'rep') %>%
+    bind_rows(.id = "rep") %>%
     group_by(rep) %>% nest()

### 2017-06-20 inst/paper/figs.tex
fixed quotes and a few minor changes

@@
 \KeywordTok{library}\NormalTok{(tidyverse)}
+\CommentTok{#> Loading tidyverse: ggplot2}
+\CommentTok{#> Loading tidyverse: tibble}
+\CommentTok{#> Loading tidyverse: tidyr}
+\CommentTok{#> Loading tidyverse: readr}
+\CommentTok{#> Loading tidyverse: purrr}
+\CommentTok{#> Loading tidyverse: dplyr}
+\CommentTok{#> Conflicts with tidy packages ----------------------------------------------}
+\CommentTok{#> filter(): dplyr, stats}
+\CommentTok{#> lag():    dplyr, stats}
 \KeywordTok{library}\NormalTok{(cowplot)}
+\CommentTok{#> }
+\CommentTok{#> Attaching package: 'cowplot'}
+\CommentTok{#> The following object is masked from 'package:ggplot2':}
+\CommentTok{#> }
+\CommentTok{#>     ggsave}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\CommentTok{# `valr_example()` identifies the path of example files}
-\NormalTok{bedfile <-}\StringTok{ }\KeywordTok{valr_example}\NormalTok{(}\StringTok{'genes.hg19.chr22.bed.gz'}\NormalTok{)}
-\NormalTok{genomefile <-}\StringTok{ }\KeywordTok{valr_example}\NormalTok{(}\StringTok{'hg19.chrom.sizes.gz'}\NormalTok{)}
-\NormalTok{bgfile  <-}\StringTok{ }\KeywordTok{valr_example}\NormalTok{(}\StringTok{'hela.h3k4.chip.bg.gz'}\NormalTok{)}
+\NormalTok{bedfile <-}\StringTok{ }\KeywordTok{valr_example}\NormalTok{(}\StringTok{"genes.hg19.chr22.bed.gz"}\NormalTok{)}
+\NormalTok{genomefile <-}\StringTok{ }\KeywordTok{valr_example}\NormalTok{(}\StringTok{"hg19.chrom.sizes.gz"}\NormalTok{)}
+\NormalTok{bgfile  <-}\StringTok{ }\KeywordTok{valr_example}\NormalTok{(}\StringTok{"hela.h3k4.chip.bg.gz"}\NormalTok{)}
 
@@
 \StringTok{  }\KeywordTok{scale_x_continuous}\NormalTok{(}\DataTypeTok{labels =} \NormalTok{x_labels, }\DataTypeTok{breaks =} \NormalTok{x_breaks) +}\StringTok{ }
-\StringTok{  }\KeywordTok{xlab}\NormalTok{(}\StringTok{'Position (bp from TSS)'}\NormalTok{) +}\StringTok{ }\KeywordTok{ylab}\NormalTok{(}\StringTok{'Signal'}\NormalTok{) +}\StringTok{ }
+\StringTok{  }\KeywordTok{xlab}\NormalTok{(}\StringTok{"Position (bp from TSS)"}\NormalTok{) +}\StringTok{ }\KeywordTok{ylab}\NormalTok{(}\StringTok{"Signal"}\NormalTok{) +}\StringTok{ }
 \StringTok{  }\KeywordTok{theme_classic}\NormalTok{() +}
-\StringTok{  }\NormalTok{pub_theme}
+\StringTok{  }\NormalTok{pub_theme +}
+\StringTok{  }\KeywordTok{theme}\NormalTok{(}\DataTypeTok{axis.text.x =} \KeywordTok{element_text}\NormalTok{(}\DataTypeTok{angle =} \DecValTok{90}\NormalTok{))}
 
-\KeywordTok{plot_grid}\NormalTok{(p, }\DataTypeTok{align =} \StringTok{'h'}\NormalTok{, }\DataTypeTok{nrow =} \DecValTok{1}\NormalTok{, }\DataTypeTok{labels=}\StringTok{"AUTO"}\NormalTok{)}
+\KeywordTok{plot_grid}\NormalTok{(p, }\DataTypeTok{align =} \StringTok{"h"}\NormalTok{, }\DataTypeTok{nrow =} \DecValTok{1}\NormalTok{, }\DataTypeTok{labels=}\StringTok{"AUTO"}\NormalTok{)}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\KeywordTok{ggsave}\NormalTok{(}\StringTok{"figure2.pdf"}\NormalTok{, }\DataTypeTok{width =} \FloatTok{5.6}\NormalTok{, }\DataTypeTok{height =} \DecValTok{4}\NormalTok{)}
+\KeywordTok{ggsave}\NormalTok{(}\StringTok{"figure2.pdf"}\NormalTok{, }\DataTypeTok{width =} \FloatTok{5.6}\NormalTok{, }\DataTypeTok{height =} \FloatTok{5.6}\NormalTok{)}
 \end{Highlighting}
@@
 \CommentTok{# Set-up test genome}
-\NormalTok{genome <-}\StringTok{ }\KeywordTok{read_genome}\NormalTok{(}\KeywordTok{valr_example}\NormalTok{(}\StringTok{'hg19.chrom.sizes.gz'}\NormalTok{))}
+\NormalTok{genome <-}\StringTok{ }\KeywordTok{read_genome}\NormalTok{(}\KeywordTok{valr_example}\NormalTok{(}\StringTok{"hg19.chrom.sizes.gz"}\NormalTok{))}
 \KeywordTok{write_tsv}\NormalTok{(genome, }\StringTok{"genome.txt"}\NormalTok{, }\DataTypeTok{col_names=}\OtherTok{FALSE}\NormalTok{)}
@@
 \NormalTok{plot_a <-}\StringTok{ }\KeywordTok{ggplot}\NormalTok{(res, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x=}\KeywordTok{reorder}\NormalTok{(expr, time), }\DataTypeTok{y =} \NormalTok{time)) +}
-\StringTok{  }\KeywordTok{geom_boxplot}\NormalTok{(}\DataTypeTok{width =} \FloatTok{0.75}\NormalTok{, }\DataTypeTok{size =} \FloatTok{0.33}\NormalTok{, }\DataTypeTok{outlier.shape =} \OtherTok{NA}\NormalTok{, }\DataTypeTok{fill =} \NormalTok{RColorBrewer::}\KeywordTok{brewer.pal}\NormalTok{(}\DecValTok{3}\NormalTok{, }\StringTok{"Set1"}\NormalTok{)[}\DecValTok{2}\NormalTok{]) +}
+\StringTok{  }\KeywordTok{geom_boxplot}\NormalTok{(}\DataTypeTok{width =} \FloatTok{0.75}\NormalTok{, }\DataTypeTok{size =} \FloatTok{0.33}\NormalTok{, }\DataTypeTok{outlier.shape =} \OtherTok{NA}\NormalTok{, }\DataTypeTok{fill =} \NormalTok{grDevices::}\KeywordTok{grey.colors}\NormalTok{(}\DecValTok{2}\NormalTok{)[}\DecValTok{2}\NormalTok{]) +}
 \StringTok{  }\KeywordTok{coord_flip}\NormalTok{() +}
@@
 \StringTok{  }\KeywordTok{geom_boxplot}\NormalTok{(}\DataTypeTok{width =} \FloatTok{0.75}\NormalTok{, }\DataTypeTok{size =} \FloatTok{0.33}\NormalTok{, }\DataTypeTok{outlier.shape =} \OtherTok{NA}\NormalTok{) +}
-\StringTok{  }\KeywordTok{scale_fill_brewer}\NormalTok{(}\DataTypeTok{palette =} \StringTok{"Set1"}\NormalTok{) +}\StringTok{ }
+\StringTok{  }\KeywordTok{scale_fill_grey}\NormalTok{() +}\StringTok{ }
 \StringTok{  }\KeywordTok{coord_flip}\NormalTok{() +}
+\StringTok{  }\KeywordTok{geom_vline}\NormalTok{(}\DataTypeTok{xintercept =} \KeywordTok{seq_along}\NormalTok{(}\KeywordTok{unique}\NormalTok{(dat$func)) +}\StringTok{ }\FloatTok{0.5}\NormalTok{,}
+             \DataTypeTok{color =} \StringTok{"grey"}\NormalTok{) +}
 \StringTok{  }\KeywordTok{theme_bw}\NormalTok{() +}
@@
 \subsection{Code demonstrations}\label{code-demonstrations}
+
+\subsubsection{file I/O}\label{file-io}
 
@@
 
-Using mysql
+\subsubsection{syntax demo}\label{syntax-demo}
 
@@
 \begin{Highlighting}[]
-\KeywordTok{library}\NormalTok{(dplyr, }\DataTypeTok{warn.conflicts =} \NormalTok{F)}
-\NormalTok{hg19 <-}\StringTok{ }\KeywordTok{db_ucsc}\NormalTok{(}\StringTok{"hg19"}\NormalTok{)}
-\KeywordTok{tbl}\NormalTok{(hg19, }\StringTok{"refGene"}\NormalTok{)}
-\CommentTok{#> Warning in .local(conn, statement, ...): Unsigned INTEGER in col 0 imported}
-\CommentTok{#> as numeric}
-\CommentTok{#> Warning in .local(conn, statement, ...): Unsigned INTEGER in col 4 imported}
-\CommentTok{#> as numeric}
-\CommentTok{#> Warning in .local(conn, statement, ...): Unsigned INTEGER in col 5 imported}
-\CommentTok{#> as numeric}
-\CommentTok{#> Warning in .local(conn, statement, ...): Unsigned INTEGER in col 6 imported}
-\CommentTok{#> as numeric}
-\CommentTok{#> Warning in .local(conn, statement, ...): Unsigned INTEGER in col 7 imported}
-\CommentTok{#> as numeric}
-\CommentTok{#> Warning in .local(conn, statement, ...): Unsigned INTEGER in col 8 imported}
-\CommentTok{#> as numeric}
-\CommentTok{#> Warning in .local(conn, statement, ...): Unsigned INTEGER in col 0 imported}
-\CommentTok{#> as numeric}
-\CommentTok{#> Warning in .local(conn, statement, ...): Unsigned INTEGER in col 4 imported}
-\CommentTok{#> as numeric}
-\CommentTok{#> Warning in .local(conn, statement, ...): Unsigned INTEGER in col 5 imported}
-\CommentTok{#> as numeric}
-\CommentTok{#> Warning in .local(conn, statement, ...): Unsigned INTEGER in col 6 imported}
-\CommentTok{#> as numeric}
-\CommentTok{#> Warning in .local(conn, statement, ...): Unsigned INTEGER in col 7 imported}
-\CommentTok{#> as numeric}
-\CommentTok{#> Warning in .local(conn, statement, ...): Unsigned INTEGER in col 8 imported}
-\CommentTok{#> as numeric}
-\CommentTok{#> # Source:   table<refGene> [?? x 16]}
-\CommentTok{#> # Database: mysql 5.6.26-log [genomep@genome-mysql.cse.ucsc.edu:/hg19]}
-\CommentTok{#>      bin         name chrom strand   txStart     txEnd  cdsStart    cdsEnd}
-\CommentTok{#>    <dbl>        <chr> <chr>  <chr>     <dbl>     <dbl>     <dbl>     <dbl>}
-\CommentTok{#>  1   585    NR_024540  chr1      -     14361     29370     29370     29370}
-\CommentTok{#>  2  1759    NR_110022  chrX      + 153991016 154005964 154005964 154005964}
-\CommentTok{#>  3   604    NR_033380  chrX      +   2527305   2575270   2575270   2575270}
-\CommentTok{#>  4    75    NR_033381  chrY      +   2477305   2506370   2506370   2506370}
-\CommentTok{#>  5   915    NM_182508 chr13      +  43355685  43365685  43358203  43362926}
-\CommentTok{#>  6   923    NM_181845 chr19      +  44331472  44353050  44337656  44352793}
-\CommentTok{#>  7   586    NR_039983  chr1      -    134772    140566    140566    140566}
-\CommentTok{#>  8  1964    NR_028327  chr5      + 180750506 180755196 180755196 180755196}
-\CommentTok{#>  9   126 NM_001271875 chr17      -  55938603  56032684  55943837  55962925}
-\CommentTok{#> 10  2278    NR_125989  chr1      - 222001007 222014008 222014008 222014008}
-\CommentTok{#> # ... with more rows, and 8 more variables: exonCount <dbl>,}
-\CommentTok{#> #   exonStarts <chr>, exonEnds <chr>, score <int>, name2 <chr>,}
-\CommentTok{#> #   cdsStartStat <chr>, cdsEndStat <chr>, exonFrames <chr>}
+\KeywordTok{library}\NormalTok{(valr)}
+\KeywordTok{library}\NormalTok{(dplyr)}
+
+\NormalTok{snps <-}\StringTok{ }\KeywordTok{read_bed}\NormalTok{(}\KeywordTok{valr_example}\NormalTok{(}\StringTok{"hg19.snps147.chr22.bed.gz"}\NormalTok{), }\DataTypeTok{n_fields =} \DecValTok{6}\NormalTok{)}
+\NormalTok{genes <-}\StringTok{ }\KeywordTok{read_bed}\NormalTok{(}\KeywordTok{valr_example}\NormalTok{(}\StringTok{"genes.hg19.chr22.bed.gz"}\NormalTok{), }\DataTypeTok{n_fields =} \DecValTok{6}\NormalTok{)}
+
+\CommentTok{# find snps in intergenic regions}
+\NormalTok{intergenic <-}\StringTok{ }\KeywordTok{bed_subtract}\NormalTok{(snps, genes)}
+\CommentTok{# distance from intergenic snps to nearest gene}
+\NormalTok{nearby <-}\StringTok{ }\KeywordTok{bed_closest}\NormalTok{(intergenic, genes)}
+
+\NormalTok{nearby %>%}
+\StringTok{  }\KeywordTok{select}\NormalTok{(}\KeywordTok{starts_with}\NormalTok{(}\StringTok{"name"}\NormalTok{), .overlap, .dist) %>%}
+\StringTok{  }\KeywordTok{filter}\NormalTok{(}\KeywordTok{abs}\NormalTok{(.dist) <}\StringTok{ }\DecValTok{1000}\NormalTok{)}
+\CommentTok{#> # A tibble: 285 x 4}
+\CommentTok{#>         name.x            name.y .overlap .dist}
+\CommentTok{#>          <chr>             <chr>    <int> <int>}
+\CommentTok{#>  1   rs2261631             P704P        0  -267}
+\CommentTok{#>  2 rs570770556             POTEH        0  -912}
+\CommentTok{#>  3 rs538163832             POTEH        0  -952}
+\CommentTok{#>  4   rs9606135            TPTEP1        0  -421}
+\CommentTok{#>  5  rs11912392 ANKRD62P1-PARP4P3        0   104}
+\CommentTok{#>  6   rs8136454          BC038197        0   355}
+\CommentTok{#>  7   rs5992556              XKR3        0  -455}
+\CommentTok{#>  8 rs114101676              GAB4        0   473}
+\CommentTok{#>  9  rs62236167             CECR7        0   261}
+\CommentTok{#> 10   rs5747023             CECR1        0  -386}
+\CommentTok{#> # ... with 275 more rows}
+\end{Highlighting}
+\end{Shaded}
+
+\subsubsection{Figure 1 code demo}\label{figure-1-code-demo}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{x <-}\StringTok{ }\NormalTok{tibble::}\KeywordTok{tribble}\NormalTok{(}
+  \NormalTok{~chrom, ~start, ~end,}
+  \StringTok{"chr1"}\NormalTok{, }\DecValTok{25}\NormalTok{,     }\DecValTok{50}\NormalTok{,}
+  \StringTok{"chr1"}\NormalTok{, }\DecValTok{100}\NormalTok{,    }\DecValTok{125}
+\NormalTok{)}
+
+\NormalTok{y <-}\StringTok{ }\NormalTok{tibble::}\KeywordTok{tribble}\NormalTok{(}
+  \NormalTok{~chrom, ~start, ~end,}
+  \StringTok{"chr1"}\NormalTok{, }\DecValTok{30}\NormalTok{,     }\DecValTok{75}
+\NormalTok{)}
+
+\KeywordTok{bed_glyph}\NormalTok{(}\KeywordTok{bed_intersect}\NormalTok{(x, y))}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{x <-}\StringTok{ }\NormalTok{tibble::}\KeywordTok{tribble}\NormalTok{(}
+  \NormalTok{~chrom, ~start, ~end,}
+  \StringTok{"chr1"}\NormalTok{,      }\DecValTok{1}\NormalTok{,      }\DecValTok{50}\NormalTok{,}
+  \StringTok{"chr1"}\NormalTok{,      }\DecValTok{10}\NormalTok{,     }\DecValTok{75}\NormalTok{,}
+  \StringTok{"chr1"}\NormalTok{,      }\DecValTok{100}\NormalTok{,    }\DecValTok{120}
+\NormalTok{)}
+
+\KeywordTok{bed_glyph}\NormalTok{(}\KeywordTok{bed_merge}\NormalTok{(x))}
+\end{Highlighting}
+\end{Shaded}
+
+\subsubsection{grouping data demo}\label{grouping-data-demo}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{x <-}\StringTok{ }\NormalTok{tibble::}\KeywordTok{tribble}\NormalTok{(}
+  \NormalTok{~chrom, ~start, ~end, ~strand,}
+  \StringTok{"chr1"}\NormalTok{, }\DecValTok{1}\NormalTok{,      }\DecValTok{100}\NormalTok{,  }\StringTok{"+"}\NormalTok{,}
+  \StringTok{"chr1"}\NormalTok{, }\DecValTok{50}\NormalTok{,     }\DecValTok{150}\NormalTok{,  }\StringTok{"+"}\NormalTok{,}
+  \StringTok{"chr2"}\NormalTok{, }\DecValTok{100}\NormalTok{,    }\DecValTok{200}\NormalTok{,  }\StringTok{"-"}
+\NormalTok{)}
+
+\NormalTok{y <-}\StringTok{ }\NormalTok{tibble::}\KeywordTok{tribble}\NormalTok{(}
+  \NormalTok{~chrom, ~start, ~end, ~strand,}
+  \StringTok{"chr1"}\NormalTok{, }\DecValTok{50}\NormalTok{,     }\DecValTok{125}\NormalTok{,  }\StringTok{"+"}\NormalTok{,}
+  \StringTok{"chr1"}\NormalTok{, }\DecValTok{50}\NormalTok{,     }\DecValTok{150}\NormalTok{,  }\StringTok{"-"}\NormalTok{,}
+  \StringTok{"chr2"}\NormalTok{, }\DecValTok{50}\NormalTok{,     }\DecValTok{150}\NormalTok{,  }\StringTok{"+"}
+\NormalTok{)}
+
+\CommentTok{# intersect tbls by strand}
+\NormalTok{x <-}\StringTok{ }\KeywordTok{group_by}\NormalTok{(x, strand)}
+\NormalTok{y <-}\StringTok{ }\KeywordTok{group_by}\NormalTok{(y, strand)}
+
+\KeywordTok{bed_intersect}\NormalTok{(x, y)}
+\CommentTok{#> # A tibble: 2 x 8}
+\CommentTok{#>   chrom start.x end.x strand.x start.y end.y strand.y .overlap}
+\CommentTok{#>   <chr>   <dbl> <dbl>    <chr>   <dbl> <dbl>    <chr>    <int>}
+\CommentTok{#> 1  chr1       1   100        +      50   125        +       50}
+\CommentTok{#> 2  chr1      50   150        +      50   125        +       75}
+\end{Highlighting}
+\end{Shaded}
+
+Comparisons between intervals on opposite strands are done using the
+\texttt{flip\_strands()} function:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{x <-}\StringTok{ }\KeywordTok{group_by}\NormalTok{(x, strand)}
+
+\NormalTok{y <-}\StringTok{ }\KeywordTok{flip_strands}\NormalTok{(y)}
+\NormalTok{y <-}\StringTok{ }\KeywordTok{group_by}\NormalTok{(y, strand)}
+
+\KeywordTok{bed_intersect}\NormalTok{(x, y)}
+\CommentTok{#> # A tibble: 3 x 8}
+\CommentTok{#>   chrom start.x end.x strand.x start.y end.y strand.y .overlap}
+\CommentTok{#>   <chr>   <dbl> <dbl>    <chr>   <dbl> <dbl>    <chr>    <int>}
+\CommentTok{#> 1  chr2     100   200        -      50   150        -       50}
+\CommentTok{#> 2  chr1       1   100        +      50   150        +       50}
+\CommentTok{#> 3  chr1      50   150        +      50   150        +      100}
+\end{Highlighting}
+\end{Shaded}
+
+\subsubsection{Figure 2 code demo}\label{figure-2-code-demo}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{bedfile <-}\StringTok{ }\KeywordTok{valr_example}\NormalTok{(}\StringTok{"genes.hg19.chr22.bed.gz"}\NormalTok{)}
+\NormalTok{genomefile <-}\StringTok{ }\KeywordTok{valr_example}\NormalTok{(}\StringTok{"hg19.chrom.sizes.gz"}\NormalTok{)}
+\NormalTok{bgfile  <-}\StringTok{ }\KeywordTok{valr_example}\NormalTok{(}\StringTok{"hela.h3k4.chip.bg.gz"}\NormalTok{)}
+
+\NormalTok{genes <-}\StringTok{ }\KeywordTok{read_bed}\NormalTok{(bedfile, }\DataTypeTok{n_fields =} \DecValTok{6}\NormalTok{)}
+\NormalTok{genome <-}\StringTok{ }\KeywordTok{read_genome}\NormalTok{(genomefile)}
+\NormalTok{y <-}\StringTok{ }\KeywordTok{read_bedgraph}\NormalTok{(bgfile)}
+\end{Highlighting}
+\end{Shaded}
+
+Then we generate 1 bp intervals to represent transcription start sites
+(TSSs). We focus on \texttt{+} strand genes, but \texttt{-} genes are
+easily accomodated by filtering them and using
+\texttt{bed\_makewindows()} with \texttt{reversed} window numbers.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{# generate 1 bp TSS intervals, "+" strand only}
+\NormalTok{tss <-}\StringTok{ }\NormalTok{genes %>%}
+\StringTok{  }\KeywordTok{filter}\NormalTok{(strand ==}\StringTok{ "+"}\NormalTok{) %>%}
+\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{end =} \NormalTok{start +}\StringTok{ }\DecValTok{1}\NormalTok{)}
+
+\CommentTok{# 1000 bp up and downstream}
+\NormalTok{region_size <-}\StringTok{ }\DecValTok{1000}
+\CommentTok{# 50 bp windows}
+\NormalTok{win_size <-}\StringTok{ }\DecValTok{50}
+
+\CommentTok{# add slop to the TSS, break into windows and add a group}
+\NormalTok{x <-}\StringTok{ }\NormalTok{tss %>%}
+\StringTok{  }\KeywordTok{bed_slop}\NormalTok{(genome, }\DataTypeTok{both =} \NormalTok{region_size) %>%}
+\StringTok{  }\KeywordTok{bed_makewindows}\NormalTok{(genome, win_size)}
+
+\NormalTok{x}
+\CommentTok{#> # A tibble: 13,530 x 7}
+\CommentTok{#>    chrom    start      end      name score strand .win_id}
+\CommentTok{#>    <chr>    <int>    <int>     <chr> <chr>  <chr>   <int>}
+\CommentTok{#>  1 chr22 16161065 16161115 LINC00516     3      +       1}
+\CommentTok{#>  2 chr22 16161115 16161165 LINC00516     3      +       2}
+\CommentTok{#>  3 chr22 16161165 16161215 LINC00516     3      +       3}
+\CommentTok{#>  4 chr22 16161215 16161265 LINC00516     3      +       4}
+\CommentTok{#>  5 chr22 16161265 16161315 LINC00516     3      +       5}
+\CommentTok{#>  6 chr22 16161315 16161365 LINC00516     3      +       6}
+\CommentTok{#>  7 chr22 16161365 16161415 LINC00516     3      +       7}
+\CommentTok{#>  8 chr22 16161415 16161465 LINC00516     3      +       8}
+\CommentTok{#>  9 chr22 16161465 16161515 LINC00516     3      +       9}
+\CommentTok{#> 10 chr22 16161515 16161565 LINC00516     3      +      10}
+\CommentTok{#> # ... with 13,520 more rows}
+\end{Highlighting}
+\end{Shaded}
+
+Now we use the \texttt{.win\_id} group with \texttt{bed\_map()} to
+caluclate a sum by mapping \texttt{y} signals onto the intervals in
+\texttt{x}. These data are regrouped by \texttt{.win\_id} and a summary
+with \texttt{mean} and \texttt{sd} values is calculated.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{# map signals to TSS regions and calculate summary statistics.}
+\NormalTok{res <-}\StringTok{ }\KeywordTok{bed_map}\NormalTok{(x, y, }\DataTypeTok{win_sum =} \KeywordTok{sum}\NormalTok{(value, }\DataTypeTok{na.rm =} \OtherTok{TRUE}\NormalTok{)) %>%}
+\StringTok{  }\KeywordTok{group_by}\NormalTok{(.win_id) %>%}
+\StringTok{  }\KeywordTok{summarize}\NormalTok{(}\DataTypeTok{win_mean =} \KeywordTok{mean}\NormalTok{(win_sum, }\DataTypeTok{na.rm =} \OtherTok{TRUE}\NormalTok{),}
+            \DataTypeTok{win_sd =} \KeywordTok{sd}\NormalTok{(win_sum, }\DataTypeTok{na.rm =} \OtherTok{TRUE}\NormalTok{))}
+
+\NormalTok{res}
+\CommentTok{#> # A tibble: 41 x 3}
+\CommentTok{#>    .win_id win_mean    win_sd}
+\CommentTok{#>      <int>    <dbl>     <dbl>}
+\CommentTok{#>  1       1 100.8974  85.83423}
+\CommentTok{#>  2       2 110.6829  81.13521}
+\CommentTok{#>  3       3 122.9070  99.09635}
+\CommentTok{#>  4       4 116.2800  96.30098}
+\CommentTok{#>  5       5 116.3500 102.33773}
+\CommentTok{#>  6       6 124.9048  95.08887}
+\CommentTok{#>  7       7 122.9437  94.39792}
+\CommentTok{#>  8       8 127.5946  91.47407}
+\CommentTok{#>  9       9 130.2051  95.71309}
+\CommentTok{#> 10      10 130.1220  88.82809}
+\CommentTok{#> # ... with 31 more rows}
+\end{Highlighting}
+\end{Shaded}
+
+Finally, these summary statistics are used to construct a plot that
+illustrates histone density surrounding TSSs.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{library}\NormalTok{(ggplot2)}
+
+\NormalTok{x_labels <-}\StringTok{ }\KeywordTok{seq}\NormalTok{(-region_size, region_size, }\DataTypeTok{by =} \NormalTok{win_size *}\StringTok{ }\DecValTok{5}\NormalTok{)}
+\NormalTok{x_breaks <-}\StringTok{ }\KeywordTok{seq}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{41}\NormalTok{, }\DataTypeTok{by =} \DecValTok{5}\NormalTok{)}
+
+\NormalTok{sd_limits <-}\StringTok{ }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{ymax =} \NormalTok{win_mean +}\StringTok{ }\NormalTok{win_sd, }\DataTypeTok{ymin =} \NormalTok{win_mean -}\StringTok{ }\NormalTok{win_sd)}
+
+\NormalTok{p <-}\StringTok{ }\KeywordTok{ggplot}\NormalTok{(res, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x =} \NormalTok{.win_id, }\DataTypeTok{y =} \NormalTok{win_mean)) +}
+\StringTok{  }\KeywordTok{geom_point}\NormalTok{(}\DataTypeTok{size =} \FloatTok{0.25}\NormalTok{) +}\StringTok{ }\KeywordTok{geom_pointrange}\NormalTok{(sd_limits, }\DataTypeTok{size =} \FloatTok{0.1}\NormalTok{) +}\StringTok{ }
+\StringTok{  }\KeywordTok{scale_x_continuous}\NormalTok{(}\DataTypeTok{labels =} \NormalTok{x_labels, }\DataTypeTok{breaks =} \NormalTok{x_breaks) +}\StringTok{ }
+\StringTok{  }\KeywordTok{xlab}\NormalTok{(}\StringTok{"Position (bp from TSS)"}\NormalTok{) +}\StringTok{ }\KeywordTok{ylab}\NormalTok{(}\StringTok{"Signal"}\NormalTok{) +}\StringTok{ }
+\StringTok{  }\KeywordTok{theme_classic}\NormalTok{()}
 \end{Highlighting}
@@
 
-\NormalTok{repeats <-}\StringTok{ }\KeywordTok{read_bed}\NormalTok{(}\KeywordTok{valr_example}\NormalTok{(}\StringTok{'hg19.rmsk.chr22.bed.gz'}\NormalTok{), }\DataTypeTok{n_fields =} \DecValTok{6}\NormalTok{) }
-\NormalTok{genome <-}\StringTok{ }\KeywordTok{read_genome}\NormalTok{(}\KeywordTok{valr_example}\NormalTok{(}\StringTok{'hg19.chrom.sizes.gz'}\NormalTok{))}
-\NormalTok{genes <-}\StringTok{ }\KeywordTok{read_bed12}\NormalTok{(}\KeywordTok{valr_example}\NormalTok{(}\StringTok{'hg19.refGene.chr22.bed.gz'}\NormalTok{))}
+\NormalTok{repeats <-}\StringTok{ }\KeywordTok{read_bed}\NormalTok{(}\KeywordTok{valr_example}\NormalTok{(}\StringTok{"hg19.rmsk.chr22.bed.gz"}\NormalTok{), }\DataTypeTok{n_fields =} \DecValTok{6}\NormalTok{) }
+\NormalTok{genome <-}\StringTok{ }\KeywordTok{read_genome}\NormalTok{(}\KeywordTok{valr_example}\NormalTok{(}\StringTok{"hg19.chrom.sizes.gz"}\NormalTok{))}
+\NormalTok{genes <-}\StringTok{ }\KeywordTok{read_bed12}\NormalTok{(}\KeywordTok{valr_example}\NormalTok{(}\StringTok{"hg19.refGene.chr22.bed.gz"}\NormalTok{))}
 \CommentTok{# convert bed12 to bed with exons}
@@
   \KeywordTok{replicate}\NormalTok{(n, }\KeywordTok{bed_shuffle}\NormalTok{(.data, genome), }\DataTypeTok{simplify =} \OtherTok{FALSE}\NormalTok{) %>%}
-\StringTok{    }\KeywordTok{bind_rows}\NormalTok{(}\DataTypeTok{.id =} \StringTok{'rep'}\NormalTok{) %>%}
+\StringTok{    }\KeywordTok{bind_rows}\NormalTok{(}\DataTypeTok{.id =} \StringTok{"rep"}\NormalTok{) %>%}
 \StringTok{    }\KeywordTok{group_by}\NormalTok{(rep) %>%}\StringTok{ }\KeywordTok{nest}\NormalTok{()}
@@
 \CommentTok{#>    <chr>                <list>        <dbl>}
-\CommentTok{#>  1     1 <tibble [10,000 x 6]> 0.0004337515}
-\CommentTok{#>  2     2 <tibble [10,000 x 6]> 0.0003501136}
-\CommentTok{#>  3     3 <tibble [10,000 x 6]> 0.0006080318}
-\CommentTok{#>  4     4 <tibble [10,000 x 6]> 0.0005768465}
-\CommentTok{#>  5     5 <tibble [10,000 x 6]> 0.0002537739}
-\CommentTok{#>  6     6 <tibble [10,000 x 6]> 0.0004480599}
-\CommentTok{#>  7     7 <tibble [10,000 x 6]> 0.0001777942}
-\CommentTok{#>  8     8 <tibble [10,000 x 6]> 0.0003448734}
-\CommentTok{#>  9     9 <tibble [10,000 x 6]> 0.0003303327}
-\CommentTok{#> 10    10 <tibble [10,000 x 6]> 0.0003312376}
+\CommentTok{#>  1     1 <tibble [10,000 x 6]> 0.0003388967}
+\CommentTok{#>  2     2 <tibble [10,000 x 6]> 0.0004965988}
+\CommentTok{#>  3     3 <tibble [10,000 x 6]> 0.0002974843}
+\CommentTok{#>  4     4 <tibble [10,000 x 6]> 0.0006899870}
+\CommentTok{#>  5     5 <tibble [10,000 x 6]> 0.0004678412}
+\CommentTok{#>  6     6 <tibble [10,000 x 6]> 0.0001726937}
+\CommentTok{#>  7     7 <tibble [10,000 x 6]> 0.0004694941}
+\CommentTok{#>  8     8 <tibble [10,000 x 6]> 0.0004660410}
+\CommentTok{#>  9     9 <tibble [10,000 x 6]> 0.0006846643}
+\CommentTok{#> 10    10 <tibble [10,000 x 6]> 0.0002143829}
 \CommentTok{#> # ... with 90 more rows}

### 2017-06-20 inst/paper/main.tex
fixed quotes and a few minor changes

@@
 %% Default: numerical citations
-\usepackage[numbers]{natbib}
+%\usepackage[numbers]{natbib}
 
 %% Uncomment this lines for superscript citations instead
-% \usepackage[super]{natbib}
+\usepackage[super]{natbib}
 
@@
 
-\title{valr: Reproducible genome interval analysis in R}
+\title{valr: Reproducible Genome Interval Analysis in R}
 \author[1]{Kent A. Riemondy}
+\author[2]{Ryan M. Sheridan}
 \author[1]{Austin Gillen}
-\author[2]{Ryan M. Sheridan}
 \author[2]{Yinni Yu}
@@
 
-	A routine bioinformatic task is the analysis of the relationships between sets of genomic intervals including the identification of DNA variants within genetic regions, annotation of regions enriched for nucleic acid binding proteins, and computation of read density within a set of exons.
-Command-line tools for interval analysis such as BEDtools \cite{quinlan_bedtools:_2010} and BEDOPS \cite{neph_bedops:_2012} enable analyses of genome-wide datasets and are keys components of analysis pipelines.
+	A routine bioinformatic task is the analysis of the relationships between sets of genomic intervals including the identification of DNA variants within protein coding regions, annotation of regions enriched for nucleic acid binding proteins, and computation of read density within a set of exons.
+Command-line tools for interval analysis such as \texttt{BEDtools} \cite{quinlan_bedtools:_2010} and \texttt{BEDOPS} \cite{neph_bedops:_2012} enable analyses of genome-wide datasets and are key components of analysis pipelines.
 Analyses with these tools commonly combine processing intervals on the command-line with visualization and statistical analysis in R.
-However, the need to master both the command-line and R hinders exploratory data analysis, and the development reproducible research workflows built in the RMarkdown framework.
+However, the need to master both the command-line and R hinders exploratory data analysis, and the development of reproducible research workflows built in the RMarkdown framework.
 
-	Existing R packages developed for interval analysis include IRanges \cite{lawrence_software_2013}, bedr \cite{haider_bedr_2016}, and GenometriCorr \cite{favorov_exploring_2012}.
-IRanges is a Bioconductor package that provides  interval classes and methods to perform interval arithmetic and is used by many Bioconductor packages.
-bedr is a CRAN-distributed package that provides wrapper R functions to call the BEDTools, BedOPS, and tabix command-line utilities, providing out-of-memory support for interval analysis.
-Finally,  GenometriCorr provides a set of statistical tests to determine the relationships between interval sets using IRanges data structures.
-These packages provide  functionality for processing and statistical inference of interval data, but they do not easily integrate with interactive data visualization features developed for R including Shiny\cite{chang_shiny_2017}, and the RStudio IDE.
-Moreover, the S4 data structures implemented by IRanges do not easily interface with the recent advances provided by the popular tidyverse suite of data processing and visualization tools (e.g., dplyr, purrr, broom, and ggplot2) \cite{wickham_tidyverse_2017}.
-We sought to develop an flexible R package for genomic interval arithmetic built to incorporate new R programming, visualization, and interactivity features.
+	Existing R packages developed for interval analysis include \texttt{IRanges} \cite{lawrence_software_2013}, \texttt{bedr} \cite{haider_bedr_2016}, and \texttt{GenometriCorr} \cite{favorov_exploring_2012}.
+\texttt{IRanges} is a Bioconductor package that provides interval classes and methods to perform interval arithmetic and is used by many Bioconductor packages.
+\texttt{bedr} is a CRAN-distributed package that provides wrapper R functions to call the \texttt{BEDtools}, \texttt{BEDOPS}, and \texttt{tabix} command-line utilities, providing out-of-memory support for interval analysis.
+Finally, \texttt{ GenometriCorr} provides a set of statistical tests to determine the relationships between interval sets using \texttt{IRanges} data structures.
+These packages provide functionality for processing and statistical inference of interval data, however they require a detailed understanding of S4 classes (\texttt{IRanges}) or the installation of external command-line dependencies (\texttt{bedr}).
+Additionally, these packages do not easily integrate with the recent advances provided by the popular \texttt{tidyverse} suite of data processing and visualization tools (e.g.
+\texttt{dplyr}, \texttt{purrr}, \texttt{broom} and \texttt{ggplot2}) \cite{wickham_tidyverse_2017}.
+We therefore sought to develop a flexible R package for genomic interval arithmetic built to incorporate new R programming, visualization, and interactivity features.
 
@@
 \subsection*{Implementation}
-valr is an R package that makes extensive use of dplyr, a flexible and high-performance framework for data manipulation in R \cite{wickham_dplyr_2016}.
-Compute intensive functions in valr are written in C++ using Rcpp to enable fluid interactive analysis of large datasets\cite{eddelbuettel_rcpp_2011}.
-Interval intersections and related operations use an interval tree algorithm to efficiently search for overlapping intervals\cite{cormen_2001}.
-BED files are imported and handled in R as data\_frame objects, requiring minimal pre or post-processing to integrate with additional R packages or command line tools.
+\texttt{valr} is an R package that makes extensive use of \texttt{dplyr}, a flexible and high-performance framework for data manipulation in R \cite{wickham_dplyr_2016}.
+Additionally, compute intensive functions in \texttt{valr} are written in C++ using \texttt{Rcpp} to enable fluid interactive analysis of large datasets \cite{eddelbuettel_rcpp_2011}.
+Interval intersections and related operations use an interval tree algorithm to efficiently search for overlapping intervals \cite{cormen_2001}.
+BED files are imported and handled in R as \texttt{data\_frame} objects, requiring minimal pre or post-processing to integrate with additional R packages or command line tools.
 
 \subsection*{Operation}
-valr is distributed as part of the CRAN R package repository and is compatible with Mac OS X, Windows, and major Linux operating systems.
-Package dependencies and system requirements are documented in the valr CRAN repository (\url{https://cran.r-project.org/web/packages/valr/index.html}).
+\texttt{valr} is distributed as part of the CRAN R package repository and is compatible with Mac OS X, Windows, and major Linux operating systems.
+Package dependencies and system requirements are documented in the \href{https://cran.r-project.org/web/packages/valr/index.html}{\texttt{valr} CRAN repository}.
 
 \section*{Use Cases}
-To demonstrate the functionality and utility of \texttt{valr}, presented below is a basic tutorial for using valr and additional common use cases for genomic interval analysis.
+To demonstrate the functionality and utility of \texttt{valr}, presented below is a basic tutorial for using \texttt{valr} and additional common use cases for genomic interval analysis.
 
@@
 
-\texttt{valr} provides a set of functions to read BED, BEDgraph, and VCF formats into R as convenient \texttt{tibble} data\_frame objects.
+\texttt{valr} provides a set of functions to read BED, BEDgraph, and VCF formats into R as convenient \texttt{tibble} (tbl) \texttt{data\_frame} objects.
 All tbls have \texttt{chrom}, \texttt{start}, and \texttt{end} columns, and tbls from multi-column formats have additional pre-determined column names.
@@
 \texttt{read.table}, \texttt{readr::read\_tsv}) are also supported provided the constructed dataframes contain the requisite column names (\texttt{chrom}, \texttt{start}, \texttt{end}).
-Additionally \texttt{valr} supports connections to remote databases to programatically access the UCSC and Ensembl databases via the \texttt{db\_ucsc} and \texttt{db\_ensembl} functions.
+Additionally, \texttt{valr} supports connections to remote databases to access the UCSC and Ensembl databases via the \texttt{db\_ucsc} and \texttt{db\_ensembl} functions.
 
@@
 The functions in \texttt{valr} have similar names to their \texttt{BEDtools} counterparts, and so will be familiar to users of the \texttt{BEDtools} suite.
-Also, similar to pybedtools\cite{dale_pybedtools:_2011}, a python wrapper for \texttt{BEDTools}, \texttt{valr} has a terse syntax.
+Also, similar to pybedtools\cite{dale_pybedtools:_2011}, a python wrapper for \texttt{BEDtools}, \texttt{valr} has a terse syntax.
 For example, shown below is a demonstration of how to find all intergenic SNPs within 1 kilobase of genes using \texttt{valr}.
@@
 
-\NormalTok{snps <-}\StringTok{ }\KeywordTok{read_bed}\NormalTok{(}\KeywordTok{valr_example}\NormalTok{(}\StringTok{'hg19.snps147.chr22.bed.gz'}\NormalTok{), }\DataTypeTok{n_fields =} \DecValTok{6}\NormalTok{)}
-\NormalTok{genes <-}\StringTok{ }\KeywordTok{read_bed}\NormalTok{(}\KeywordTok{valr_example}\NormalTok{(}\StringTok{'genes.hg19.chr22.bed.gz'}\NormalTok{), }\DataTypeTok{n_fields =} \DecValTok{6}\NormalTok{)}
+\NormalTok{snps <-}\StringTok{ }\KeywordTok{read_bed}\NormalTok{(}\KeywordTok{valr_example}\NormalTok{(}\StringTok{"hg19.snps147.chr22.bed.gz"}\NormalTok{), }\DataTypeTok{n_fields =} \DecValTok{6}\NormalTok{)}
+\NormalTok{genes <-}\StringTok{ }\KeywordTok{read_bed}\NormalTok{(}\KeywordTok{valr_example}\NormalTok{(}\StringTok{"genes.hg19.chr22.bed.gz"}\NormalTok{), }\DataTypeTok{n_fields =} \DecValTok{6}\NormalTok{)}
 
@@
 \NormalTok{nearby %>%}
-\StringTok{  }\KeywordTok{select}\NormalTok{(}\KeywordTok{starts_with}\NormalTok{(}\StringTok{'name'}\NormalTok{), .overlap, .dist) %>%}
+\StringTok{  }\KeywordTok{select}\NormalTok{(}\KeywordTok{starts_with}\NormalTok{(}\StringTok{"name"}\NormalTok{), .overlap, .dist) %>%}
 \StringTok{  }\KeywordTok{filter}\NormalTok{(}\KeywordTok{abs}\NormalTok{(.dist) <}\StringTok{ }\DecValTok{1000}\NormalTok{)}
@@
 To aid in demonstrating the interval operations available in \texttt{valr}, we developed the \texttt{bed\_glyph()} tool which produces plots demonstrating the input and output of operations in \texttt{valr} in a manner similar to those found in the \texttt{BEDtools} documentation.
-Shown below is the code required to produce glyphs displaying the results of intersecting \texttt{x} and \texttt{y} intervals with \texttt{bed\_intersect()} , and the result of merging \texttt{x} intervals with \texttt{bed\_merge()} (Figure 1 \ref{fig:Figure 1}).
+Shown below is the code required to produce glyphs displaying the results of intersecting \texttt{x} and \texttt{y} intervals with \texttt{bed\_intersect()} , and the result of merging \texttt{x} intervals with \texttt{bed\_merge()} (Figure \ref{fig:Figure 1}).
 
@@
   \NormalTok{~chrom, ~start, ~end,}
-  \StringTok{'chr1'}\NormalTok{, }\DecValTok{25}\NormalTok{,     }\DecValTok{50}\NormalTok{,}
-  \StringTok{'chr1'}\NormalTok{, }\DecValTok{100}\NormalTok{,    }\DecValTok{125}
+  \StringTok{"chr1"}\NormalTok{, }\DecValTok{25}\NormalTok{,     }\DecValTok{50}\NormalTok{,}
+  \StringTok{"chr1"}\NormalTok{, }\DecValTok{100}\NormalTok{,    }\DecValTok{125}
 \NormalTok{)}
@@
   \NormalTok{~chrom, ~start, ~end,}
-  \StringTok{'chr1'}\NormalTok{, }\DecValTok{30}\NormalTok{,     }\DecValTok{75}
+  \StringTok{"chr1"}\NormalTok{, }\DecValTok{30}\NormalTok{,     }\DecValTok{75}
 \NormalTok{)}
@@
   \NormalTok{~chrom, ~start, ~end,}
-  \StringTok{'chr1'}\NormalTok{,      }\DecValTok{1}\NormalTok{,      }\DecValTok{50}\NormalTok{,}
-  \StringTok{'chr1'}\NormalTok{,      }\DecValTok{10}\NormalTok{,     }\DecValTok{75}\NormalTok{,}
-  \StringTok{'chr1'}\NormalTok{,      }\DecValTok{100}\NormalTok{,    }\DecValTok{120}
+  \StringTok{"chr1"}\NormalTok{,      }\DecValTok{1}\NormalTok{,      }\DecValTok{50}\NormalTok{,}
+  \StringTok{"chr1"}\NormalTok{,      }\DecValTok{10}\NormalTok{,     }\DecValTok{75}\NormalTok{,}
+  \StringTok{"chr1"}\NormalTok{,      }\DecValTok{100}\NormalTok{,    }\DecValTok{120}
 \NormalTok{)}
@@
 \end{Highlighting}
-% * <jay.hesselberth@gmail.com> 2017-06-10T17:17:56.304Z:
-%
-% Is it possible to insert the figure inline in the document?
-% * KR
-% Yes, but they request that figures are listed at the end of the document.
-% Once the document is formatted by F1000 I think that the document will look similar to this article:
-% https://f1000research.com/articles/5-2122/v2
-% ^.
 
@@
 
-The \texttt{group\_by} function in dplyr can be used to perform functions
+The \texttt{group\_by} function in \texttt{dplyr} can be used to execute functions
 on subsets of single and multiple \texttt{data\_frame}s. Functions in
@@
   \NormalTok{~chrom, ~start, ~end, ~strand,}
-  \StringTok{'chr1'}\NormalTok{, }\DecValTok{1}\NormalTok{,      }\DecValTok{100}\NormalTok{,  }\StringTok{'+'}\NormalTok{,}
-  \StringTok{'chr1'}\NormalTok{, }\DecValTok{50}\NormalTok{,     }\DecValTok{150}\NormalTok{,  }\StringTok{'+'}\NormalTok{,}
-  \StringTok{'chr2'}\NormalTok{, }\DecValTok{100}\NormalTok{,    }\DecValTok{200}\NormalTok{,  }\StringTok{'-'}
+  \StringTok{"chr1"}\NormalTok{, }\DecValTok{1}\NormalTok{,      }\DecValTok{100}\NormalTok{,  }\StringTok{"+"}\NormalTok{,}
+  \StringTok{"chr1"}\NormalTok{, }\DecValTok{50}\NormalTok{,     }\DecValTok{150}\NormalTok{,  }\StringTok{"+"}\NormalTok{,}
+  \StringTok{"chr2"}\NormalTok{, }\DecValTok{100}\NormalTok{,    }\DecValTok{200}\NormalTok{,  }\StringTok{"-"}
 \NormalTok{)}
@@
   \NormalTok{~chrom, ~start, ~end, ~strand,}
-  \StringTok{'chr1'}\NormalTok{, }\DecValTok{50}\NormalTok{,     }\DecValTok{125}\NormalTok{,  }\StringTok{'+'}\NormalTok{,}
-  \StringTok{'chr1'}\NormalTok{, }\DecValTok{50}\NormalTok{,     }\DecValTok{150}\NormalTok{,  }\StringTok{'-'}\NormalTok{,}
-  \StringTok{'chr2'}\NormalTok{, }\DecValTok{50}\NormalTok{,     }\DecValTok{150}\NormalTok{,  }\StringTok{'+'}
+  \StringTok{"chr1"}\NormalTok{, }\DecValTok{50}\NormalTok{,     }\DecValTok{125}\NormalTok{,  }\StringTok{"+"}\NormalTok{,}
+  \StringTok{"chr1"}\NormalTok{, }\DecValTok{50}\NormalTok{,     }\DecValTok{150}\NormalTok{,  }\StringTok{"-"}\NormalTok{,}
+  \StringTok{"chr2"}\NormalTok{, }\DecValTok{50}\NormalTok{,     }\DecValTok{150}\NormalTok{,  }\StringTok{"+"}
 \NormalTok{)}
@@
 \subsubsection*{API}\label{api}
-The major functions available in valr are shown in Table \ref{tab:tools}.
+The major functions available in \texttt{valr} are shown in Table \ref{tab:tools}.
 
@@
 \begin{Highlighting}[]
-\CommentTok{# `valr_example()` identifies the path of example files in the valr package}
-\NormalTok{bedfile <-}\StringTok{ }\KeywordTok{valr_example}\NormalTok{(}\StringTok{'genes.hg19.chr22.bed.gz'}\NormalTok{)}
-\NormalTok{genomefile <-}\StringTok{ }\KeywordTok{valr_example}\NormalTok{(}\StringTok{'hg19.chrom.sizes.gz'}\NormalTok{)}
-\NormalTok{bgfile  <-}\StringTok{ }\KeywordTok{valr_example}\NormalTok{(}\StringTok{'hela.h3k4.chip.bg.gz'}\NormalTok{)}
+\NormalTok{bedfile <-}\StringTok{ }\KeywordTok{valr_example}\NormalTok{(}\StringTok{"genes.hg19.chr22.bed.gz"}\NormalTok{)}
+\NormalTok{genomefile <-}\StringTok{ }\KeywordTok{valr_example}\NormalTok{(}\StringTok{"hg19.chrom.sizes.gz"}\NormalTok{)}
+\NormalTok{bgfile  <-}\StringTok{ }\KeywordTok{valr_example}\NormalTok{(}\StringTok{"hela.h3k4.chip.bg.gz"}\NormalTok{)}
 
@@
 \begin{Highlighting}[]
-\CommentTok{# generate 1 bp TSS intervals, `+` strand only}
+\CommentTok{# generate 1 bp TSS intervals, "+" strand only}
 \NormalTok{tss <-}\StringTok{ }\NormalTok{genes %>%}
-\StringTok{  }\KeywordTok{filter}\NormalTok{(strand ==}\StringTok{ '+'}\NormalTok{) %>%}
+\StringTok{  }\KeywordTok{filter}\NormalTok{(strand ==}\StringTok{ "+"}\NormalTok{) %>%}
 \StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{end =} \NormalTok{start +}\StringTok{ }\DecValTok{1}\NormalTok{)}
@@
 Finally, these summary statistics are used to construct a plot that
-illustrates histone density surrounding TSSs (Figure 2\ref{fig:Figure 2}).
+illustrates histone density surrounding TSSs (Figure \ref{fig:Figure 2}).
 
@@
 
-\KeywordTok{ggplot}\NormalTok{(res, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x =} \NormalTok{.win_id, }\DataTypeTok{y =} \NormalTok{win_mean)) +}
-\StringTok{  }\KeywordTok{geom_point}\NormalTok{() +}\StringTok{ }\KeywordTok{geom_pointrange}\NormalTok{(sd_limits) +}\StringTok{ }
+\NormalTok{p <-}\StringTok{ }\KeywordTok{ggplot}\NormalTok{(res, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x =} \NormalTok{.win_id, }\DataTypeTok{y =} \NormalTok{win_mean)) +}
+\StringTok{  }\KeywordTok{geom_point}\NormalTok{(}\DataTypeTok{size =} \FloatTok{0.25}\NormalTok{) +}\StringTok{ }\KeywordTok{geom_pointrange}\NormalTok{(sd_limits, }\DataTypeTok{size =} \FloatTok{0.1}\NormalTok{) +}\StringTok{ }
 \StringTok{  }\KeywordTok{scale_x_continuous}\NormalTok{(}\DataTypeTok{labels =} \NormalTok{x_labels, }\DataTypeTok{breaks =} \NormalTok{x_breaks) +}\StringTok{ }
-\StringTok{  }\KeywordTok{xlab}\NormalTok{(}\StringTok{'Position (bp from TSS)'}\NormalTok{) +}\StringTok{ }\KeywordTok{ylab}\NormalTok{(}\StringTok{'Signal'}\NormalTok{) +}\StringTok{ }
-\StringTok{  }\KeywordTok{ggtitle}\NormalTok{(}\StringTok{'Human H3K4me3 signal near transcription start sites'}\NormalTok{) +}
+\StringTok{  }\KeywordTok{xlab}\NormalTok{(}\StringTok{"Position (bp from TSS)"}\NormalTok{) +}\StringTok{ }\KeywordTok{ylab}\NormalTok{(}\StringTok{"Signal"}\NormalTok{) +}\StringTok{ }
 \StringTok{  }\KeywordTok{theme_classic}\NormalTok{()}
@@
 \begin{Highlighting}[]
-\KeywordTok{library}\NormalTok{(tidyverse, }\DataTypeTok{warn.conflicts =} \NormalTok{F)}
+\KeywordTok{library}\NormalTok{(tidyverse)}
 
-\NormalTok{repeats <-}\StringTok{ }\KeywordTok{read_bed}\NormalTok{(}\KeywordTok{valr_example}\NormalTok{(}\StringTok{'hg19.rmsk.chr22.bed.gz'}\NormalTok{), }\DataTypeTok{n_fields =} \DecValTok{6}\NormalTok{) }
-\NormalTok{genome <-}\StringTok{ }\KeywordTok{read_genome}\NormalTok{(}\KeywordTok{valr_example}\NormalTok{(}\StringTok{'hg19.chrom.sizes.gz'}\NormalTok{))}
-\NormalTok{genes <-}\StringTok{ }\KeywordTok{read_bed12}\NormalTok{(}\KeywordTok{valr_example}\NormalTok{(}\StringTok{'hg19.refGene.chr22.bed.gz'}\NormalTok{))}
+\NormalTok{repeats <-}\StringTok{ }\KeywordTok{read_bed}\NormalTok{(}\KeywordTok{valr_example}\NormalTok{(}\StringTok{"hg19.rmsk.chr22.bed.gz"}\NormalTok{), }\DataTypeTok{n_fields =} \DecValTok{6}\NormalTok{) }
+\NormalTok{genome <-}\StringTok{ }\KeywordTok{read_genome}\NormalTok{(}\KeywordTok{valr_example}\NormalTok{(}\StringTok{"hg19.chrom.sizes.gz"}\NormalTok{))}
+\NormalTok{genes <-}\StringTok{ }\KeywordTok{read_bed12}\NormalTok{(}\KeywordTok{valr_example}\NormalTok{(}\StringTok{"hg19.refGene.chr22.bed.gz"}\NormalTok{))}
 \CommentTok{# convert bed12 to bed with exons}
@@
   \KeywordTok{replicate}\NormalTok{(n, }\KeywordTok{bed_shuffle}\NormalTok{(.data, genome), }\DataTypeTok{simplify =} \OtherTok{FALSE}\NormalTok{) %>%}
-\StringTok{    }\KeywordTok{bind_rows}\NormalTok{(}\DataTypeTok{.id =} \StringTok{'rep'}\NormalTok{) %>%}
+\StringTok{    }\KeywordTok{bind_rows}\NormalTok{(}\DataTypeTok{.id =} \StringTok{"rep"}\NormalTok{) %>%}
 \StringTok{    }\KeywordTok{group_by}\NormalTok{(rep) %>%}\StringTok{ }\KeywordTok{nest}\NormalTok{()}
 \NormalTok{\}}
+
 \NormalTok{nreps <-}\StringTok{ }\DecValTok{100}
+
 \NormalTok{shuffled <-}\StringTok{ }\KeywordTok{shuffle_intervals}\NormalTok{(}\DataTypeTok{n =} \NormalTok{nreps, repeats, genome) %>%}
@@
 \subsection*{Benchmarking against bedtools}
-In order to ensure that valr performs fast enough to enable interactive analysis, key functionality is implemented in C++.
-To test the speed of major valr functions we generated two data\_frames containing 1 million random 1 kilobase intervals derived from the human genome.
-Most of the major valr functions complete execution in less than 1 second, demonstrating that valr can process large interval datasets efficiently (Figure 3A\ref{fig:Figure 3}).
+In order to ensure that \texttt{valr} performs fast enough to enable interactive analysis, key functionality is implemented in C++.
+To test the speed of major \texttt{valr} functions we generated two \texttt{data\_frames} containing 1 million randomly selected 1 kilobase intervals derived from the human genome.
+Most of the major \texttt{valr} functions complete execution in less than 1 second, demonstrating that \texttt{valr} can process large interval datasets efficiently (Figure \ref{fig:Figure 3}{A}).
 
-We also benchmarked major valr functions against corresponding commands in BEDTools.
-valr operates on data\_frames already loaded into RAM, whereas BEDTools performs file-reading, processing, and writing.
-To compare valr against BEDTools we generated two BED files containing 1 million random 1 kilobase intervals derived from the human genome.
-For valr functions, we timed reading the table into R (e.g.
-with read\_bed()) and performing the respective function.
-For BEDTools commands we timed executing the command with the output written to /dev/null.
-valr functions performed similarly or faster than BEDTools commands with the exception of the \texttt{bed\_map} and \texttt{bed\_fisher} (Figure 3B\ref{fig:Figure 3}).
+We also benchmarked major \texttt{valr} functions against corresponding commands in \texttt{BEDtools}.
+\texttt{valr} operates on \texttt{data\_frames} already loaded into RAM, whereas \texttt{BEDtools} performs file-reading, processing, and writing.
+To compare \texttt{valr} against \texttt{BEDtools} we generated two BED files containing 1 million randomly selected 1 kilobase intervals derived from the human genome.
+For \texttt{valr} functions, we timed reading the table into R (e.g.
+with \texttt{read\_bed()}) and performing the respective function.
+For \texttt{BEDtools} commands we timed executing the command with the output written to \texttt{/dev/null}.
+\texttt{valr} functions performed similarly or faster than \texttt{BEDtools} commands with the exception of \texttt{bed\_map} and \texttt{bed\_fisher} (Figure \ref{fig:Figure 3}{B}).
 
-\subsection{Reproducible Reports and Interactive Visualizations}
+\subsection*{Reproducible Reports and Interactive Visualizations}
 Command-line tools like \texttt{BEDtools} and \texttt{bedops} can be incorporated into reproducible workflows (e.g., with
@@
 
-Additionally, \texttt{valr} seamlessly integrates into R shiny applications allowing for complex interactive visualizations relating to genomic interval analyses.
-We have developed a shiny application available on github \href{https://github.com/rnabioco/valrdata} that explores ChiP-Seq signal density surrounding transcription start sites and demonstrates the ease of implementing \texttt{valr} to power dynamic visualizations.
+Additionally, \texttt{valr} seamlessly integrates into R \texttt{shiny}{\cite{chang_shiny_2017}} applications allowing for complex interactive visualizations relating to genomic interval analyses.
+We have developed a \texttt{shiny} application (available on \href{https://github.com/rnabioco/valrdata}{\texttt{github}}) that explores ChiP-Seq signal density surrounding transcription start sites and demonstrates the ease of implementing \texttt{valr} to power dynamic visualizations.
 
 \section*{Summary} % Optional - only if NO new datasets are included
-valr provides a flexible framework for interval arithmetic in R/Rstudio.
-valr functions are written with a simple and terse syntax that promotes flexible interactive analysis.
-Additionally by providing an easy-to-use interface for interval arithmetic in R, valr is also a useful teaching tool to introduce the analyses necessary to investigate correlations between genomic intervals, without requiring familiarity with the command-line.
+\texttt{valr} provides a flexible framework for interval arithmetic in R/Rstudio.
+\texttt{valr} functions are written with a simple and terse syntax that promotes flexible interactive analysis.
+Additionally by providing an easy-to-use interface for interval arithmetic in R, \texttt{valr} is also a useful teaching tool to introduce the analyses necessary to investigate correlations between genomic intervals, without requiring familiarity with the command-line.
 We envision that \texttt{valr} will help researchers quickly and reproducibly analyze genome interval datasets.
@@
 \begin{enumerate}
-\item URL valr can be installed via CRAN using "install.packages('valr')".
-\item valr is maintained at http://github.com/rnabioco/valr
-\item valr source code is available at http://github.com/rnabioco/valr
+\item \texttt{valr} can be installed via CRAN using \texttt{install.packages('valr')}.
+\item \texttt{valr} is maintained at http://github.com/rnabioco/valr
+\item \texttt{valr} source code is available at http://github.com/rnabioco/valr
 \item Latest stable version of source code is at: https://github.com/rnabioco/valr/archive/v0.3.0.tar.gz
@@
 
-JR conceptualized the package and directed software development.
-JR, KR, AG, and RS developed and tested the package.
+JH conceptualized the package and directed software development.
+JH, KR, AG, and RS developed and tested the package.
 YY and CB contributed workflows and benchmarks for the package.
@@
 \includegraphics[width=1\textwidth]{figure1.pdf}
-\caption{\label{fig:Figure 1}Visualizing interval operations in valr with \texttt{bed\_glyph()}.}
+\caption{\label{fig:Figure 1}Visualizing interval operations in \texttt{valr} with \texttt{bed\_glyph()}.}
 \end{figure}
@@
 \textnormal{(A) Summarized coverage of human H3K4Me3 Chip-Seq coverage across positive strand transcription start sites on chromosome 22.
-Data presented +/- SEM.}}
+Data presented +/- SD.}}
 \end{figure}
@@
 \includegraphics[width=1\textwidth]{figure3.pdf}
-\caption{\label{fig:Figure 3} Performance of major valr functionality.
+\caption{\label{fig:Figure 3} Performance of major \texttt{valr} functionality.
 \textnormal{(A)  Timings calculated by performing 10 repetitions of indicated functions on dataframes preloaded in R containing 1 million random 1 kilobase x/y intervals generated using \texttt{bed\_random()}.
-(B) Timings for executing functions in BEDTools or equivalent functions in valr using the same interval sets as in (A) written to files.
-All BEDTools function outputs were written to /dev/null/, and were timed using GNU \texttt{time}.
-Timings for valr functions in (B) include times for reading files using \texttt{read\_bed()} functions and were timed using the \texttt{microbenchmark} package.}}
+(B) Timings for executing functions in \texttt{BEDtools v2.25.0} or equivalent functions in \texttt{valr} using the same interval sets as in (A) written to files.
+All \texttt{BEDtools} function outputs were written to \texttt{/dev/null/}, and were timed using GNU \texttt{time}.
+Timings for \texttt{valr} functions in (B) include times for reading files using \texttt{read\_bed()} functions and were timed using the \texttt{microbenchmark} package.}}
 \end{figure}
@@
 \hrule \vspace{0.1cm}
-\caption{\label{tab:tools} An overview of major functions available in valr.}
+\caption{\label{tab:tools} An overview of major functions available in \texttt{valr}.}
 \centering
